### 2020-06-18 11-25/Presentation.tex
Initial Overleaf Import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}

\mode<presentation> {

% The Beamer class comes with a number of default slide themes
% which change the colors and layouts of slides. Below this is a list
% of all the themes, uncomment each in turn to see what they look like.

%\usetheme{default}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
\usetheme{Boadilla}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

% As well as themes, the Beamer class has a number of color themes
% for any slide theme. Uncomment each of these in turn to see how it
% changes the colors of your current slide theme.

%\usecolortheme{albatross}
%\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{orchid}
%\usecolortheme{rose}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{whale}
\usecolortheme{wolverine}

%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\setbeamertemplate{bibliography item}{\insertbiblabel}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
%\usepackage {tikz}
\usepackage{tkz-graph}
\usepackage{bibentry}

\usepackage[
backend=biber,
style=numeric,
sorting=ynt
]{biblatex}

\documentclass[xcolor=table]{beamer}

\bibliography{bibfile}

\graphicspath{ {images/} }

\GraphInit[vstyle = Shade]
\tikzset{
  LabelStyle/.style = { rectangle, rounded corners, draw,
                        minimum width = 2em, fill = yellow!50,
                        text = red, font = \bfseries },
  VertexStyle/.append style = { inner sep=5pt,
                                font = \normalsize\bfseries},
  EdgeStyle/.append style = {->, bend left} }
\usetikzlibrary {positioning}
%\usepackage {xcolor}
\definecolor {processblue}{cmyk}{0.96,0,0,0}

% show sections only
\setcounter{tocdepth}{1}
% show subsections
%\setcounter{tocdepth}{2}
% show subsubsections
%\setcounter{tocdepth}{3}
\AtBeginSection[]
{
    \begin{frame}
        \frametitle{Table of Contents}
        \tableofcontents[currentsection]
    \end{frame}
}

%\AtBeginSubsection[]
%{
%    \begin{frame}
%        \frametitle{Table of Contents}
%        \tableofcontents[currentsection,currentsubsection]
%    \end{frame}
%}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Short title]{Thesis Pre-final presentation} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Azcarraga, Araya} % Your name
\institute[Department of Computer Science, University of the Philippine - Diliman] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
University of the Philippine - Diliman\\ % Your institution for the title page
\medskip
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}
\frametitle{Overview} % Table of contents slide, comment this block out to remove it
\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
\section{Introduction}

% to do sa intro
% list all jargon
% explain most jargon
% iklian bio part
% pahabain tech part esp p2p
% explain CDN

    %%%%%%%%%%%%%%%%%%% BIO SLIDES %%%%%%%%%%%%%%%%%%%%%%%%%%

    \subsection{Definition} %keep daw
    \begin{frame}{Human Karyotype}
        \centering
\includegraphics[scale=1]{karyotype.jpg}
    \end{frame}
    \begin{frame}{Definition of Genomics}
        According to the National Institute of Health \textit{Specifically for the Human Genome Project}
                \begin{block} {\textbf{Definition of Genomics}}
        Genomics is the study of all of a person's genes (the genome), including interactions of those genes with each other and with the person's environment. \cite{genomics-definition}
                \end{block}
    The genomics definition we will use is for \textbf{all} organisms
    \end{frame}

    \subsection{Introduction}
    \begin{frame}{Introduction}
        \begin{itemize}
            \item DNA comprises of the components A,T,G,C
            \begin{itemize}
                \item \textbf{A} - Adenine
                \item \textbf{T} - Thymine
                \item \textbf{G} - Guanine
                \item \textbf{C} - Cytosine
            \end{itemize}
            \item There needs to be a lot of these components to describe complex organisms
            \begin{itemize}
                \item The full human genome contains around $3.2 x 10^9$ base pairs \cite[p.~4]{introgenomics}
            \end{itemize}
        \end{itemize}
        \centering
        \includegraphics[scale=0.1]{gene-base-pair.png}
    \end{frame}

    \subsection{Genomics Data}
    \begin{frame}{Genomics Data}
            \textbf{Data Type}
    \begin{itemize}
        \item Store DNA sequences digitally
        \item A simple data file format is called FASTA format
            \begin{itemize}
                \item starts with $>$
                \item then can add keywords and attributes separated by $|$
            \end{itemize}
    \end{itemize}
    \begin{block}{Sample.FASTA}
    $>$AB000263 $|$acc=AB000263$|$descr=Homo sapiens mRNA for prepro cortistatin like peptide, complete cds.$|$len=70
ACAAGATGCCATTGTCCCCCGGCCTCCTGCTGCTGCTGCTCTCCGGGGCCACGGCCACCGCTGCCCTGCC

    \end{block}
    \end{frame}

    \subsection{Sequencing}
    \begin{frame}{Sequencing}
        \textbf{Sequencing}
    \begin{itemize}
        \item Sequencers are machines used to look at the DNA of a specimen (A,G,C,T)
        \begin{itemize}
            \item Sequencers look at the exact order of a single strand \cite{genomics-definition}. 
        \end{itemize}
        \item DNA is \textbf{replicated} and fragmented before processing
        \item Sequencers then reads (or sequences) them
        \item Output of the read fragments is contained in a .FASTA file
    \end{itemize}
            \centering
        \includegraphics[scale=0.2]{sequencing.png}
    \end{frame}

    \begin{frame}{Next Generation Sequencing}
    (2003) The first human genome took ten years of work and $3,000,000,000$ \$ to finish.
    \begin{itemize}
        \item (2019) Currently the technology today in Japan can make 10,000 human genomes per day (Timewise: 36.5 million times faster) \cite[p.~19]{introgenomics}
    \end{itemize}
    \textbf{Next Generation Sequencers}
        \begin{itemize}
        \item Can read more faster, and cheaper
        \item More data
        \item Technology will get faster, there needs to be systems in preparation for that
    \end{itemize}
    \end{frame}
    %%%%%%%%%%%%%%%%%%% TECH SLIDES %%%%%%%%%%%%%%%%%%%%%%%%%%

    \subsection{Data Storage}
    \begin{frame}{Data Storage}
        A \textbf{database} (\textbf{DBMS}, Database Management System) is defined by 
        \begin{block} {\textbf{Definition}}

        DBMS contains information about a particular enterprise
        \begin{itemize}
            \item Collection of interrelated data
            \item Set of programs to access the data 
            \item An environment that is both convenient and efficient to use
        \end{itemize}
        \cite{Silberschatz2010}
        \end{block}
    \end{frame}

    \subsection{Database Principles}
    \begin{frame}{Database Principles}
        \begin{itemize}
            \item \textbf{Principles} to remember when creating a database (CAP). Idea is that only 2 of the three can be focused on. \cite[Ch.~19]{Silberschatz2010}
            \begin{itemize}
                \item \textbf{Consistency}
                \begin{itemize}
                    \item Update one part, will all other (replicated) parts be updated quickly as well
                    \item Data is accurate and copied faithfully
                \end{itemize}
                \item \textbf{Accessibility}
                \begin{itemize}
                    \item How easy it is to access data
                    \item API protocols
                \end{itemize}
                \item  \textbf{Partition Tolerance}
                \begin{itemize}
                    \item If one portion breaks, the other portions are still usable
                    \centering 
                    \includegraphics[scale=0.2]{partition_tolerance.png}
                    \end{itemize} 
             \end{itemize}
        \end{itemize}
    \end{frame}

    \subsection{Special Database}
    \begin{frame}{Special Databases}
        \textbf{Large Data}
        \begin{itemize}
            \item \textbf{Distributed Database}
            \begin{itemize}
                \item A distributed database system consists of loosely coupled sites that share no physical component             
                \item Database systems that run on each site are independent of each other
                \item Transactions may access data at one or more sites
                \cite[Ch.~19]{Silberschatz2010}
                \item Beneficial for large data with a certain speed requirement
            \end{itemize}
        \end{itemize}
    \end{frame}

    \begin{frame}{Special Databases}
        \textbf{Genomics Data}
        \texttt{needs a different set of features to easily conduct processes (alignment, assembly, comparison).}
        \begin{itemize}
            \item Need a special database to store all of these data
            \item Designed for more efficient and optimal use
            \item Has special needs for the users of genomic data
            \item Discuss in the RRL the different ways researchers approached this
        \end{itemize}
    \end{frame}

        \begin{frame}{Network Layers}
         \textbf{Networks Layers} how computers pass data to each other
        \begin{itemize}
            \item \begin{itemize}
            \item Level 1: Medium used to send signals (physical, electrical signals of 0 and 1), through wire or by air
            \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together different . Goal of sending a group of signals.
            \item Level 3: Introduction of networks: IP Address to uniquely identify entities. 
            \item Level 4: Management of error free data transmission: Sending a group of messages via TCP, UDP, FTP
            \item \textbf{Level 5}: Abstraction of Level 4: Function calls programmers will use to do level 4 processes.
            \end{itemize}
        \end{itemize}
    \end{frame}

    \subsection{Application Architecture}
    \begin{frame}{Application Architecture}
        \begin{itemize}
            \item The \textbf{application architecture} is how a networked app is structured across the various end systems (i.e. computers, servers, etc.) \cite{kurose}
            \item Falls into either client-server, peer-to-peer, or hybrid architecture
        \end{itemize}
    \end{frame}
        \subsubsection{Client-Server}
        \begin{frame}{Client-Server}
            \begin{itemize}
                \item A \textbf{client-server} architecture relies on a Web server that is always available and has a fixed IP address\\
                \item A single server connects to many clients
            \end{itemize}
            \centering
            \includegraphics[scale=0.5]{Server-based-network.png}
        \end{frame}
        \subsubsection{Peer-to-Peer \& Hybrid}
        \begin{frame}{Peer-to-Peer \& Hybrid}
            \begin{itemize}
                \item A \textbf{peer-to-peer (P2P)} architecture has minimal to no reliance on always-on, dedicated servers
                \item Peers are desktops/laptops that communicate to each other without passing through a dedicated server
                \item A \textbf{hybrid} architecture combines elements of P2P and client-server architectures, attempting to utilize advantages of both
            \end{itemize}
            \centering
            \includegraphics[scale=0.5]{P2P-network.png}
        \end{frame}
        \begin{frame}{BitTorrent terminology}
            \begin{itemize} % siguro breeze through what these mean 
                \item \textbf{Peers} are users that connect to the BitTorrent network.
                \item \textbf{Torrent files} contain metadata about files \& folders, trackers, and hash values.
                \item \textbf{Trackers} are servers that keep track of peers and availability of files
                \item \textbf{Seeds} are peers which have downloaded the entire file and are making it available
            \end{itemize}
        \end{frame}

\section{Review of Related Literature}
    \subsection{BioTorrents}
    \begin{frame}{BioTorrents}
        % outline
        \begin{itemize}
        \item System for legally sharing scientific data that works on top of the BitTorrent protocol. It is much like a public tracker.
        \item The main website hosts .torrent and metadata files
        \item Info about each dataset includes categories, license, filenames, etc. that helps users in searching for relevant datasets
        \item Torrent file has list of trackers, servers that "know" which peers are serving which files.
        \item Torrent client downloads from multiple peers simultaneously.
        \item To upload a torrent, the client makes a .torrent file using the torrenting software, then uploads it to the main website along with metadata. The client must have their computer continuously turned on to keep the file available as a peer. \cite{biotorrents}
        \end{itemize}
    \end{frame}

    \begin{frame}{BioTorrents}
        \begin{itemize}
            \item As with all public trackers, BioTorrents requires "good souls" that will keep many torrents available as often as possible.
            \item The datasets must be moderated to prevent abuse.
            \item The BioTorrents website is no longer up.
        \end{itemize}
    \end{frame}

    \subsection{PeerDB}
    \begin{frame}{PeerDB}
        \begin{itemize}
            \item Proposed P2P system for sharing general data
            \item Each node is equipped with a database
            \item Queries are passed around neighbors, with a maximum number of hops
            \item Caches information about other peers to speed up queries
            \item Large number of nodes & relatively small number of lookup servers
            \cite{peerdb}
        \end{itemize}
    \end{frame}

    \subsection{SeqTorr}
    \begin{frame}{SeqTorr}
    \textbf{SeqTorr} \cite{seqtorr}
    \begin{itemize}
        \item Utilizes distributed database
        \item Master node
        \begin{itemize}
            \item containing metadata of all the sequences
            \item where user authentication is handled
        \end{itemize}
        \item Data node
        \begin{itemize}
            \item contain a portion of all the sequences
            \item needs to replicate a sequence \texttt{r} times before a sequence is considered available
            \item proximity of data nodes can increase download speed but isn't true P2P
        \end{itemize}
        \item Defined API Implementation
        \begin{itemize}
            \item POST
            \item GET
            \item PUT
            \item DELETE
        \end{itemize}

    \end{itemize}
    \end{frame}

    \begin{frame}{SeqTorr}
        Problems:
        \begin{itemize}
            \item program not currently available
            \begin{itemize}
                \item cannot measure any statistics on this
            \end{itemize}
            \item not Peer2Peer
        \end{itemize}
    \end{frame}

    \begin{frame}{Table}
   \begin{table}[]
\begin{tabular}{|l|l|l|l|l|}
\hline
System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ \hline
BioTorrents & P2P           & No         & Depends                    & Bio-related  \\ \hline
SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ \hline
PeerDB      & P2P           & No         & Depends                    & General      \\ \hline
NCBI        & CS            & Yes        & Slow (US server)           & Bio-related  \\ \hline
\end{tabular}
\end{table} 
    \end{frame}

% NCBI Biosystems
% pros
% cons

% BitTorrious
% pros
% \cite{bittorious}
% cons

\section{Problem}
\begin{frame}{Problem}
The different gaps or problems found with other implementations

\begin{itemize}
    \item Genome researchers download entire database of genomic data
    \item Genome researchers only need a portion of the database, but end up downloading all
    \item Client-Server approach can face bottlenecks (server speed)
    \item Code has no documentation, thus harder to maintain
%% add more

\end{itemize}
\end{frame}

\section{Objectives}
\begin{frame}{Objectives}
Based on the problems we found earlier, we propose the following objectives
    \begin{itemize}
        \item Create information system that speeds up the download and upload of genomic data by at least 3x compared to existing systems
        \item The system should have the main API's
        \begin{itemize}
            \item POST genome data
            \item GET genome data (implement hybrid P2P)
            \item PUT genome metadata (optional: PUT data)
            \item DELETE genome data
        \end{itemize}
        \item Write documentation for the system
    \end{itemize}
\end{frame}

\section{Scope}
\begin{frame}{Scope}
The current scope and coverage of the project is only
\begin{itemize}
    \item Security and Authentication is not part of the scope
    \item Use only FASTA file format 
    \begin{itemize}
        \item DNA nucleotides only
    \end{itemize}
    \item UI will be purely for functional purposes (very basic design)
\end{itemize}
\end{frame}

\section{Theoretical Framework}

\begin{frame}{Theoretical Framework: System Description}
    \begin{itemize}
        \item A \textbf{master node} stores metadata of all sequences, may or may not contain sequence files
        \item \textbf{Data nodes} stores sequence files & metadata of all sequence files in its filesystem
        \item Both master node & data nodes have user interface that allow users to upload, search, and download sequence files
    \end{itemize}
\end{frame}
% DRAW.IO FILE: https://drive.google.com/file/d/1Nbp0Kmgcr2BZ13LBT3SGLMBoA6JA_Eth/view
\begin{frame}{Theoretical Framework: Schematics}
\textbf{App network diagram} \\ \medskip
\centering
\includegraphics[scale=0.5]{thesis1.png}
\end{frame}

\begin{frame}{Theoretical Framework: Schematics}
\textbf{POST: Uploading a sequence to the master node} \\ \medskip
\centering
\includegraphics[scale=0.5]{thesis3.png}
\end{frame}

\begin{frame}{Theoretical Framework: Schematics}
\textbf{POST: Uploading a sequence to a data node} \\ \medskip
\centering
\includegraphics[scale=0.5]{thesis2.png}
\end{frame}

\begin{frame}{Theoretical Framework: Schematics}
\textbf{GET: Downloading a sequence} \\ \medskip
\centering
\includegraphics[scale=0.5]{thesis4.png}
\end{frame}

\section{Proposed Methodology}
\begin{frame}{Proposed Methodology}
\begin{enumerate}
    \item Creating the system
    \begin{enumerate}
        \item Make prototype for UI
        \item Code data node
        \item Deploy \& test functionality of data node/s
        \item Code master node
        \item Deploy \& test functionality of master node/s
        \item Test interoperability of nodes
    \end{enumerate}
    \item Benchmarking
    \begin{enumerate}
        \item Create test for testing original speed
        \item Test original download speeds
        \item Create test for testing new speed
        \item Test new speeds
    \end{enumerate}
        \item Documentation
    \begin{enumerate}
        \item PDF & Github Style
        \item How to use (all users)
        \item How to set up
        \item How to maintain & add features
        \item How to test?
    \end{enumerate}
\end{enumerate}
\end{frame}

\subsection{Proposed Methodology Timeline}
\begin{frame}{Proposed Methodology: Timeline 1}
    \begin{itemize}
        \item Mid January: Start
        \item Mid April: End
    \end{itemize}
    \begin{enumerate}
        \item End January Deliverable: 
        \begin{itemize}
            \item Wireframe Mockup
            \item Pseudocode / Flowchart of Entire System (to be reviewed by the adviser)
            \item Finalized feature list
            \item Documentation Version 1 (Cid)
        \end{itemize}
        \item End of February Deliverable:
            \begin{itemize}
                \item Benchmarking specifications finalization \& Set up
                \item Start of bench marking of existing tech
                \item Start Coding
                \item Documentation Version 2 (Cid)
            \end{itemize}
    \end{enumerate}            
            \end{frame}

    \begin{frame}{Proposed Methodology: Timeline 2}
        \begin{enumerate}  
        \item Mid March Deliverable:
            \begin{itemize}
                \item Finish Coding version 1 of system
                \item Implement basic UI
                \item Finish bench marking existing 
                \item Start Alpha Testing (us and other undergrad thesis students)
                \item Documentation Version 3 (Cid)
            \end{itemize}
        \item End of March Deliverable:
            \begin{itemize}
                \item Revise version 1 to version 2 of system based on Alpha Test
                \item Revise basic UI
                \item Testing of system for feature 
                \item Start Beta Testing (public, DCS \& PGC community)
                \item Documentation Version 4 (Cid)
            \end{itemize}
        \item Mid April Deliverable
            \begin{itemize}
                \item Revise version 2 to version 3 of system based on Beta test
                \item Documentation Version 5 (Cid)
            \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}
\Huge{\centerline{The End}}
\small{\centerline{Thank you for coming}}
\end{frame}

\begin{frame}
\Huge{\centerline{Discussion points}}
\small{

\begin{itemize}
    \item What principles of database design will be focused here?
        \begin{itemize}
            \item Consistency
            \item Availability
            \item Partition Tolerance
        \end{itemize}
    \item Focus of the thesis?
    \begin{itemize}
        \item Not be focused on peer to peer? But peer to peer principles that solves the problems?
        \item Focused on making the data easier for users to get?
    \end{itemize}
    \item How will bench marking and testing be done?
    \begin{itemize}
        \item Download entire NCBI database?
        \item Download a portion?
        \item Download a specific genome (chicken / rice)?
    \end{itemize}
\end{itemize}

}
\end{frame}

\section{Bibliography}

\begin{frame}[allowframebreaks]
\printbibliography[heading=none]
\end{frame}

\end{document}

### 2020-06-18 11-25/todo.tex
Initial Overleaf Import

\documentclass{article}

\begin{document}
\title{Week Nov 19-23}

\maketitle
\section{To Do?}

%idk haha \\
%outline natin mga things to do from the minutes \\
%try to fix what we can pero no pressure \\
Intro \\ \\
- Lessen Bio Slides \\
- Explain client server \\ 
- Have more tech slides \\
- Explain Database core principles CIA? \\
- APIs? \\
- Add to P2P explain it more \\
- Find out all jargon used \\
- After that define those jargon \\
- Describe Distributed Databases more \\
- add more pictures \\
- explain Client Server and Peer 2 Peer \\
- explain network layers \\

Problem \\ \\
- Describe the problem why genomic researchers make these systems in the first place \\
- Describe problem where researchers download the whole database kasi mabagal internet nila or something \\

RRL \\ \\
- may one table comparing all the systems may check and ex \\

Objectives \\ \\
- Add some way of measuring objectives \\
- More concise \\
- Example: Implement an info delivery system that speedups download of data \\

\section{NOTES}
\subsection{Problem Statement}

\subsection{Theoretical Framework}
from seqtorr
POST
PUT
GET
DELETE

POST
1) choose file
2) turn file to torrent file
3) upload torrent file

PUT
1) choose metadata to edit
2) edit metadata
3) save metadata

GET (may expound the system more)
1) get torrent file
2) from torrent file get IP address of peers and parts they have
3) measure speed of each part from each peer
4) download the part from the fastest peer
5) combine files into the file
6) checksum??

DELETE
1) delete torrent?

\subsection{Proposed Methodology}
CREATING THE SYSTEM \\
1) code pseudocode / prototype for UI \\
2) code network shit / test out network shit \\
3) code base code \\

4) code frontend \\
5) deploy \\

TESTING \\
6) create a test for testing original speeds \\
7) test original speeds \\
8) create a test for testing new speeds \\
9) test new speeds \\
10) test UI??? \\
11) test security?? \\
12) test checksum???? \\

%Ideas kaya comment pa
%Theoretical Framework
% main server
%

%Proposed Methodology
% Mid Jan - Mid April
% Per Week?

\subsection{Saturday}
10pm - 12 mn \\
revise presentation \\
12mn - 2am \\
start theoretical framework \\
start proposed methodology \\

\subsection{Sunday}

9am-6pm wala si cid pero Carl you can start

\subsection{Sunday}
let's make a goal kaya na rest day natin sunday (and monday morning) para di tayo

\end{document}

### 2020-06-18 12-02/Minutes.tex
Initial Overleaf Import

MINUTES 2

Speedup number where did it come from?

How do you test the speedup? rn Mbps

Plan for locking mechanism? read locks write locks? Integrity of data. Permissions producer consumer problem. Consistency of data? Sabay put and get who gets it first?

By next year magpoprovide ng PCs. Physical PC will be from PGC.

UI? Di problem yung UI.
one for upload.
one for download.
one for file list, uTorrent.

FOCUS IN TRANSFER

Pros cons ng distributed ng peer to peer
How they apply to needs of problem
Maraming online di lang NCBI it is a common framework na.

Not clear. Why not have it on a smaller scale? Abstract pros and cons for diff approaches. May proposed schematic. Di diniscuss bakit yun and how it addresses needs. Di justified why ito yung architecture.

Ano yung problem? And why this specific problem is being solved by this theore framework

Di clear. We went straight to special databases and not normal databases. Why not current databases lang. Why couldnt pgc not just have an ftp server.

May plan but it isnt clear why there is a plan? What exactly are you trying ifix? What has been tried to fix before.

Does the user select between data node and master node? Can choose a mirror esque

Cant see the important aspect. Data model of database. What data is being distributed? What you capture information? What you store?

Copy it from Seqtorr -> Also Cite it

How would you store them? How you distribute? May preprocessing of data ba? Types of files? How you store?

A more specific layer. File name. Header. Sequence. Date uploaded? Species? Length of sequence?

Do we have multiple tables? Use muna data model of seqtorr? Master node that contains the metadata? If it only has metadata? Does it mirror to other data nodes?

Multiple metadata nodes? Can data nodes do what master nodes does? What if master node fails? Can you mirror the metadata to other data nodes? Master node distributes the data and kahit magfail yung main node. Suggest: remove master node keep metadata for all.

Why master node?

### 2020-06-18 12-09/Presentation.tex
Initial Overleaf Import

--- 2020-06-18 11-25/Presentation.tex
@@
 %----------------------------------------------------------------------------------------
-%	PACKAGES AND THEMES
+%	PACKAGES AND THEMEShttps://www.overleaf.com/project/5dbfadb88b2560000102ab44
 %----------------------------------------------------------------------------------------
@@
 
-\title[Short title]{Thesis Pre-final presentation} % The short title appears at the bottom of every slide, the full title is only on the title page
+\title[Short title]{caDDS: Community Assisted Distributed Database for Sequences} % The short title appears at the bottom of every slide, the full title is only on the title page
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%----------------------------------------------------------------------------------------
+%	PRESENTATION SLIDES
+%----------------------------------------------------------------------------------------
+
 \begin{document}
 
-\begin{frame}
-\titlepage % Print the title page as the first slide
-\end{frame}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OVERVIEW %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}
@@
 
-%----------------------------------------------------------------------------------------
-%	PRESENTATION SLIDES
-%----------------------------------------------------------------------------------------
+%%%%% TODO
+%% READ COMMENTS
+%% REARRANGE SOME SLIDES
+%% ADD MISSING SLIDES (KAHIT NO CONTENT) FOR NARRATIVE PURPOSES
+%% REVISE SLIDES AND ADD NOTES
 
-%------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO + RRL %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
 
-% to do sa intro
-% list all jargon
-% explain most jargon
-% iklian bio part
-% pahabain tech part esp p2p
-% explain CDN
+    \subsection{Definition} %keep daw
+    \begin{frame}{Human DNA Length}
+    % mahirap maintindihan, di madali iconnect sa complexity of sequencing
+    % not as connected
+    % put instead another slide
+    % if you unwind all the letters it reaches to the moon and back
+    % represent through a sequence of letters
+    % see both the complexity and the size of data
 
-    %%%%%%%%%%%%%%%%%%% BIO SLIDES %%%%%%%%%%%%%%%%%%%%%%%%%%
+        \centering
+        \includegraphics[scale=0.3]{dna-length.jpg}
+    \end{frame}
 
-    \subsection{Definition} %keep daw
-    \begin{frame}{Human Karyotype}
-        \centering
-\includegraphics[scale=1]{karyotype.jpg}
-    \end{frame}
     \begin{frame}{Definition of Genomics}
@@
 
-    \subsection{Introduction}
+    \subsection{Introduction + RRL}
     \begin{frame}{Introduction}
+    % kadugtong ng first slide
+    % ATCG na paulit ulit
+    % Sequence of letters codes for instructions to determine the traits of the organism
         \begin{itemize}
@@
     \begin{frame}{Genomics Data}
+    % can have billions of these...
             \textbf{Data Type}
@@
     \end{itemize}
+
     \begin{block}{Sample.FASTA}
@@
 ACAAGATGCCATTGTCCCCCGGCCTCCTGCTGCTGCTGCTCTCCGGGGCCACGGCCACCGCTGCCCTGCC
+    \end{block}
 
-    \end{block}
     \end{frame}
@@
     \begin{frame}{Sequencing}
+
         \textbf{Sequencing}
@@
     \end{frame}
-    %%%%%%%%%%%%%%%%%%% TECH SLIDES %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+   \begin{frame}{Human genome cost over time}
+
+    Graph of human genome cost over time \cite{genomics-cost} \\
+    \centering
+    \includegraphics[scale=0.3]{human-gen-cost.jpg}
+
+    \end{frame}
+
+    \begin{frame}{Sequence data cost over time}
+
+    Graph of a megabase sequence cost over time \cite{genomics-cost}
+    \begin{itemize}
+        \item Megabase means one million bases
+    \end{itemize}
+    % show how data storage vs production of data
+    % get this from hangouts
+    \centering
+    \includegraphics[scale=0.1]{seq-cost.jpeg}
+
+    \end{frame}
+
+    \begin{frame}{NGS vs Hard Disk Storage}
+            Computational sense, there is a need to store, upload, and retrieve sequence files that are growing larger. 
+            \begin{itemize}
+            \item Slope of orange line (NGS) is larger than slope of yellow line (pre-NGS)
+            \item Hard disk storage (blue) graph is greater than yellow (pre-NGS) but is now less than orange (NGS)
+            \end{itemize}
+            \centering
+                \includegraphics[scale=0.2]{seq-data.jpeg}
+
+    \end{frame}
 
@@
 
-    \subsection{Special Database}
-    \begin{frame}{Special Databases}
-        \textbf{Large Data}
+    %given these efforts, one way to solve it is to have distributed databases, dito pwede pumasok
+
+    % dito pwede pumasok yung nireview na technology
+    % in presetation di to yung specific section (RRL, Intro)
+    % presentation, mas mahalaga yung flow
+    % show subsection in a footer, no defined header
+
+    \begin{frame}{Normal: NCBI}
+        \begin{itemize}
+            \item NCBI has a large collection of genomic datasets in its website, which can be searched and downloaded.
+        \end{itemize}
+    \end{frame}
+    \begin{frame}{Normal: NCBI}
+        \begin{itemize}
+            \item Search page https://www.ncbi.nlm.nih.gov/nucleotide
+            \includegraphics[scale=0.5]{ncbi1.png}
+        \end{itemize}
+    \end{frame}
+    \begin{frame}{Normal: NCBI}
+        \begin{itemize}
+            \item List of results
+            \includegraphics[scale=0.5]{ncbi2.png}
+        \end{itemize}
+    \end{frame}
+    \begin{frame}{Normal: NCBI}
+        \begin{itemize}
+            \item Download page
+            \includegraphics[scale=0.5]{ncbi3.png}
+        \end{itemize}
+    \end{frame}
+
+    % cinomment out ko kasi meron naman explanation sa taas
+    %\subsection{P2P Database}
+    %\begin{frame}{P2P Databases}
+    %    something explaining briefly p2p databases
+    %\end{frame}
+
+        \subsection{BioTorrents}
+    \begin{frame}{P2P: BioTorrents}
+        \begin{itemize}
+            \item Scientific data continues to grow, and so does the demand for easier accessibility
+            \item Centralized servers using HTTP or FTP cannot keep up with concurrent requests
+            \item Peer-to-peer protocols does not scale well for large files
+            \item BitTorrent handles both these problems
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}{P2P: BioTorrents}
+        \begin{itemize}
+        \item System for legally sharing scientific data that works on top of the BitTorrent protocol. It is much like a public tracker.
+        \item The main website hosts .torrent and metadata files
+        \item Info about each dataset includes categories, license, filenames, etc. that helps users in searching for relevant datasets
+        \item Torrent file has list of trackers, servers that "know" which peers are serving which files.
+        \item Torrent client downloads from multiple peers simultaneously.
+        \item To upload a torrent, the client makes a .torrent file using the torrenting software, then uploads it to the main website along with metadata.
+The client must have their computer continuously turned on to keep the file available as a peer.
+\cite{biotorrents}
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}{P2P: BioTorrents}
+        \includegraphics[scale=0.33]{biotorrents-page.png}
+    \end{frame}
+
+    \begin{frame}{P2P: BioTorrents}
+        \begin{itemize}
+            \item As with all public trackers, BioTorrents requires "good souls" that will keep many torrents available as often as possible.
+            \item The datasets must be moderated to prevent abuse.
+            \item The BioTorrents website is no longer up.
+        \end{itemize}
+    \end{frame}
+
+    \subsection{PeerDB}
+    \begin{frame}{P2P: PeerDB}
+    % try adding screenshots
+        \begin{itemize}
+            \item Proposed P2P system for sharing general data
+            \item Each node is equipped with a database
+            \item Queries are passed around neighbors, with a maximum number of hops
+            \item Caches information about other peers to speed up queries
+            \item Large number of nodes \& relatively small number of lookup servers
+            \cite{peerdb}
+        \end{itemize}
+    \end{frame}
+
+    \subsection{Distributed Database}
+    \begin{frame}{Distributed Databases}
         \begin{itemize}
@@
 
-    \begin{frame}{Special Databases}
-        \textbf{Genomics Data}
-        \texttt{needs a different set of features to easily conduct processes (alignment, assembly, comparison).}
+    \subsection{SeqTorr}
+    \begin{frame}{Distributed: SeqTorr}
+    \textbf{SeqTorr} \cite{seqtorr}
     \begin{itemize}
-            \item Need a special database to store all of these data
-            \item Designed for more efficient and optimal use
-            \item Has special needs for the users of genomic data
-            \item Discuss in the RRL the different ways researchers approached this
+        \item Standard genomics workflow consists of uploading and downloading data on an international database like NCBI.
+        \item Having a distributed scalable local infrastructure to store genomic data instead of relying on NCBI would be beneficial since
+        \begin{itemize}
+            \item only certain sequences in NCBI are relevant to the institutions (e.g. Asian sequences are needed more than Caucasian ones)
+            \item researchers within the country can share data with fast speeds
+            \item institutions can share in the hosting of data
+        \end{itemize}
+    \end{itemize}
+    \end{frame}
+    \begin{frame}{Distributed: SeqTorr}
+    \begin{itemize}
+        \item Utilizes distributed database
+        \item Master node
+        \begin{itemize}
+            \item containing metadata of all the sequences
+            \item where user authentication is handled
+        \end{itemize}
+        \item Data node
+        \begin{itemize}
+            \item contain a portion of all the sequences
+            \item needs to replicate a sequence \texttt{r} times before a sequence is considered available
+            \item proximity of data nodes can increase download speed
+        \end{itemize}
+        \item Defined API Implementation
+        \begin{itemize}
+            \item POST
+            \item GET
+            \item PUT
+            \item DELETE
+        \end{itemize}
+
     \end{itemize}
@@
 
+    \begin{frame}{Distributed: SeqTorr}
+    % have a unified way of discussing each database
+    % what aspects of the database to discuss each
+    % extra features will come after
+        Problems: % have one for each database system
+        \begin{itemize}
+            \item program not currently available
+            \begin{itemize}
+                \item cannot measure any statistics on this
+            \end{itemize}
+        \end{itemize}
+    \end{frame}
+
+\begin{frame}{Table}
+% Bio-related -> biological
+% What's the difference between biological vs general vs FASTA
+% What does general mean?
+% What does biological mean? -> explain pa to
+% May 3 types of data ...
+   \begin{table}[]
+        \begin{tabular}{|l|l|l|l|l|}
+        \hline
+        System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ \hline
+        BioTorrents & P2P           & No         & Depends                    & Biological  \\ \hline
+        SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ \hline
+        PeerDB      & P2P           & No         & Depends                    & General      \\ \hline
+        NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\ \hline
+        \end{tabular}
+    \end{table} 
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PROBLEM %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Problem}
+% after discussion of dist. databases
+\begin{frame}{Problem}
+The different gaps or problems found with other implementations
+\textbf{Problems of Classical Databases}
+\begin{enumerate}
+%PROBLEMS OF CLASSICAL DATABASES (ALL FOUR)
+    \item Genome researchers download entire database of genomic data
+    \item Genome researchers only need a portion of the database, but end up downloading everything  (gigabytes size)
+    \item Client-Server approach can face bottlenecks (server speed) %is resolved by making it more distributed
+    \end{enumerate}
+% or having a hard time getting a portion of the database
+% easier to download parts vs whole, whole vs parts
+
+% not easy to navigate the database? 
+% what is the big problem? too big data to download all at once?
+% why is it a problem? speed of download? or size?
+
+\textbf{Problems of Distributed Databases}
+
+\begin{enumerate}
+%PROBLEM OF DIST. DATABASES
+    \item Only one framework on distributed database for genomes
+    \item Existing distributed database for genome code has little documentation % concept of APIs and documentation
+
+\end{enumerate}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\titlepage % Print the title page as the first slide
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OBJECTIVES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Objectives}
+% to tackle those problems
+% should be written in connection with the problems said:
+% problem x : objective x
+% para mas clear yung objectives
+
+\begin{frame}{Objectives}
+Based on the problems we found earlier, we propose the following objectives
+ \begin{itemize}
+    \item Create an information system that speeds up the download and upload of genomic data
+    \begin{itemize}
+        \item To resolve 1 \& 2 of the Problems of Classical Databases. 
+    \end{itemize}
+\end{itemize}
+
+\begin{itemize}
+    \item The system should have the main API’s
+    \begin{itemize}
+        \item POST genome data
+        \item GET genome data (implement hybrid P2P)
+        \begin{itemize}
+            \item This will aim to resolve number 3 in the Problems of Classical Databases
+        \end{itemize}
+        \item PUT genome metadata (optional: PUT data)
+        \item DELETE genome data
+    \end{itemize}
+    \item Write documentation for the system
+    \begin{itemize}
+        \item This will be used to resolve the Problems of Distributed Databases
+    \end{itemize}
+\end{itemize}
+\end{frame}
+
+\section{Scope}
+\begin{frame}{Scope}
+The current scope and coverage of the project is only
+\begin{itemize}
+    \item Security and Authentication is not part of the scope
+    \item Use only FASTA file format 
+    \begin{itemize}
+        \item DNA nucleotides only
+    \end{itemize}
+    \item Only a proof of concept, UI will be purely for functional purposes (very basic design)
+\end{itemize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% THEORETICAL FRAMEWORK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Theoretical Framework}
+% dito idiscuss yung tech
+% ito ipopropose namin 
+% lahat ng network layers, client-servers
+
         \begin{frame}{Network Layers}
+        % not part of intro na 
+        % pwede sa methods / theoretical framework na iintroduce
+        % medyo off sa flow
          \textbf{Networks Layers} how computers pass data to each other
@@
     \subsection{Application Architecture}
+    % part ng theoretical framework
+    % pwedeng walang slide of c/s
+    % pwedeng ibanggit nalang ang client-server architecture
+
+    % methods: ito framework
+    % implemens client-server...
     \begin{frame}{Application Architecture}
@@
     \end{frame}
+
         \subsubsection{Client-Server}
@@
         \end{frame}
+
         \subsubsection{Peer-to-Peer \& Hybrid}
@@
         \end{frame}
+
+        \begin{frame}{Bittorent terminology}
+            % ADD SCREENSHOT
+            Peers \\ \\ 
+             \centering
+            \includegraphics[scale=0.5]{bit-2.PNG}
+        \end{frame}
+
+        \begin{frame}{Bittorent terminology}
+            % ADD SCREENSHOT
+            Trackers \\ \\
+             \centering
+            \includegraphics[scale=0.5]{bit-3.PNG}
+        \end{frame}
+
         \begin{frame}{BitTorrent terminology}
+            % visually show each term din
+            % screenshot of torrent
             \begin{itemize} % siguro breeze through what these mean 
@@
 
-\section{Review of Related Literature}
-    \subsection{BioTorrents}
-    \begin{frame}{BioTorrents}
-        % outline
+\begin{frame}{Theoretical Framework: System Description}
     \begin{itemize}
-        \item System for legally sharing scientific data that works on top of the BitTorrent protocol. It is much like a public tracker.
-        \item The main website hosts .torrent and metadata files
-        \item Info about each dataset includes categories, license, filenames, etc. that helps users in searching for relevant datasets
-        \item Torrent file has list of trackers, servers that "know" which peers are serving which files.
-        \item Torrent client downloads from multiple peers simultaneously.
-        \item To upload a torrent, the client makes a .torrent file using the torrenting software, then uploads it to the main website along with metadata.
-The client must have their computer continuously turned on to keep the file available as a peer.
-\cite{biotorrents}
+        \item A \textbf{master node} stores metadata of all sequences, may or may not contain sequence files
+        \item \textbf{Data nodes} stores sequence files \& metadata of all sequence files in its filesystem
+        \item Both master node \& data nodes have user interface that allow users to upload, search, and download sequence files
     \end{itemize}
@@
 
-    \begin{frame}{BioTorrents}
-        \begin{itemize}
-            \item As with all public trackers, BioTorrents requires "good souls" that will keep many torrents available as often as possible.
-            \item The datasets must be moderated to prevent abuse.
-            \item The BioTorrents website is no longer up.
-        \end{itemize}
-    \end{frame}
-
-    \subsection{PeerDB}
-    \begin{frame}{PeerDB}
-        \begin{itemize}
-            \item Proposed P2P system for sharing general data
-            \item Each node is equipped with a database
-            \item Queries are passed around neighbors, with a maximum number of hops
-            \item Caches information about other peers to speed up queries
-            \item Large number of nodes & relatively small number of lookup servers
-            \cite{peerdb}
-        \end{itemize}
-    \end{frame}
-
-    \subsection{SeqTorr}
-    \begin{frame}{SeqTorr}
-    \textbf{SeqTorr} \cite{seqtorr}
-    \begin{itemize}
-        \item Utilizes distributed database
-        \item Master node
-        \begin{itemize}
-            \item containing metadata of all the sequences
-            \item where user authentication is handled
-        \end{itemize}
-        \item Data node
-        \begin{itemize}
-            \item contain a portion of all the sequences
-            \item needs to replicate a sequence \texttt{r} times before a sequence is considered available
-            \item proximity of data nodes can increase download speed but isn't true P2P
-        \end{itemize}
-        \item Defined API Implementation
-        \begin{itemize}
-            \item POST
-            \item GET
-            \item PUT
-            \item DELETE
-        \end{itemize}
-
-    \end{itemize}
-    \end{frame}
-
-    \begin{frame}{SeqTorr}
-        Problems:
-        \begin{itemize}
-            \item program not currently available
-            \begin{itemize}
-                \item cannot measure any statistics on this
-            \end{itemize}
-            \item not Peer2Peer
-        \end{itemize}
-    \end{frame}
-
-    \begin{frame}{Table}
-   \begin{table}[]
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ \hline
-BioTorrents & P2P           & No         & Depends                    & Bio-related  \\ \hline
-SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ \hline
-PeerDB      & P2P           & No         & Depends                    & General      \\ \hline
-NCBI        & CS            & Yes        & Slow (US server)           & Bio-related  \\ \hline
-\end{tabular}
-\end{table} 
-    \end{frame}
-
-% NCBI Biosystems
-% pros
-% cons
-
-% BitTorrious
-% pros
-% \cite{bittorious}
-% cons
-
-\section{Problem}
-\begin{frame}{Problem}
-The different gaps or problems found with other implementations
-
-\begin{itemize}
-    \item Genome researchers download entire database of genomic data
-    \item Genome researchers only need a portion of the database, but end up downloading all
-    \item Client-Server approach can face bottlenecks (server speed)
-    \item Code has no documentation, thus harder to maintain
-%% add more
-
-\end{itemize}
-\end{frame}
-
-\section{Objectives}
-\begin{frame}{Objectives}
-Based on the problems we found earlier, we propose the following objectives
-    \begin{itemize}
-        \item Create information system that speeds up the download and upload of genomic data by at least 3x compared to existing systems
-        \item The system should have the main API's
-        \begin{itemize}
-            \item POST genome data
-            \item GET genome data (implement hybrid P2P)
-            \item PUT genome metadata (optional: PUT data)
-            \item DELETE genome data
-        \end{itemize}
-        \item Write documentation for the system
-    \end{itemize}
-\end{frame}
-
-\section{Scope}
-\begin{frame}{Scope}
-The current scope and coverage of the project is only
-\begin{itemize}
-    \item Security and Authentication is not part of the scope
-    \item Use only FASTA file format 
-    \begin{itemize}
-        \item DNA nucleotides only
-    \end{itemize}
-    \item UI will be purely for functional purposes (very basic design)
-\end{itemize}
-\end{frame}
-
-\section{Theoretical Framework}
-
-\begin{frame}{Theoretical Framework: System Description}
-    \begin{itemize}
-        \item A \textbf{master node} stores metadata of all sequences, may or may not contain sequence files
-        \item \textbf{Data nodes} stores sequence files & metadata of all sequence files in its filesystem
-        \item Both master node & data nodes have user interface that allow users to upload, search, and download sequence files
-    \end{itemize}
-\end{frame}
 % DRAW.IO FILE: https://drive.google.com/file/d/1Nbp0Kmgcr2BZ13LBT3SGLMBoA6JA_Eth/view
@@
 
+\subsection{Data Model}
+\begin{frame}{Data Model}
+%% GET THIS FROM SEQTORR
+
+% So far the data stored will be 1 table
+% Sequence, Author, Institution, Species, Chromosome, Chromosome Position, Date -> CHECK SEQTORR
+% Part of Theo Framework
+This data model is lifted from SeqTorr \cite{seqtorr}
+
+\small
+\begin{center}
+
+\begin{tabular}{c|c}
+\hline
+    key & description \\
+\hline
+\hline
+    seq\textunderscore id & unique ID assigned to all sequences \\
+    organism & organism where the sequence was extracted from \\
+    quality & the rating of the data based on the number of downloads or another criteria \\
+    uploader & user that uploaded the data \\
+    institute & institute where the uploader or sequence was taken from \\
+    upload\textunderscore date & date the data was uploaded to the database \\
+    last\textunderscore modified & the date the metadata or data was edited \\
+    data\textunderscore nodes & list of data nodes that this data contains
+\end{tabular}
+\end{center}
+
+\end{frame}
+
+\subsection{Data Permissions}
+\begin{frame}{Data Permissions}
+% Uploading can be done by anyone (easily)
+% But only master (we can call it moderator) can screen and verify
+% Anyone can download -> has to be a hash sum to verify integrity of contents
+% Users - Moderator, Users / Peers
+% Part of Theo Framework
+% EXPLAIN ^^ NALANG
+
+% Data Integrity - Makes sure data downloaded is always consistent -> use hashing algorithms, save the hash in the master file, checksum / hash
+
+% merong private 
+% -> kahit sa UI lang siya magauthenticate
+% -> sa table lang nakasabi sino may access (just an added column)
+% merong public
+\begin{center}
+\begin{tabular}{c|c}
+\hline
+    user & description \\
+\hline
+\hline
+   Moderator & Has ability to upload, download, store data, verify \\
+    Peer & Has ability to upload, download, store data \\
+    Guest & Has ability to upload, download \\
+
+\end{tabular}
+\end{center}
+
+\end{frame}
+
+% where does data pass through (schematic)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PROPOSED METHODOLOGY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Proposed Methodology}
@@
 \begin{enumerate}
+
+% have 3 subsections, each of those tasks, turn into sentences
+% try to identify tools per task (or possible tools)
     \item Creating the system
     \begin{enumerate}
-        \item Make prototype for UI
         \item Code data node
@@
     \begin{enumerate}
-        \item PDF & Github Style
+        \item PDF \& Github Style
         \item How to use (all users)
         \item How to set up
-        \item How to maintain & add features
+        \item How to maintain \& add features
         \item How to test?
@@
 \end{frame}
+
+%Proposed tech background -> paper
+% Visualize Gantt Chart
+% Cluster together the parts
+% Gantt chart nalang
+% As early as possible
+% Try transfer between nodes, com sci -> PGC
+
+%GAMEPLAN
+%pagnaset up kaagad
+%check speedup, kahit sobrang pangit basta may speedup
+%pag may speedup graduate na
+
+%GAMEPLAN
+% 1) Working System
+% 2) May Speedup
+% 3) Tama Data na kinocopy 100% accuracy
+% 4) Tama Distribution, ay feature to find fastest route, way to find fastest download piece
+% 5) Usability & Documentation
+% 6) Authenticaion
+% 7) UI
+
+% FIRST WEEK
+% choose dataset from ncbi
+% time the dataset download
+% set this as initial time to beat
+% compare between normal file transfer
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TIMELINE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
         \begin{itemize}
+        % start having API calls na
             \item Wireframe Mockup
@@
 
+    \begin{frame}{Proposed Methodology: Gantt Chart}
+        \centering
+        \includegraphics[scale=1]{jan.png}
+    \end{frame}
+
+    \begin{frame}{Proposed Methodology: Gantt Chart}
+        \centering
+        \includegraphics[scale=1]{feb.png}
+    \end{frame}
+
+    \begin{frame}{Proposed Methodology: Gantt Chart}
+        \centering
+        \includegraphics[scale=1]{mar.png}
+    \end{frame}
+
+    \begin{frame}{Proposed Methodology: Gantt Chart}
+        \centering
+        \includegraphics[scale=1]{april.png}
+    \end{frame}
+
 \begin{frame}
@@
 \small{\centerline{Thank you for coming}}
-\end{frame}
-
-\begin{frame}
-\Huge{\centerline{Discussion points}}
-\small{
-
-\begin{itemize}
-    \item What principles of database design will be focused here?
-        \begin{itemize}
-            \item Consistency
-            \item Availability
-            \item Partition Tolerance
-        \end{itemize}
-    \item Focus of the thesis?
-    \begin{itemize}
-        \item Not be focused on peer to peer? But peer to peer principles that solves the problems?
-        \item Focused on making the data easier for users to get?
-    \end{itemize}
-    \item How will bench marking and testing be done?
-    \begin{itemize}
-        \item Download entire NCBI database?
-        \item Download a portion?
-        \item Download a specific genome (chicken / rice)?
-    \end{itemize}
-\end{itemize}
-
-}
 \end{frame}

### 2020-06-18 12-09/minutes.tex
Initial Overleaf Import

scalability
user interface
upload speed

specific objectives:
benchmarking objectives address those objectives

??accuracy, ??error rate

download conditions?

can test if a data node goes down? robustness

raw download speeds? is there a standard way of downloading that incorporates the network used.

properly define metrics na

ano yung expected type of environment we would be testing our system? 
expected: test it on natural environment
use network on ny given day?

controlled network infrastructure? or to show unperturbed download speed

NCBI, there is another downloading way of whole database download

and use that for benchmarking

how much resources do you see.

resources:
- set up a workstation, gawing master server, won't be disturbing for others
- specify more the projected resources we need
: clear is one master na always
: one data node sa baba
: one sa dcs
: ask sir cabarle, sir mario, a private network
: public, tapos firewall IP table block,
: Cpanel ...

revised proposal:
- include resources needed
- put it in methodology

comm. assisted. dist. database for sequences

caDDS

- PGC data node, is a workstation sa baba, within the network.
- if testing outside, DCS.

WED Dec 11

FRIDAY Dec 13, 12 nn. Official Rev. Proposal na Paper

FINAL FINAL DEADLINE BY 18 12 nn dapat satisfactory na.

Kung may revision, incomplete pa yun.

### 2020-06-18 DEPRECATED PAPERS/(DEPRECATED) Paper.tex
Initial Overleaf Import

% https://www.overleaf.com/6622695572gyrgqvhzstkh

\documentclass{article}

%\documentclass[conference]{IEEEtran}
%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{xcolor}    
\usepackage{graphicx} % Allows including images
\usepackage[backend=bibtex,style=verbose-trad2]{biblatex}
\bibliography{bibfile}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\color{blue}

\title{caDDS: Community Assisted Distributed Database for Sequences}

\author{\IEEEauthorblockN{1\textsuperscript{st} Carl Araya}
\IEEEauthorblockA{\textit{Algorithms and Complexity Laboratory } \\
\textit{Philippine Genome Center}\\
\textit{University of the Philippines - Diliman}\\
Quezon City, Philippines \\
jjazcarraga@up.edu.ph}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Cid Azcarraga}
\IEEEauthorblockA{\textit{Algorithms and Complexity Laboratory } \\
\textit{Philippine Genome Center}\\
\textit{University of the Philippines - Diliman}\\
Quezon City, Philippines \\
jjazcarraga@up.edu.ph}
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
%(done)CID: Start the idea that inside each body is the following:
Inside every human body are the instructions to create and maintain it. These instructions are found inside the genome. The entire human genome, when laid out is around 2 meters in length. \autocite{ency_sci_tech}
%(done) KING: cite this
But once all of the DNA of all the cells in the human body are combined. The length can span the earth to moon 6000 times. This shows the length and amount of DNA inside the human body. This describes as well the vast amounts of information in the body.

%CID: Add a paragraph elaborating further on the amount of information here.

%(partial done) CID: This intro lacks references and needs better visual description

The word \textit{genome} was used briefly. But first, what does genomes or genomics actually mean?

\subsection{Definition}
National Institute of Health \autocite{genomics-definition} defines genomics as \\ \\
\textit{Genomics is the study of all of a person's genes (the genome), including interactions of those genes with each other and with the person's environment.} \\ \\

%(partially done) CID: Mej abrupt pagpasok kaagad sa genomics na definition. Walang clean transition

In the definition, there was also a word called genes. But what are genes? Genes comprises of DNA. DNA is an acronym that stands for deoxyribonucleic acid. DNA is usually characterized by the letters A,T,G,C. Where A is for Adenine, T is for Thymine, G is for Guanine, C is for Cytosine. These are the instructions passed on from one generation to another, this is also what is used to form the organism. \autocite[p.~2]{hartl2018}

%CID: Explain what the acronym DNA means. Then describe why that acronym becomes ATGC

These nucleotides
%(done)KING: Bases are the nucleotide
%CID: Explain this part more, it doesnt really explain the importance
form the base pairs (or \textit{bases}). These bases act as a blueprint to describe and instruct the body on how to build it and function.\autocite{alberts_mole}
%CID: Expound this sentence more, perhaps explain the build and functioning process

There are these components, and these describe complex organisms. Which means there is a long set of instructions to detail these complex organisms. For humans, the full human genome contain around $3.2x10^9$ base pairs \autocite{introgenomics}. This is a lot of information inside the human organism. But before analyzing that information, these have to be taken from the cells and read.

%CID: Separate the paragraphs to explain what genes are, to the paragraph explaining the complexity of the human genome (also elaborate on this)

\subsection{Sequencing}

%(partially done) CID: Have a clear transition to sequencing, you ended at complexity.
To see the information of the DNA of a cell. There are machines called \textbf{Sequencers} that look (or more appropriately, \textit{read}) the DNA of a specimen and outputs A,G,C,T. Sequencers look at the exact order of a single strand of DNA. The process of sequencing DNA isn't a simple process of a machine reading one strand of DNA. In most cases, one strand is too little for one machine to read. So before reading, the DNA is replicated multiple times in the laboratory. DNA is also fragmented before processing. Fragmented meaning it will be cut up into pieces.

%CID: Describe what fragmented means

These are fragmented since the machine cannot handle reading the entire length of the DNA, it has to read it in parts. Sequencers then reads (or sequences) them.  These reads of fragments from the replicated DNA will then be put back together computationally to form the \textit{correct or reference} DNA.

%CID: The paragraph on sequencing is good. Maybe cite more.

Back in 2003 The first human genome \autocite{introgenomics} had been first completely read. It took the project ten years of work, a dozen institutions, and 3,000,000,000 dollars to finish. Currently, in 2019, the technology today in Japan is much faster than the technology back in 2003. The technology in Japan can read 10,000 human genomes per day. \autocite[p.~19]{introgenomics} 
%(done) KING: Cite

% (partially done) CID: Get the transition of the NGS data graph from the slides

\begin{figure}[h]
\caption{Cost per Genome Data Over Time}
\centering
\includegraphics[width=0.65\textwidth]{images/human-gen-cost.jpg} 
\label{fig:human_gen_cost_fig}
\end{figure}

Looking at the Figure~\ref{fig:human_gen_cost_fig} on Page~\pageref{fig:human_gen_cost_fig} \autocite{genomics-cost}, we note the expensive cost at the initial stages of sequencing the human genome. But as time progresses, the cost substantially went down to what it is now. The big fall of the prices around 2004 - 2005 corresponds to the introduction of the Next Generation Sequencers.

\subsubsection{Next Generation Sequencing}

%(done) CID: Describe the speedup in terms of what and how the speedup was computed

The introduction of Next Generation Sequencers can be seen as a significant event in the history of genomics. This is a new technology which has allowed the machines to read faster and cheaper. These utilize a variety of technologies, some make use of light, some use ion semiconductors, others use a special DNA marker with a fluorescent probe \autocite[~p.67]{paulselzer2018}. This tech has increased the read speed and speed in which data is produced. But note as the data increases, storage demand therefore also increases. Since each read requires storage. This poses a problem to projects involving sequences, the data volumes is usually very large\autocite{bon_compression}. There exists a concept of \textit{coverage} as well. It means the number of reads it needs in order to produce a reference sequence (the correct, standard sequence). Meaning the sequence isn't read only once. This contributes to the large data size of sequences. Currently, for a coverage of around 60, the file size for a compressed human genome can easily reach 200 GB. And a project of 10-20 genomes would need approximately 4 TB of storage. Transferring these data across research groups would indeed be a challenge. \autocite[~p.68]{paulselzer2018}
%(done) KING: cite
% more space is used.

%(done) CID: Visually explain how more space is needed (cite numbers). May separate the sentences describing NGS and space utilization

\subsection{Database}

\begin{figure}[h]
\caption{Megabase Cost Over Time}
\centering
\includegraphics[width=0.65\textwidth]{images/seq-cost.jpeg} 
\label{fig:megabase_cost_fig}
\end{figure}

Similar to Figure~\ref{fig:human_gen_cost_fig} on Page~\pageref{fig:human_gen_cost_fig}, we also note the same pattern for the cost of sequencing (not just humans) in Figure~\ref{fig:megabase_cost_fig}\autocite{genomics-cost}. The cost of sequencing dropped 1000 fold from 2008 to 2012 alone \autocite{bon_compression}. This lowering of cost, makes the idea of sequencing more available. And thus produces a lot more data. But these terabytes of data need to be stored somewhere.

\begin{center}
\textit{Moore's law states that the number of transistors on a chip doubles every two years.}
\end{center}

%CID: An idea: make a quote stating Moore's Law, cite, then explain how this relates to thesis
%KING: Cite and check if true

This means, that as time passes by, technology for processing and storage will improve. And the overall capacity for what technology can do will increase as well. But the speed in which the number of sequence data increases seems to exceed Moore's Law. And this will pose a problem since the technology doesn't seem to be catching up. The immediate problem is the storage of all the data being produced. A certain technology needs to be made for the purpose of holding this data. This technology is referred to as Databases.

%CARL: change "flow" of ideas, maybe reorder/rewrite subsection names
% currently, the subsections are:
% - Database
%   - Principles
%   - Database Types
% - App Archi
%   - CS
%   - P2P
%   - Hybrid

\subsubsection{Definition}
A database (DBMS, Database Management System) is defined by containing information about a particular enterprise. Functionalities include: \autocite{Silberschatz2010}
\begin{itemize}
    \item Collection of interrelated data
    \item Set of programs to access the data
    \item An environment that is both convenient and efficient to use
\end{itemize}

These form a system that is made for accessing and updating data.

%CID: Summarize a bit on why this was discussed and why this is needed in the thesis

\subsubsection{Principles} 
There are also 3 principles to remember when creating a database. The idea is that only 2 of the three can be focused on at any given time. And these need to be taken into consideration when designing the system. \autocite{Silberschatz2010}

\begin{itemize}
    \item Consistency - Update one part, will all other (replicated) parts be updated quickly as well; Data is accurate and copied faithfully. An example would be that genes for specimen A, when copied to another data node should contain the exact copy of the data, with no minor changes.
    \item Accessibility - How easy it is to access data, API protocols. An example would be the number of commands  a user needs to do to access every data.
    \item Partition Tolerance - If one portion breaks, the other portions are still usable. An example would be a power failure for one data node, how would this affect the other nodes?
\end{itemize}
%CID: Use explanations from the slides
%CID: Explain CRUD

\subsubsection{Database Types} 
Two types of databases will be discussed in this paper. Some of these databases will use technical terms, these terms will be discussed later in the Theoretical Framework. Namely: \autocite{centralizedvsdistributed}

\begin{itemize}
    \item Centralized - In a centralized database, all data are managed by a single DBMS in a single node to be distributed to users. As a result, transactions are easily manageable because everything can be processed in the single node, but the cost of communication is high and reliability is low, as an error in the database will mean disaster to the network as a whole.
    \item Distributed - In a distributed database, data are stored in multiple servers in different geographical locations. Each node may contain a part or the entirety of the data. The difficulty of managing transactions will go up, but the result is higher availability, faster response, and modular growth, among other advantages. 
\end{itemize}

%\begin{itemize}
%    \item Client-Server Database - Utilizes a client-server architecture. One server to sends and receives the data to all the clients. 
%    \item P2P Database - Utilizes a peer architecture. Each peer has the ability to send and receive the data to all the other peers. There is no hierarchy.

    %KING: One server that sends
    %KING: One server for sending
    %KING: One server sends
    %CID: Basically reexplain client and server architecture, wha does client mean, what does server mean

%    \item Hybrid (Or Distributed) Database - Utilizes master node and data node. Unlike the normal database, this maintains a different structure, where not only one server has all the data. And a concept of replication of data is happening.
%\end{itemize}

%CID: I dont see any problems here, perhaps cite nalang a bit more

%CID: Explain muna P2P CS H in intro.

\subsection{Application Architecture}
The application architecture is how a networked app is structured
across the various end systems (i.e. computers, servers, etc.) \autocite{kurose}
Falls into either client-server, peer-to-peer, or hybrid architecture
%CID: Explain a bit how application architecture affects the app? Like does this change the features or speed or traffic management?

\subsubsection{Client-Server}
A client-server architecture relies on a Web server that is always
available and has a fixed IP address. When a client sends a request for data, the server responds by sending the requested data. The client-server architecture may not be able to keep up with the increasing number of connections.

\subsubsection{P2P}
A peer-to-peer (P2P) architecture has minimal to no reliance on
always-on, dedicated servers. Peers are desktops/laptops that communicate to each other without passing through a dedicated server. These include internet telephony, BitTorrent, and some other file sharing protocols. An advantage of P2P networks is their scalability. Each user can benefit the whole network by sharing his/her own resources. However, P2P has its own problems, namely security and incentivizing users to share their resources.

\subsubsection{Hybrid}
A hybrid architecture combines elements of P2P and client-server architectures, attempting to utilize advantages of both.

%CID: Explain some problems din kaya?
%CARL: wala sa kurose eh what if gusto ko magdagdag ng sarili ko haha

%KING: Fix periods in a sentence

% this is only a particular implementation of hybrid kasi so cinomment out ko
 %A master node stores metadata of all sequences, may or may not contain sequence files. Data nodes stores sequence files and metadata of all sequence files in its filesystem. Both master node & data nodes have user interface that allow users to upload, search, and download sequence files.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% RRL %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Review of Related Literature}

We review the technologies on data storage done before. 
%CID: Make a transition to go from the diff database types to the explanation of the diff databases

\subsection{SeqTorr: A distributed scalable database for genomic information}
The standard genomics workflow consists of uploading and downloading data on an international database like NCBI. Having a distributed scalable local infrastructure to store genomic data instead of relying on NCBI would be beneficial since: only certain sequences in NCBI are relevant to the institutions (e.g. Asian sequences are needed more than Caucasian ones), researchers within the country can share data with fast speeds, and  institutions can share in the hosting of data.

SeqTorr consists of a master node and multiple data nodes. The master node contains metadata of all the sequences and is where the user authentication is handled, while the data node is where the actual sequences are stored. All user uploads happen in the master node. When the master node receives a sequence from a user, it distributes the sequence to any available data nodes. 
\autocite{seqtorr}

Currently, SeqTorr is not available, and the code base does not have any documentation which makes future work difficult.

\subsection{BioTorrents}
Scientific data continues to grow, and so does the demand for easier accessibility. Centralized servers using HTTP or FTP cannot keep up with concurrent requests, and peer-to-peer protocols does not scale well for large files. BitTorrent handles both these problems.

BioTorrents is a system for legally sharing scientific data that works on top of the BitTorrent protocol. It is much like a public tracker. The main website hosts .torrent and metadata files. Info about each dataset includes categories, license, filenames, etc. that helps users in searching for relevant datasets. The torrent file has a list of trackers, servers that "know" which peers are serving which files, so it can download from those peers simultaneously. To upload a torrent, the client makes a .torrent file using the torrenting software, then uploads it to the main website along with metadata. The client must have their computer continuously turned on to keep the file available as a peer. The main issue with BioTorrents is that it requires lots of users that will keep many torrents available as often as possible.
\autocite{biotorrents}

%CID: Have a paragraph explaining the problems of BioTorrents
However, BioTorrents has been down for years. As with all public trackers, it requires users to always keep the data publicly available, and the datasets must be moderated to prevent abuse.

\subsection{PeerDB}
% summarize intro
PeerDB is a proposed P2P system for sharing general data. Each node is equipped with a MySQL database that it keeps available for other nodes to download from. Each node also keeps track of its neighbors, and any useful information about them. Queries are passed around neighbors, with a maximum number of hops. A node may cache information about other peers to speed up future queries.
\autocite{peerdb}

%CID: Have a paragraph explaining the problems of PeerDB
% The issues faced by PeerDB is ...

%CID: Explain NCBI

\subsection{NCBI}

%CID: Discuss further how NCBI works
%This is how NCBI works... (Insert Sir El King's demonstration of how to bulk download).

%CID: Explain a few problems faced by NCBI
%The problems seen based on bulk downloading are ...

%describe more the problems per database
%normal vs p2p vs dist databases

%CID: Explain muna P2P CS H in intro.
%CID: Have a sentence explaining this table in word form and adding all the other problems discussed above

\begin{table}[h]
    \caption{Comparisons of Different Databases}
    \begin{tabular}{|l|l|l|l|l|}
    \hline
    System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ \hline
    BioTorrents & P2P           & No         & Depends                    & Biological  \\ \hline
    SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ \hline
    PeerDB      & P2P           & No         & Depends                    & General      \\ \hline
    NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\ \hline
    \end{tabular}
    \label{table:database_comparison_table}
\end{table}

Table~\ref{table:database_comparison_table} on page ~\pageref{table:database_comparison_table} describes the different technologies done. This will also be used to decide what are the main problems of the different databases.

% (done) KING: 1) Needs a table header
%(done) KING: 2) Fix latex positioning, show the table after the PeerDB Section
%(done) KING: 3) When you show a table / figure, make sure you reference that somewhere in the discussion

\section{Problem}
%CID: Have some discussion then somehow the discussion concludes to the idea that the following problems occur, these problems mej came out of nowhere kasi
%CID: Expound more on each category

The different gaps or problems found with other implementations
\subsection{Problems of Classical Databases}
\begin{enumerate}
\item Genome researchers download entire database of genomic data
\item Genome researchers only need a portion of the database, but end up downloading everything
\item Client-Server approach can face bottlenecks (server speed)
\end{enumerate}

\subsection{Problems of Distributed Databases}
\begin{enumerate}
    \item No existing usable framework on distributed database for genomes
    \item Existing distributed database for genome code has no documentation to start with % concept of APIs and documentation
\end{enumerate}
%CID: Evaluate if these problems are accurate. Maybe cite earlier paragraphs?? I dont know if that's possible

\section{Objectives}

%CID: Have a discussion, like what I said sa problem to make the reader want to say the objectives we will state. Like we want to lead the reader to what we will say
%CID: Maybe restate the problems super briefly, and then discuss what tech is available
%CID: Explain more on the metrics to be used

Based on the problems we found earlier, we propos to:
%CID: Hmmm parang okay naman, the wording nalang

\begin{itemize}
    \item Create an information system that speeds up the download and upload of genomic data
    \begin{itemize}
        \item To resolve 1 \& 2 of the Problems of Classical Databases. 
    \end{itemize}
\end{itemize}

\begin{itemize}
    \item The system should have the main API’s
    \begin{itemize}
        \item POST genome data
        \item GET genome data (implement hybrid P2P)
        \begin{itemize}
            \item This will aim to resolve number 3 in the Problems of Classical Databases
        \end{itemize}
        \item PUT genome metadata (optional: PUT data)
        \item DELETE genome data
    \end{itemize}
    \item Write documentation for the system
    \begin{itemize}
        \item This will be used to resolve the Problems of Distributed Databases
    \end{itemize}
\end{itemize}

%KING: Use Genomic Data instead of genome data

\section{Scope}
%(done) CID: pahabain pa konti

The project will focus mainly on the functionalities of the system, and not focus much yet on the different possible data types and data in the system. Therefore the current scope and coverage of the project:

%(done) KING: This sentence does not make sense (the coverage is only: security and authenticaion is not part of the scope), make it sound that it will work when it isnt in bullet form
%(done) CID: Weird sentence, just rephrase

\begin{itemize}
    \item will be the use of only the FASTA file format
    \item will be the use of only the DNA nucleotides
    \item will only a proof of concept, UI will be purely for functional purposes
(very basic design)
\end{itemize}

The current scope and coverage of the project will NOT contain:

\begin{itemize}
    \item Security and Authentication
\end{itemize}

Inside the university there are many problems associated with having security when using network resources, that is why authentication will not be part of the scope.

\section{Theoretical Framework}

\subsection{Networks}

%KING: everyone is assumed to know networks
%(done) KING: Please cite
%CID: Can reduce the discussion here. Then emphasize the importants parts
Networks Layers are the abstraction of how computers pass data to each other. This is needed to maximize the speed in which data is transferred between computers. A network, specifically the Internet, is comprised of the following levels of abstraction. \autocite{kurose}
\begin{enumerate}
    \item Level 1: Medium used to send signals (physical, electrical signals of 0
and 1), through wire or by air
    \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together
different . Goal of sending a group of signals.
    \item Level 3: Introduction of networks: IP Address to uniquely identify
entities.
    \item Level 4: Management of error free data transmission: Sending a group
of messages via TCP, UDP, FTP
    \item Level 5: Abstraction of Level 4: Function calls programmers will use
to do level 4 processes.
\end{enumerate}

Level 1 through 3 focus on building a system for sending messages from one computer entity to another. And this isn't important for the thesis, but important to understanding the scope of what the thesis will do.

Level 4 and 5 here is where this thesis will dwell upon. Since this is more specific to what the thesis will be needing, such as file transfer, checking if the files are error free, and more.
%CID: expound this sentence more

% (partially done) CID: Explain POST PUT GET DELETE in relation to CRUD, idk if needed to

In networks, when there is data being accessed over the network, it is not called C.R.U.D., instead it is called POST, GET, PUT, DELETE. POST is analogous to CREATE meaning the user submit data to be stored (in the server). GET is similar to READ but is in the context of reading a file over a network (from a server). PUT is analogous to UPDATE as DELETE remains as DELETE. These are codes used to denote network commands that do the same thing as the CRUD functions discussed earlier. This is important since there is an added layer of difficulty when it comes to transmitting data over a network.

%(partially done) CID: Have a paragraph summarizing what the system does (Like expanded version ng objectives pero mas may technical shizz)

\subsection{Summary of the System}
The system is essentially a database. The database stores sequence data. It has the normal CRUD functions meaning Create, Read, Update, Delete. It can let anyone download sequence data. But only registered users can upload sequence data. When a user uploads data, the moderator will verify the data's authenticity before allowing it to be uploaded. Then for the downloading the data. The user will experience a normal download, but under the hood the system will be doing a lot of work. It will check multiple "mirrors" of the download, by looking at the different available data nodes, then let the user choose the fastest or best option. File transfer will occur, then the system will be the one checking if the file were downloaded correctly. There is a simple organization of available files as well.

%(partially done) CID: the paragraph above is still informal

%(done) CID: Also explain why title is that

The system is described as \textit{community assisted}. Since ideally, institutions can contribute to sharing the data by hosting servers that will become the data nodes. This sharing allows a greater bandwidth, and hopefully speed in which this data can be transferred. The speed will be measured in order to measure if this project achieved its objective. The database is a distributed database, which will aim to utilize the best features from Peer2Peer and Client/Servers. This will also aim to store sequences. Thus the "Distributed Database for Sequences" portion of the title.

\begin{figure}[h]
\caption{App network diagram}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis1.png} 
\end{figure}

%CID: Explain briefly each figure

\begin{figure}[h]
\caption{POST: Uploading a sequence to the master node}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis3.png} 
\end{figure}

\begin{figure}[h]
\caption{POST: Uploading a sequence to the data node}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis2.png} 
\end{figure}

\begin{figure}[h]
\caption{GET: Downloading a sequence}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis4.png} 
\end{figure}

%CID: Explain the data model and data permissions

%CID: Data Model, make sure to cite this in a paragraph

\begin{table}[h]
\caption{Data Model to be Used in the System}

\begin{tabular}{c|c}
\hline
    key & description \\
\hline
\hline
    seq\textunderscore id & unique ID assigned to all sequences \\
    organism & organism where the sequence was extracted from \\
    quality & the rating of the data based on the number of downloads or another criteria \\
    uploader & user that uploaded the data \\
    institute & institute where the uploader or sequence was taken from \\
    upload\textunderscore date & date the data was uploaded to the database \\
    last\textunderscore modified & the date the metadata or data was edited \\
    data\textunderscore nodes & list of data nodes that this data is contained to

\end{tabular}
\label{table:data_model_table}

\end{table}

Table~\ref{table:data_model_table} on Page~\pageref{table:data_model_table} aims to show the metadata that will be initially stored in the database.

%CID: Data Permissions

\begin{table}[h]
\caption{Data Permissions and Users in the System}
\begin{tabular}{c|c}
\hline
    user & description \\
\hline
\hline
   Moderator & Has ability to upload, download, store data, verify \\
    Peer & Has ability to upload, download, store data \\
    Guest & Has ability to upload, download \\

\end{tabular}
\label{table:data_perm_table}
\end{table}

Table~\ref{table:data_perm_table} on Page~\pageref{table:data_perm_table} shows the different set of users and the permissions allowed to them in the system. The moderator is a special user who has the ability to not allow some data to be uploaded. This is to make sure there is a means to verifying the data being uploaded. All users can upload, download, and store (or host) data. All other users, which are called guests has an ability to download and upload data.

\section{Proposed Methodology}

% CID: explain the 3 main phases (system creation, testing, documentation)
% CID: explain the importance of each phase
% CID: explain

\subsection{Creating the system}
The system will be needed to be done before any of the following steps can occur. 
\begin{enumerate}
    \item Set-up database architecture
    \item Code the POST API
    \item Code the DELETE API
    \item Code the PUT API
    \item Code the GET API
    \begin{enumerate}
        \item As a classic database
        \item Revise to have distributed database functionality (may be moved to item 1) 
        \begin{enumerate}
            \item Code data node
            \item Deploy \& test functionality of data node/s
            \item Code master node
            \item Deploy \& test functionality of master node/s
            \item Test interoperability of nodes
        \end{enumerate}
    \end{enumerate}
\end{enumerate}

\subsection{Testing}
Part of the goal of the project is to implement a better data transfer mechanism. And one way to verify this is to see the speed in which the system compares to other systems.

\begin{enumerate}
    \item Create test for testing original speed
    \item Test original download speeds
    \item Create test for testing new speed
    \item Test new speeds
\end{enumerate}

\subsection{Documentation}
Another part of this project's goal as well is to make this available and usable for the community. And documentation will help make the system easy to use, manage, and develop.
\begin{enumerate}
    \item Compiled to a PDF that is easy to read
    \item How to use (all users)
    \item How to set up
    \item How to maintain and add features
\end{enumerate}

\subsection{Resources needed}

Servers must have networking capability and at the very least have their HTTP port configured to be open to each other, or optionally, the HTTP port can be open to the Internet.
\begin{enumerate}
    \item 1-2 servers in PGC to serve as master node (1 minimum; the other server will serve as data node)
    \item 1-2 servers in DCS to serve as data nodes (1 minimum)
    \item 2 workstations where both of us can program the system
    \end{enumerate}
\end{enumerate}

\printbibliography

\end{document}

### 2020-06-18 DEPRECATED PAPERS/(DEPRECATED) Paper2020OldFormat.tex
Initial Overleaf Import

--- 2020-06-18 DEPRECATED PAPERS/(DEPRECATED) Paper.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+% START OF PROGRESS FOR CID MAY 22 2020
 \section{Introduction}
-%(done)CID: Start the idea that inside each body is the following:
+The goal of this introduction is to explain the idea of genomes and their size and magnitude.
 Inside every human body are the instructions to create and maintain it.
 These instructions are found inside the genome.
-The entire human genome, when laid out is around 2 meters in length.
+The the instructions inside the human genome are incredibly long.
+The entire human genome,inside a singular cell, when laid out is around 2 meters in length.
 \autocite{ency_sci_tech}
-%(done) KING: cite this
-But once all of the DNA of all the cells in the human body are combined.
+
+But once all cells are accounted, when all of the DNA of all the cells in the human body are combined.
 The length can span the earth to moon 6000 times.
-This shows the length and amount of DNA inside the human body.
-This describes as well the vast amounts of information in the body.
+This describes as well the vast amounts of DNA information in the body.
+This highlights as well the importance of genomes and genomics.
+They constitute and contribute important information for all the cells in an organism.
 
-%CID: Add a paragraph elaborating further on the amount of information here.
+Defining \textit{genome}
+% CID PROGRESS ENDS HEREE FOR MAY 22 2020
+%------------------------------------------
+% CID PROGRESS STARTS HEREE FOR MAY 29 2020
 
-%(partial done) CID: This intro lacks references and needs better visual description
-
-The word \textit{genome} was used briefly. But first, what does genomes or genomics actually mean?
-
-\subsection{Definition}
+\subsection{Definition of Genomics}
 National Institute of Health \autocite{genomics-definition} defines genomics as \\ \\
@@
 
-%(partially done) CID: Mej abrupt pagpasok kaagad sa genomics na definition. Walang clean transition
-
-In the definition, there was also a word called genes.
-But what are genes?
-Genes comprises of DNA.
+In the definition, one main component of genomics are genes.
+Genes comprise of DNA.
 DNA is an acronym that stands for deoxyribonucleic acid.
-DNA is usually characterized by the letters A,T,G,C.
-Where A is for Adenine, T is for Thymine, G is for Guanine, C is for Cytosine.
-These are the instructions passed on from one generation to another, this is also what is used to form the organism.
+These are more specific information in the cells of organism.
+DNA is usually denoted by the letters A,T,G,C.
+These letters symbolize the chemical compounds (or nucleotides) that encode information in the DNA.
+Another important feature of genes is thar these  are the instructions passed on from one generation to another.
 \autocite[p.~2]{hartl2018}
 
-%CID: Explain what the acronym DNA means. Then describe why that acronym becomes ATGC
+These nucleotides form the base pairs (or \textit{bases}).
+These bases encode information as stated earlier.
+A clearer picture is that these bases combined act as a blueprint to describe and instruct the body on how to build it and function.\autocite{alberts_mole}
 
-These nucleotides
-%(done)KING: Bases are the nucleotide
-%CID: Explain this part more, it doesnt really explain the importance
-form the base pairs (or \textit{bases}). These bases act as a blueprint to describe and instruct the body on how to build it and function.\autocite{alberts_mole}
-%CID: Expound this sentence more, perhaps explain the build and functioning process
+If the bases, genes, genomics combined form the blueprint of the organism.
+There is a need to understand the complexity and interaction to further understand how an organism works.
+Genomics play a role in understanding different components of an organism by paying attention to the genes and nucleotide bases.
 
-There are these components, and these describe complex organisms.
-Which means there is a long set of instructions to detail these complex organisms.
+The instructions inside the genes describe the complexity which describe an organism.
+This also means more complex organisms have more complex genes and therefore have a lot more information.
+Meaning the moment the genomics of complex organisms are studied, more information will be needed and gathered.
+Humans, being complex, should also have a long set of instructions to detail the complex behaviors.
 For humans, the full human genome contain around $3.2x10^9$ base pairs \autocite{introgenomics}.
 This is a lot of information inside the human organism.
-But before analyzing that information, these have to be taken from the cells and read.
 
-%CID: Separate the paragraphs to explain what genes are, to the paragraph explaining the complexity of the human genome (also elaborate on this)
+Storing information in the information is one aspect, but before analyzing that information, these have to be taken from the cells and read.
+Reading these information from genes will need certain technology.
+The technology is called \textit{Sequencing}.
 
@@
 
-%(partially done) CID: Have a clear transition to sequencing, you ended at complexity.
-To see the information of the DNA of a cell.
-There are machines called \textbf{Sequencers} that look (or more appropriately, \textit{read}) the DNA of a specimen and outputs A,G,C,T.
+Reading genetic information found in the DNA uses the technology of Sequencing.
+The machines are aptly called \textbf{Sequencers} that look (or more appropriately, \textit{read}) the DNA of a specimen.
+The machine outputs A,G,C,T, sometimes with uncertainties.
 Sequencers look at the exact order of a single strand of DNA.
-The process of sequencing DNA isn't a simple process of a machine reading one strand of DNA.
-In most cases, one strand is too little for one machine to read.
-So before reading, the DNA is replicated multiple times in the laboratory.
-DNA is also fragmented before processing.
-Fragmented meaning it will be cut up into pieces.
+The process of sequencing DNA seems straightforward that a machine reads only one strand of DNA.
+In actuality, one strand is too little for one machine to read.
+So before reading, the DNA is \textit{replicated} multiple times in the laboratory.
 
-%CID: Describe what fragmented means
+Sequencers also can't read the entire length of DNA.
+The DNA is also \textit{fragmented} before processing.
+By fragmented meaning it will be cut up into pieces.
+These are fragmented since the technology cannot handle reading the entire length of the DNA.
+Sequencers can only read it in parts.
 
-These are fragmented since the machine cannot handle reading the entire length of the DNA, it has to read it in parts.
-Sequencers then reads (or sequences) them.
-These reads of fragments from the replicated DNA will then be put back together computationally to form the \textit{correct or reference} DNA.
+The idea of \textit{replication} and \textit{fragmentation} are important because these increase the data taken up when sequencing.
+Replication causes there are a lot of redundant data when the sequence.
+Fragmentation causes there to be a need to piece back the information together.
+Now with those concepts in mind sequencers then reads (or sequences) the replicated and fragmented data.
 
-%CID: The paragraph on sequencing is good. Maybe cite more.
+These reads or sequences taken will need an extra step before the data becomes useful.
+It will then be needed to be put back together computationally to form the \textit{correct or reference} DNA.
+These are one of the reasons why genomic data is large because even the fragmented replicated data is still stored, not just the reference DNA.
 
-Back in 2003 The first human genome \autocite{introgenomics} had been first completely read.
+\subsection{Human Genome Project}
+
+Another way to see the way genomic data is large is by looking at the \textit{Human Genome Project}.
+The project's aim was to sequence the entire human genome.
+At the time it needed a lot of collaborations, a lot of financing, and a lot of sequencing, and new computation methods.
+Back in 2003 was a milestone because the first human genome \autocite{introgenomics} had been first completely read.
 It took the project ten years of work, a dozen institutions, and 3,000,000,000 dollars to finish.
-Currently, in 2019, the technology today in Japan is much faster than the technology back in 2003.
+The technology for sequencing and genomics was still at its infancy.
+
+\subsubsection{Next Generation Sequencing}
+
+%(done) CID: Describe the speedup in terms of what and how the speedup was computed
+
+The introduction of Next Generation Sequencers can be seen as a another significant event in the history of genomics.
+This is a new technology which has allowed the machines to read faster and cheaper.
+These utilize a variety of technologies, some make use of light, some use ion semiconductors, others use a special DNA marker with a fluorescent probe \autocite[~p.67]{paulselzer2018}.
+
+Many countries that research have started employing NGS.
+Sixteen years later after the HGP, the technology in Japan is very much faster than the technology back in 2003 used for the HGP.
 The technology in Japan can read 10,000 human genomes per day.
 \autocite[p.~19]{introgenomics}
-%(done) KING: Cite
 
-% (partially done) CID: Get the transition of the NGS data graph from the slides
+These show how the technology for reading genomes has been developed over the years.
+This tech has increased the read speed.
+Which directly affects the speed in which data is produced.
+Since each read requires storage.
+This poses a problem to projects involving sequences, the data volumes is usually very large\autocite{bon_compression}.
+
+There exists a concept of \textit{coverage} as well.
+Referring back to the \textit{replication} and \textit{fragmentation} concept earlier.
+There needs to be a specific number of reads a machine needs in order to produce a reference sequence (the correct, standard sequence).
+Meaning the sequence isn't read only once.
+
+This contributes to the large data size of sequences.
+Currently, for a coverage of around 60, the file size for a compressed human genome can easily reach 200 GB.
+And a project of 10-20 genomes would need approximately 4 TB of storage.
+Transferring these data across research groups would indeed be a challenge.
+\autocite[~p.68]{paulselzer2018} These highlight the important and growing problem of storage in genomic research.
 
@@
 The big fall of the prices around 2004 - 2005 corresponds to the introduction of the Next Generation Sequencers.
-
-\subsubsection{Next Generation Sequencing}
-
-%(done) CID: Describe the speedup in terms of what and how the speedup was computed
-
-The introduction of Next Generation Sequencers can be seen as a significant event in the history of genomics.
-This is a new technology which has allowed the machines to read faster and cheaper.
-These utilize a variety of technologies, some make use of light, some use ion semiconductors, others use a special DNA marker with a fluorescent probe \autocite[~p.67]{paulselzer2018}.
-This tech has increased the read speed and speed in which data is produced.
-But note as the data increases, storage demand therefore also increases.
-Since each read requires storage.
-This poses a problem to projects involving sequences, the data volumes is usually very large\autocite{bon_compression}.
-There exists a concept of \textit{coverage} as well.
-It means the number of reads it needs in order to produce a reference sequence (the correct, standard sequence).
-Meaning the sequence isn't read only once.
-This contributes to the large data size of sequences.
-Currently, for a coverage of around 60, the file size for a compressed human genome can easily reach 200 GB.
-And a project of 10-20 genomes would need approximately 4 TB of storage.
-Transferring these data across research groups would indeed be a challenge.
-\autocite[~p.68]{paulselzer2018}
-%(done) KING: cite
-% more space is used.
-
-%(done) CID: Visually explain how more space is needed (cite numbers). May separate the sentences describing NGS and space utilization
 
@@
 This lowering of cost, makes the idea of sequencing more available.
-And thus produces a lot more data.
-But these terabytes of data need to be stored somewhere.
+The number of  research projects should also increase as well contributing to the growing number of genomic data.
 
@@
 
-This means, that as time passes by, technology for processing and storage will improve.
+The idea of Moore's law should give relief to genomic researchers since this means, that as time passes by, technology for processing and storage will improve.
 And the overall capacity for what technology can do will increase as well.
-But the speed in which the number of sequence data increases seems to exceed Moore's Law.
-And this will pose a problem since the technology doesn't seem to be catching up.
+
+But, by paying attention to the speed in which the number of sequence data increases seems to exceed Moore's Law.
+And this will pose an additional problem since the technology doesn't seem to be catching up.
 The immediate problem is the storage of all the data being produced.
 A certain technology needs to be made for the purpose of holding this data.
-This technology is referred to as Databases.
+This technology is called \textit{Databases}.
+But there is a greater need well in making sure the data stays secure, consistent, easily accessible, and fast.
+These are other aspects needed by a genomic researcher as well.
+And this is the reason why a specific class of Databases called \textit{Distributed Databases}.
+Will be needed in this paper.
 
-%CARL: change "flow" of ideas, maybe reorder/rewrite subsection names
-% currently, the subsections are:
-% - Database
-%   - Principles
-%   - Database Types
-% - App Archi
-%   - CS
-%   - P2P
-%   - Hybrid
-
-\subsubsection{Definition}
+\subsubsection{Databases Definition}
 A database (DBMS, Database Management System) is defined by containing information about a particular enterprise. Functionalities include: \autocite{Silberschatz2010}
@@
 These form a system that is made for accessing and updating data.
+The next section discusses the key points when considering how to choose or create a good database system (or DBMS).
+The key points are important to discuss many important features of the proposed system later.
+Certain database principles are needed to make sure the data is available, secure, and consistent.
 
-%CID: Summarize a bit on why this was discussed and why this is needed in the thesis
-
-\subsubsection{Principles} 
-There are also 3 principles to remember when creating a database.
-The idea is that only 2 of the three can be focused on at any given time.
-And these need to be taken into consideration when designing the system.
+\subsubsection{Principles of Databases} 
+There are 3 principles needed when creating a database.
+The idea is that only two of the three can be focused on at any given time.
+Only two since the current technology can only handle two.
+These principles and limitations need to be taken into consideration when designing the system.
 \autocite{Silberschatz2010}
@@
 \begin{itemize}
-    \item Consistency - Update one part, will all other (replicated) parts be updated quickly as well; Data is accurate and copied faithfully.
-An example would be that genes for specimen A, when copied to another data node should contain the exact copy of the data, with no minor changes.
-    \item Accessibility - How easy it is to access data, API protocols. An example would be the number of commands  a user needs to do to access every data.
-    \item Partition Tolerance - If one portion breaks, the other portions are still usable. An example would be a power failure for one data node, how would this affect the other nodes?
+    \item Consistency - Does the data have integrity?
+To update one part, all other (replicated) parts should be updated quickly as well; Data is accurate and copied faithfully.
+An example would be that genes for specimen A, when copied to another data node should contain the exact copy of the data, with no changes.
+    \item Accessibility - How easy it is to access data, API protocols? An example would be the number of commands  a user needs to do to access every data.
+    \item Partition Tolerance - How resilient is the system when it comes to failures?
+If one portion breaks, the other portions should still be entirely usable.
+An example would be a power failure for one system, how would this affect the other systems?
 \end{itemize}
-%CID: Use explanations from the slides
-%CID: Explain CRUD
 
@@
 Two types of databases will be discussed in this paper.
-Some of these databases will use technical terms, these terms will be discussed later in the Theoretical Framework.
-Namely: \autocite{centralizedvsdistributed}
+Each have their specific niche and use.
+One is more efficient and easier to make, another is better suited for collaborative work.
+Note some technical jargon needs to be used to explain some concepts.
+These are namely: \autocite{centralizedvsdistributed}
 
@@
     \item Centralized - In a centralized database, all data are managed by a single DBMS in a single node to be distributed to users.
-As a result, transactions are easily manageable because everything can be processed in the single node, but the cost of communication is high and reliability is low, as an error in the database will mean disaster to the network as a whole.
+As a result, transactions are easily manageable because everything can be processed in the single node.
+Here, the cost of communication is high and reliability is low, as an error in the database will mean disaster to the network as a whole.
+These are easier to create.
+
     \item Distributed - In a distributed database, data are stored in multiple servers in different geographical locations.
@@
 The difficulty of managing transactions will go up, but the result is higher availability, faster response, and modular growth, among other advantages.
+These databases are better creating spread out databases, and are better for collaborative work.
 \end{itemize}
 
-%\begin{itemize}
-%    \item Client-Server Database - Utilizes a client-server architecture. One server to sends and receives the data to all the clients. 
-%    \item P2P Database - Utilizes a peer architecture. Each peer has the ability to send and receive the data to all the other peers. There is no hierarchy.
-
-    %KING: One server that sends
-    %KING: One server for sending
-    %KING: One server sends
-    %CID: Basically reexplain client and server architecture, wha does client mean, what does server mean
-
-%    \item Hybrid (Or Distributed) Database - Utilizes master node and data node.
-Unlike the normal database, this maintains a different structure, where not only one server has all the data.
-And a concept of replication of data is happening.
-%\end{itemize}
-
-%CID: I dont see any problems here, perhaps cite nalang a bit more
-
-%CID: Explain muna P2P CS H in intro.
+For collaborative genomic work, especially in this paper, distributed databases are most appropriate.
 
 \subsection{Application Architecture}
-The application architecture is how a networked app is structured
-across the various end systems (i.e. computers, servers, etc.) \autocite{kurose}
-Falls into either client-server, peer-to-peer, or hybrid architecture
-%CID: Explain a bit how application architecture affects the app? Like does this change the features or speed or traffic management?
+Another important consideration is how a system is built.
+These describe the blueprint and interaction within a database system.
+These provide insights similar to the questions asked when looking between database types.
+Database types describe how the data is handled inside the system.
+But application architecture describes how the users interact with the system.
+The application architecture is how a networked app is structured across the various end systems (i.e.
+computers, servers, etc.) \autocite{kurose} The end systems are another way to call users or clients.
+The application architecture types fall into either client-server, peer-to-peer, or hybrid architecture.
 
@@
 available and has a fixed IP address.
-When a client sends a request for data, the server responds by sending the requested data.
-The client-server architecture may not be able to keep up with the increasing number of connections.
+When a client (or user) sends a request for data, the server (or DBMS in this example) responds by sending the requested data.
+The client-server architecture is mainly built to respond to one client at a time.
+The architecture may handle multiple connections but as these grow bigger, it may not be able to keep up with the increasing number of connections since the server might not be able to handle all the requests.
 
@@
 always-on, dedicated servers.
-Peers are desktops/laptops that communicate to each other without passing through a dedicated server.
-These include internet telephony, BitTorrent, and some other file sharing protocols.
-An advantage of P2P networks is their scalability.
+Peers are desktops/laptops (or users) that communicate to each other without passing through a dedicated server (or DBMS).
+These include internet, BitTorrent, and some other file sharing protocols.
+An advantage of P2P networks is their scalability, unlike the earlier type.
 Each user can benefit the whole network by sharing his/her own resources.
-However, P2P has its own problems, namely security and incentivizing users to share their resources.
+
+However, since P2P doesn't have a central data management system. P2P has its own problems, which are security and incentivizing users to share their resources.
 
@@
 
-%CID: Explain some problems din kaya?
-%CARL: wala sa kurose eh what if gusto ko magdagdag ng sarili ko haha
-
-%KING: Fix periods in a sentence
-
-% this is only a particular implementation of hybrid kasi so cinomment out ko
- %A master node stores metadata of all sequences, may or may not contain sequence files.
-Data nodes stores sequence files and metadata of all sequence files in its filesystem.
-Both master node & data nodes have user interface that allow users to upload, search, and download sequence files.
+For a collaborative genomic system, it would be best to utilize Hybrid Application Architecture.
+Since there needs to be certain features that Client-Server of P2P might not be able to deliver.
+One would be scalability, another is security and centralized management of the data.
 
@@
 
-We review the technologies on data storage done before. 
+We review the technologies on data storage done before. Some have employed \textit{Client-Server}, some \textit{P2P}, others have tried \textit{Hybrid} (but are still few and unsuccessful).
 %CID: Make a transition to go from the diff database types to the explanation of the diff databases
@@
 \subsection{SeqTorr: A distributed scalable database for genomic information}
+The focus on SeqTorr was finding a way to scale databases meant specifically for genomic information.
+There was also a focus on being able to choose information to share and use.
 The standard genomics workflow consists of uploading and downloading data on an international database like NCBI.
@@
 
-SeqTorr consists of a master node and multiple data nodes.
+The architecture of SeqTorr consists of a master node and multiple data nodes.
 The master node contains metadata of all the sequences and is where the user authentication is handled, while the data node is where the actual sequences are stored.
@@
 
-Currently, SeqTorr is not available, and the code base does not have any documentation which makes future work difficult.
+SeqTorr seems to go in the right direction and focus for genomic databses. But currently, SeqTorr is not available, and the code base does not have any documentation which makes future work difficult.
 
 \subsection{BioTorrents}
+BioTorrents focused on P2P and making it easy to share data across biological scientific communities.
 Scientific data continues to grow, and so does the demand for easier accessibility.
-Centralized servers using HTTP or FTP cannot keep up with concurrent requests, and peer-to-peer protocols does not scale well for large files.
+Centralized servers, as stated earlier, using HTTP or FTP cannot keep up with concurrent requests, and peer-to-peer protocols does not scale well for large files.\autocite{biotorrents} One example of these centralized servers is NCBI, which will be discussed later.
 BitTorrent handles both these problems.
 
-BioTorrents is a system for legally sharing scientific data that works on top of the BitTorrent protocol.
+BioTorrents is a system for legally sharing scientific data that works on top of the BitTorrent protocol (or P2P).
 It is much like a public tracker.
 The main website hosts .torrent and metadata files.
-Info about each dataset includes categories, license, filenames, etc.
+The main website doesn't store data, rather it lets the users keep the data and share it to other users.
+The main website only facilitates the sharing of these data (via the idea of P2P and torrents).
+
+The main website stores info about each dataset includes categories, license, filenames, etc.
 that helps users in searching for relevant datasets.
-The torrent file has a list of trackers, servers that "know" which peers are serving which files, so it can download from those peers simultaneously.
-To upload a torrent, the client makes a .torrent file using the torrenting software, then uploads it to the main website along with metadata.
-The client must have their computer continuously turned on to keep the file available as a peer.
-The main issue with BioTorrents is that it requires lots of users that will keep many torrents available as often as possible.
-\autocite{biotorrents}
+There is also a torrent file has a list of trackers, servers that "know" which peers are serving which files, so it can download from those peers simultaneously.
+These torrents act like a directory telling the user who currently has the data for a piece of the file.
+The system then helps the user download many pieces of the file from many different users.
 
-%CID: Have a paragraph explaining the problems of BioTorrents
-However, BioTorrents has been down for years. As with all public trackers, it requires users to always keep the data publicly available, and the datasets must be moderated to prevent abuse.
+The client (or user) must have their computer continuously turned on to keep their file available to others.
+The main issue with BioTorrents, which is a problem as well in P2P, is that it requires lots of users that will keep many torrents available as often as possible.\autocite{biotorrents} These entail that for the system to work, many people need to use it and all these people need to keep their computers on all the time.
+
+BioTorrents was successful, many people had used it to share data.
+It was active for a while.
+However, BioTorrents has been down for years.
+This also shows what problems a system will face when doing a pure P2P application architecture.
 
@@
 
-%CID: Have a paragraph explaining the problems of PeerDB
-% The issues faced by PeerDB is ...
+PeerDB is a proposal for a distributed system using P2P.
+For this to be used in the context of genomic biological work, a main website which facilitates sharing will be need to be put up.
+PeerDB needs to be installed as well on all the users who will be participating.
 
-%CID: Explain NCBI
+A future PeerDB system will also face similar problems faced by BioTorrents since it will need to rely on numerous active users for it to work.
 
 \subsection{NCBI}
+NCBI was discussed briefly in the section on \textit{BioTorrents}.
+It uses a client-server architecture.
+Users may download data via the graphics interface (website), or may be done via FTP or HTTP \autocite{biotorrents}.
+The latter is preferred since it may be done to automate the download of data.
 
-%CID: Discuss further how NCBI works
-%This is how NCBI works... (Insert Sir El King's demonstration of how to bulk download).
+The SeqTorr authors\autocite{seqtorr} stated that NCBI is the main and current DBMS used to get and share genomic data.
+But one difficulty using that is updating the current genomic data.
+There currently is no way to check which data (on your computer) has been updated, which sequences were corrected.
+And the way the authors solved this was to regularly re-download the entire NCBI database.
 
-%CID: Explain a few problems faced by NCBI
-%The problems seen based on bulk downloading are ...
+To state again some problems of NCBI, it has the limitations of \textit{Client-Server} architecture meaning the speed of the download is heavily reliant on the speed of the main NCBI server.
+Which also puts a lot of stress on the main server.
+Another problem discussed earlier was the updating of data, which have no technology yet to do so.
 
-%describe more the problems per database
-%normal vs p2p vs dist databases
-
-%CID: Explain muna P2P CS H in intro.
-%CID: Have a sentence explaining this table in word form and adding all the other problems discussed above
+% CID PROGRESS ENDS HEREE FOR MAY 29 2020
+% CID PROGRESS STARTS HEREE FOR JUNE 2 2020
 
@@
 
-% (done) KING: 1) Needs a table header
-%(done) KING: 2) Fix latex positioning, show the table after the PeerDB Section
-%(done) KING: 3) When you show a table / figure, make sure you reference that somewhere in the discussion
-
 \section{Problem}
-%CID: Have some discussion then somehow the discussion concludes to the idea that the following problems occur, these problems mej came out of nowhere kasi
-%CID: Expound more on each category
-
-The different gaps or problems found with other implementations
+To restate the different gaps or problems found with other implementations
 \subsection{Problems of Classical Databases}
 \begin{enumerate}
-\item Genome researchers download entire database of genomic data
-\item Genome researchers only need a portion of the database, but end up downloading everything
+\item Genome researchers download an entire database of genomic data, especially when uncertain how to update their data
+\item Genome researchers only need a portion of the database, but end up downloading the entire database
 \item Client-Server approach can face bottlenecks (server speed)
+\item P2P approach has a need of a large unmoderated active user base to work
+
 \end{enumerate}
@@
 \begin{enumerate}
-    \item No existing usable framework on distributed database for genomes
-    \item Existing distributed database for genome code has no documentation to start with % concept of APIs and documentation
+    \item No existing public usable framework on distributed database for genomes
+    \item Existing distributed database for genome code has no documentation
 \end{enumerate}
-%CID: Evaluate if these problems are accurate. Maybe cite earlier paragraphs?? I dont know if that's possible
 
@@
 
-%CID: Have a discussion, like what I said sa problem to make the reader want to say the objectives we will state. Like we want to lead the reader to what we will say
-%CID: Maybe restate the problems super briefly, and then discuss what tech is available
-%CID: Explain more on the metrics to be used
-
-Based on the problems we found earlier, we propos to:
-%CID: Hmmm parang okay naman, the wording nalang
+Based on the problems we found earlier, we propose to:
 
 \begin{itemize}
-    \item Create an information system that speeds up the download and upload of genomic data
+    \item Design the framework for an information system that speeds up the download and upload of genomic data
     \begin{itemize}
@@
         \begin{itemize}
-            \item This will aim to resolve number 3 in the Problems of Classical Databases
+            \item This will aim to resolve and speed up number 3 in the Problems of Classical Databases
         \end{itemize}
@@
     \end{itemize}
-    \item Write documentation for the system
+    \item Write a documentation for the system
     \begin{itemize}
@@
     \end{enumerate}
-\end{enumerate}
 

### 2020-06-18 Paper2020ACM/Paper2020.tex
Initial Overleaf Import

%%
%% This is file `sample-acmsmall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmsmall]{acmart}

% NOTE that a single column version is required for submission and peer review. This can be done by changing the \doucmentclass[...]{acmart} in this template to 
%\documentclass[manuscript,screen]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}
\acmDOI{10.1145/1122445.1122456}

%%
%% These commands are for a JOURNAL article.
\acmJournal{JACM}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmMonth{6}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{caDDS: Community Assisted Distributed Database for Sequences}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Carl Araya}
\authornote{Both authors contributed equally to this research.}
\email{jjazcarraga@up.edu.ph}
\affiliation{%
  \institution{University of the Philippines - Diliman}
  \city{Quezon}
  \country{Philippines}
}

\author{Jose Carlos Rodrigo J. Azcarraga}
\authornotemark[1]
\affiliation{%
  \institution{University of the Philippines - Diliman}
  \city{Quezon}
  \country{Philippines}
}
\email{jjazcarraga@up.edu.ph}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Araya and Azcarraga}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
300 words
- purpose and problems:
    - design a framework
    - design a database
- basic design:
- system summary:
    - master node and data node
- conclusion:
    - time complexity wise: 
    - proposed system has this pros
    - based system has this pros

working draft (insert kulang when info is ready):
 Genomic data the past few years have grown exponentially with the technology barely keeping up. The problem faced by genome researchers is the large data set and difficulty transferring the files. Previous distributed databases are either not meant for genomic data, or difficult to replicate. This paper lays the groundwork for a distributed database that is designed to easily scale to accommodate bigger storage, and also reduces the data speed bottleneck faced by centralized databases. The proposed system has a master node and data nodes. Data nodes should theoretically increase data transfer speeds even as the number of users increase. (to be updated based on conclusion) While current databases(e.g. NCBI) have more data, more users, and a live platform. The proposed system is aimed towards a smaller community, with more frequent and localized data, which the data storage can easily be expanded as the need arises. A recommendation is to implement the system and make the code public for the use of other communities.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10010405.10010444.10010093</concept_id>
       <concept_desc>Applied computing~Genomics</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002951.10002952.10002953</concept_id>
       <concept_desc>Information systems~Database design and models</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10003120.10003130</concept_id>
       <concept_desc>Human-centered computing~Collaborative and social computing</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Applied computing~Genomics}
\ccsdesc[500]{Information systems~Database design and models}
\ccsdesc[500]{Human-centered computing~Collaborative and social computing}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{genomics, distributed, database, community}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% START OF PROGRESS FOR CID MAY 22 2020
\section{Introduction}
The goal of this introduction is to explain the idea of genomes, their size and magnitude. Inside every human body are the instructions to create and maintain it. These instructions are found inside the genome. The instructions inside the human genome are incredibly long. For example, entire human genome, inside a singular cell, when laid out is around 2 meters in length.\cite{ency_sci_tech}

Combining  and laying out all the information for all the cells in the human body. The length is incredibly long that it can span the earth to moon 6000 times. This describes as well the vast amount of DNA information in the body. These kinds of information will help scientists explain many of the understood and unknown processes inside all organisms. These may also be used to compare different organisms to each other. And to compare humans to other humans.

Defining \textit{genome}
% CID PROGRESS ENDS HEREE FOR MAY 22 2020
%------------------------------------------
% CID PROGRESS STARTS HEREE FOR MAY 29 2020

\subsection{Definition of Genomics}
Genomics is is the study of wholes sets of genes and their interactions\cite{campbell}. The efforts to study this field has led to volumes of data that need to be analyzed. Which has also spurred the field bioinformatics, which aims to use computers to aid the analysis and storage of biological data. Institutions such as the National Institute of Health and National Library of Medicine, have led this field and built a website to encourage collaboration online through their online portal called NCBI (National Center for Biotechnology Information). \cite{campbell} This website will be discussed later in more detail.

National Institute of Health \cite{genomics-definition} formally defines genomics as \\ \\
\textit{the study of all of a person's genes (the genome), including interactions of those genes with each other and with the person's environment.} \\ \\

In the definition, one main component of genomics are genes. Genes comprise of DNA. DNA, deoxyribonucleic acid, is a chemical compound that works like a computer bytes to store information for organisms. DNA's structure is composed of chemical compounds (or nucleotides) that encode information. These nucleotides are commonly written down as A,G,C,T corresponding to their chemical makeup. Another important feature of genes is that these genes are the instructions passed on from one generation to another. \cite[p.~2]{hartl2018}

These nucleotides form sets of base pairs (otherwise known as \textit{bases}). These bases encode the information stated earlier. They are pairs since, there is always a corresponding complement pair for every nucleotide. To give a bigger picture, when these bases are combined and organized, these act as a blueprint to describe and instruct the body(or organism) on how to create it and how to maintain it.\cite{alberts_mole}

If all the bases, genes, genomics combined form the blueprint of the organism. There is useful knowledge there and there will be a need to understand all the complexity and interactions to further understand how the blueprint of an organism works. Genomics play a role in understanding different components of an organism by paying attention to the genes and nucleotide bases. Different organisms have certain differences in genes and this has been a focal point of many genomic studies, which sometimes is aimed as well as establishing surprising similarities among distinct organisms.

The instructions inside the genes explain the complexity of an organism. This also means more complex organisms are likely to have more complicated genes and therefore have a lot more information. Even if this general idea is true, there are some exceptions, some plants such as the Japanese canopy plant \textit{Paris japonica} have 50 times the genome size compared to humans\cite{campbell}. These may be due to repeating non-essential DNA. But the idea of having many unknown components and interactions in a complex organism means more information will be needed and gathered. Humans, being complex, are extensively studied in order to detail the complex processes inside the body. The presence of many studies for the human genome has increased the amount of information generated. To give an idea of the amount of information, for humans, the full human genome contain around $3.2x10^9$ base pairs \cite{introgenomics}.

Storing information is one aspect, but before analyzing that information, these have to be physically taken from the organism's cells and read. Reading these information from genes will need a certain technology. The technology is called \textit{Sequencing}.

\subsection{Sequencing}

The machines to generate the sequences are aptly called \textbf{Sequencers} that look (or more appropriately, \textit{read}) the DNA of a specimen. The machine outputs A,G,C,T, sometimes with uncertainties. Sequencers look at the exact order of a single strand of DNA. The process of sequencing DNA seems straightforward that a machine will read only one strand of DNA. In actuality, one strand is too little for one machine to read. So before reading, the DNA is \textit{replicated} multiple times in the laboratory. Here there is a concept of repeated reading or coverage which will be discussed later.

Sequencers also can't read the entire length of DNA. With this, the DNA is also \textit{fragmented} before processing. By fragmented meaning it will be cut up into pieces. Sequencers can only read it in parts.

The idea of \textit{replication} and \textit{fragmentation} are important because these increase the data taken up when sequencing. Replication causes a lot of redundant data in the sequence. Fragmentation causes there to be a need to piece back the information together. Now with those concepts in mind sequencers then reads (or sequences) the replicated and fragmented data.

These reads or sequences taken will need an extra step before the data becomes useful. It will then be needed to be put back together computationally to form the \textit{correct or reference} DNA. These utilize many of the technology developed by bioinformatice. These are one of the reasons why genomic data is large because even the fragmented replicated data is still stored, not just the reference DNA.

\subsection{Human Genome Project}

Another way to see the way genomic data is large is by looking at the \textit{Human Genome Project (HGP)}. The project's aim was to sequence the entire human genome. At the time it needed a lot of collaborations, a lot of financing, and new computational tools (piecing everything together). Back in 2003 it was a milestone because the first human genome \cite{introgenomics} had been first completely read. It took the project ten years of work, a dozen institutions, and 3,000,000,000 dollars to finish. The technology for sequencing and genomics was still at its infancy.

Once the Human Genome Project was completed in 2003 and the studies fully published in 2006. The research then paved the way to study chimpanzees since their DNA is the closest to humans. And the whole genome structure of humans can easily be used to recreate the whole genome sequence of the chimpanzee. This was completed 2 years after the Human Genome Project\cite{campbell}. These paved the way for more and more whole genome studies of animals much different from humans. By April 2013, around 4,300 genomes had been studied\cite{campbell}. These contribute to the growing amount of research, analysis, and most importantly data around genomes of different organisms.

\subsubsection{Next Generation Sequencing}

The introduction of \textit{Next Generation Sequencers or (NGS)} can be seen as a another significant event in the history of genomics. This is a new technology which has allowed the machines to read faster and cheaper. These utilize a variety of technologies, some make use of light, some use ion semiconductors, others use a special DNA marker with a fluorescent probe \cite[~p.67]{paulselzer2018}.  Back in 1980 the speed in which sequencing can be done was 1000 base pairs a day. In 2000 when the human genome project had been ongoing, the speed grew to 1000 base pairs every second, 24 hours a day, 7 days a week \cite{campbell}.  In 2000 the technology had been growing faster and more reliable.

NGS has contributed to this speedup even further. Many countries that research have started employing NGS. The NGS technology in Japan can now read 10,000 human genomes per day. \cite[p.~19]{introgenomics} Compare this to the many years and institutions it took to study the first human genome back in 2003. Now a singular laboratory can achieve this with speed.

These show how the technology for reading genomes has been developed over the years. This tech has increased the speed of reading, and as well as research. Which directly affects the speed in which genome data is produced. Since each sequence read requires storage. This poses a problem to projects involving sequences, causing the data volumes to be usually very large\cite{bon_compression}.

There exists a concept of \textit{coverage} as well. Referring back to the \textit{replication} and \textit{fragmentation} concept earlier. There needs to be a specific number of reads a machine needs in order to produce a good quality reference sequence (the correct, standard sequence). Meaning the sequence isn't read only once, where multiple reads need multiple reads.

This contributes to the large data size of sequence reads. Currently, for a coverage of around 60, the file size for a compressed human genome can easily reach 200 GB. And a project of 10-20 genomes would need approximately 4 TB of storage. Transferring these data across research groups would indeed be a challenge. \cite[~p.68]{paulselzer2018} These highlight the important and growing problem of storage and file transfers in genome research.

\begin{figure}[h]
\centering
\includegraphics[width=0.65\textwidth]{images/human-gen-cost.jpg} 
\caption{Cost per Genome Data Over Time}
\label{fig:human_gen_cost_fig}
\end{figure}

To view these technology developments over time, we look at the Figure~\ref{fig:human_gen_cost_fig} on Page~\pageref{fig:human_gen_cost_fig} \cite{genomics-cost}, we note here the expensive cost at the initial stages of sequencing the human genome. But as time progresses, the cost substantially went down to what it is now. The big fall of the prices around 2004 - 2005 corresponds to the introduction of the Next Generation Sequencers.

\subsection{Database}

\begin{figure}[h]
\caption{Megabase Cost Over Time}
\centering
\includegraphics[width=0.65\textwidth]{images/seq-cost.jpeg} 
\label{fig:megabase_cost_fig}
\end{figure}

Similar to Figure~\ref{fig:human_gen_cost_fig} on Page~\pageref{fig:human_gen_cost_fig}, we also note the same pattern for the cost of sequencing (not just humans) in Figure~\ref{fig:megabase_cost_fig}\cite{genomics-cost}. After the human genome project, many more whole genome projects have been completed. And the cost of sequencing dropped 1000 fold from 2008 to 2012 alone \cite{bon_compression}. This lowering of cost, again corresponds to the idea of sequencing becoming more available. Increasing the number of research projects, therefore contributing to the growing number again of genomic data.

\begin{center}
\textit{Moore's law states that the number of transistors on a chip doubles every two years.}\cite{kurose}
\end{center}

The idea of Moore's law should give relief to genomic researchers since this means, that as time passes by, technology for processing and storage will improve. And the overall capacity for what technology can do will increase as well.

But, by paying attention to the speed in which the number of sequence data increases seems to exceed the speed of Moore's Law. And this will pose an additional problem since the technology doesn't seem to be catching up. The immediate problem is the storage of all the data being produced. A certain technology needs to be made for the purpose of holding this data. This technology is called \textit{Databases}. But there is a greater need as well in making sure the data stays secure, consistent, easily accessible, and fast. These are other aspects needed by a genomics researcher as well. And this is the reason why a specific class of Databases called \textit{Distributed Databases} will be needed in this paper.

\subsubsection{Databases Definition}
A database (DBMS, Database Management System) is defined by containing information about a particular enterprise. Functionalities include: \cite{Silberschatz2010}
\begin{itemize}
    \item Collection of interrelated data
    \item Set of programs to access the data
    \item An environment that is both convenient and efficient to use
\end{itemize}

These form a system that is made for accessing and updating data. The next section discusses the key points when considering how to  create a good database system (or DBMS). The key points are important to discuss many important features and dilemmas of the proposed system later. Certain database principles are needed to make sure the data is available, secure, and consistent.

\subsubsection{Principles of Databases} 
There are 3 principles needed when creating a database. The idea is that only two of the three can be focused on at any given time. Only two since the current technology can only handle two. These principles and limitations need to be taken into consideration when designing the system. \cite{Silberschatz2010}

\begin{itemize}
    \item Consistency - Does the data have integrity? To update one part, all other (replicated) parts should be updated quickly as well; Data is accurate and copied faithfully. An example would be that genes for specimen A, when copied to another data node should contain the exact copy of the data, with no changes.
    \item Accessibility - How easy it is to access data, API protocols? An example would be the number of commands  a user needs to do to access every data.
    \item Partition Tolerance - How resilient is the system when it comes to failures? If one portion breaks, the other portions should still be entirely usable. An example would be a power failure for one system, how would this affect the other systems?
\end{itemize}

\subsubsection{Database Types} 
Two types of databases will be discussed in this paper. Each have their specific niche and use. One is more efficient and easier to make, another is better suited for collaborative work. Note some technical jargon needs to be used to explain some concepts. These are namely: \cite{centralizedvsdistributed}

\begin{itemize}
    \item Centralized - In a centralized database, all data are managed by a single DBMS in a single node to be distributed to users. As a result, transactions are easily manageable because everything can be processed in the single node. Here, the cost of communication is high and reliability is low, as an error in the database will mean disaster to the network as a whole. These are easier to create.

    \item Distributed - In a distributed database, data are stored in multiple servers in different geographical locations. Each node may contain a part or the entirety of the data. The difficulty of managing transactions will go up, but the result is higher availability, faster response, and modular growth, among other advantages. These databases are better creating spread out databases, and are better for collaborative work.
\end{itemize}

For genomic research, when the data storage needs to be increased often, distributed databases are most appropriate.

\subsection{Application Architecture}
Another important consideration is how a system is built. These describe the blueprint and interaction within a database system. These provide insights similar to the questions asked when looking between database types. Database types describe how the data is handled inside the system. But application architecture describes how the users interact with the system. The application architecture is how a networked app is structured across the various end systems (i.e. computers, servers, etc.) \cite{kurose} The end systems are another way to call users or clients. 
The application architecture types fall into either client-server, peer-to-peer, or hybrid architecture.

\subsubsection{Client-Server}
A client-server architecture relies on a Web server that is always available and has a fixed IP address. When a client (or user) sends a request for data, the server (or DBMS in this example) responds by sending the requested data. The client-server architecture is mainly built to respond to one client at a time. The architecture may handle multiple connections but as these grow bigger, it may not be able to keep up with the increasing number of connections since the server might not be able to handle all the requests.

\subsubsection{P2P}
A peer-to-peer (P2P) architecture has minimal to no reliance on always-on, dedicated servers. Peers are desktops/laptops (or users) that communicate to each other without passing through a dedicated server (or DBMS). These include internet, BitTorrent, and some other file sharing protocols. An advantage of P2P networks is their scalability, unlike the earlier type. Each user can benefit the whole network by sharing his/her own resources.

However, since P2P doesn't have a central data management system. P2P has its own problems, which are security and incentivizing users to share their resources.

\subsubsection{Hybrid}
A hybrid architecture combines elements of P2P and client-server architectures, attempting to utilize advantages of both.

For a genomic database system, it would be best to utilize Hybrid Application Architecture. Since there needs to be certain features that Client-Server of P2P might not be able to deliver. One would be scalability, another is security and centralized management of the data, which is a feature not provided by P2P.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% RRL %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Review of Related Literature}

We review the technologies on data storage done before. Some have employed \textit{Client-Server}, some \textit{P2P}, others have tried \textit{Hybrid} (but are still few and unsuccessful).

\subsection{SeqTorr: A distributed scalable database for genomic information}
The focus on SeqTorr was finding a way to scale databases meant specifically for genomic information. There was also a focus on being able to choose information to share and use. The standard genomics workflow consists of uploading and downloading data on an international database like NCBI. Having a distributed scalable local infrastructure to store genomic data instead of relying on NCBI would be beneficial since: only certain sequences in NCBI are relevant to the institutions (e.g. Asian sequences are needed more than Caucasian ones), researchers within the country can share data with fast speeds, and  institutions can share in the hosting of data.

The architecture of SeqTorr consists of a master node and multiple data nodes. The master node contains metadata of all the sequences and is where the user authentication is handled, while the data node is where the actual sequences are stored. All user uploads happen in the master node. When the master node receives a sequence from a user, it distributes the sequence to any available data nodes. 
\cite{seqtorr}

SeqTorr seems to go in the right direction and focus for genomic databses. But currently, SeqTorr is not available, and the code base does not have any documentation which makes future work difficult.

\subsection{BioTorrents}
BioTorrents focused on P2P and making it easy to share data across biological scientific communities.  Scientific data continues to grow, and so does the demand for easier accessibility. Centralized servers, as stated earlier, using HTTP or FTP cannot keep up with concurrent requests, and peer-to-peer protocols does not scale well for large files.\cite{biotorrents} One example of these centralized servers is NCBI, which will be discussed later. BitTorrent handles both these problems.

BioTorrents is a system for legally sharing scientific data that works on top of the BitTorrent protocol (or P2P). It is much like a public tracker. The main website hosts .torrent and metadata files. The main website doesn't store data, rather it lets the users keep the data and share it to other users. The main website only facilitates the sharing of these data (via the idea of P2P and torrents).

The main website stores info about each dataset includes categories, license, filenames, etc. that helps users in searching for relevant datasets. There is also a torrent file has a list of trackers, servers that "know" which peers are serving which files, so it can download from those peers simultaneously. These torrents act like a directory telling the user who currently has the data for a piece of the file. The system then helps the user download many pieces of the file from many different users.

The client (or user) must have their computer continuously turned on to keep their file available to others. The main issue with BioTorrents, which is a problem as well in P2P, is that it requires lots of users that will keep many torrents available as often as possible.\cite{biotorrents} These entail that for the system to work, many people need to use it and all these people need to keep their computers on all the time.

BioTorrents was successful, many people had used it to share data. It was active for a while. However, BioTorrents has been down for years. This also shows what problems a system will face when doing a pure P2P application architecture.

\subsection{PeerDB}
% summarize intro
PeerDB is a proposed P2P system for sharing general data. Each node is equipped with a MySQL database that it keeps available for other nodes to download from. Each node also keeps track of its neighbors, and any useful information about them. Queries are passed around neighbors, with a maximum number of hops. A node may cache information about other peers to speed up future queries.
\cite{peerdb}

PeerDB is a proposal for a distributed system using P2P. For this to be used in the context of genomic biological work, a main website which facilitates sharing will be need to be put up. PeerDB needs to be installed as well on all the users who will be participating.

A future PeerDB system will also face similar problems faced by BioTorrents since it will need to rely on numerous active users for it to work.

\subsection{NCBI}
The National Center for Biotechnology Information or more commonly known as NCBI, which discussed briefly in the section on \textit{BioTorrents} and in the \textit{Introduction}. This is the standard genomic database used by many researches and institutions. We will refer this later on as well as the \textit{base system}. This is the website used to access and store most genomic information\cite{campbell}. It uses a client-server architecture. Users may download data via the graphics interface (website), or may be done via file transfer protocols such as FTP or HTTP \cite{biotorrents}. The latter is preferred since it may be done to automate the download of data.

The SeqTorr authors\cite{seqtorr} also stated that NCBI is the main and current DBMS used to get and share genomic data. But one difficulty using that is updating the current genomic data. There currently is no way to check which data (on your computer) has been updated, which sequences were corrected. And the way the authors solved this was to regularly re-download the entire NCBI database.

To state again some problems of NCBI, it has the limitations of \textit{Client-Server} architecture meaning the speed of the download is heavily reliant on the speed of the main NCBI server and the number of users accessing the server. Which also puts a lot of stress on the main server. Another problem discussed earlier was the updating of local user data, which have no technology yet to do so.

% CID PROGRESS ENDS HEREE FOR MAY 29 2020
% CID PROGRESS STARTS HEREE FOR JUNE 2 2020

\begin{table}[h]
    \caption{Comparisons of Different Databases}
    \label{table:database_comparison_table}
    \begin{tabular}{ccccc}
    \toprule
    System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ 
    \midrule
    BioTorrents & P2P           & No         & Depends                    & Biological  \\ 
    SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ 
    PeerDB      & P2P           & No         & Depends                    & General      \\ 
    NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\ 
    \bottomrule
    \end{tabular}
\end{table}

Table~\ref{table:database_comparison_table} on page ~\pageref{table:database_comparison_table} describes the different technologies studied. This will also be used later on to describe what are the main problems of the different databases.

\section{Problem}
To restate the different gaps or problems found with other implementations of Genomic Databases
\subsection{Problems of Classical Databases}
\begin{enumerate}
\item Genome researchers download an entire database of genomic data, especially when uncertain how to update their data
\item Genome researchers only need a portion of the database, but end up downloading the entire database
\item Client-Server approach can face bottlenecks (server speed)
\item P2P approach has a need of a large unmoderated active user base to work

\end{enumerate}

\subsection{Problems of Current Distributed Databases for Genomes}
\begin{enumerate}
    \item No public existing usable public framework on distributed databases for genomes
    \item Existing distributed database system for genomes is not replicable
\end{enumerate}

\section{Objectives}

Based on the problems we found described, we propose to:

\begin{itemize}
    \item Design the framework for an information system that speeds up the download and upload of genomic data
    \begin{itemize} 
        \item To resolve 1 \& 2 of the Problems of Classical Databases. 
    \end{itemize}
    \item  Design a distributed database system
    \begin{itemize}
        \item To resolve 3 \& 4 of the Problems of Classical Databases. 
    \end{itemize}
\end{itemize}

\begin{itemize}
    \item For the framework, the system should have the main API’s 
    \begin{itemize}
        \item POST genome data
        \item GET genome data (this will implement a hybrid P2P transfer)
        \begin{itemize}
            \item This will aim to speed up number 3 in the Problems of Classical Databases
        \end{itemize}
        \item PUT genome metadata (optional: PUT data)
        \item DELETE genome data
    \end{itemize}
\end{itemize}

\section{Scope}

The project will focus on creating a framework, and outlining key API features of the framework. The current scope of the project is as follows:

\begin{itemize}
    \item To create a framework for a proposed system
    \item To design the correctness of a system
    \item To look at the basic time complexity of the base system (NCBI) and the proposed system
    \item To compare the proposed system to the base system

\end{itemize}

The current scope and coverage of the project will NOT consider:

\begin{itemize}
    \item Security and Authentication 
    \item Specific technologies to create the system
    \item All the data stored in the system (but focus on main data only)
    \item Implementing the actual system (but pseudocodes to the author's best ability will be provided)
\end{itemize}

% CID PROGRESS ENDS HEREE FOR JUNE 2 2020

% CID PROGRESS STARTS HEREE FOR JUNE 3 2020
% REDID FORMAT
% CID PROGRESS ENDS HEREE FOR JUNE 3 2020

% CID PROGRESS STARTS HEREE FOR JUNE 4 2020

\section{Theoretical Framework}

\subsection{Networks}

An important aspect of the proposed system is the transfer of files to other computers. To achieve that, there are network technologies and systems developed. Each layer of the network has certain features, securities, and abstraction which contributes to the overall idea of sharing files across networks. Networks Layers are the abstraction of how computers pass data to each other. Understanding this is needed to maximize the speed in which data is transferred between computers. As well as to realize where certain problems or bottlenecks later on. A network, specifically the Internet, is comprised of the following levels of abstraction. \cite{kurose}
\begin{enumerate}
    \item Level 1: Medium used to send signals (physical, electrical signals of 0
and 1), through wire or by air
    \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together
different . Goal of sending a group of signals.
    \item Level 3: Introduction of networks: IP Address to uniquely identify
entities.
    \item Level 4: Management of error free data transmission: Sending a group
of messages via TCP, UDP, FTP
    \item Level 5: Abstraction of Level 4: Function calls programmers will use
to do level 4 processes.
\end{enumerate}

Level 1 through 3 focus on building a system for sending messages from one computer entity to another. Every level talks to another entity on the same level. And a higher level talks to the lower level. This later on will be important to understanding the scope of what the thesis will do. Level 1 and 2 can be vulnerable if anyone can read the data in the air, partly why HTTPS was developed to encrypt the data.

Level 4 and 5 is what the system will mainly use. In these systems, there are APIs commonly used to transfer files between computers on networks. These levels, compared to the levels earlier discussed also are more specific to the ideas of file transfer, checking if the files are error free, and more. Making sure files transferred are error free will be needed to do data synchronization later on.

In networks, when there is data being accessed over the network, it is not called C.R.U.D. (the way databases access data), instead it is called POST, GET, PUT, DELETE. POST is analogous to CREATE meaning the user submit data to be stored (in the server). GET is similar to READ but is in the context of reading a file over a network (from a server). PUT is analogous to UPDATE as DELETE remains as DELETE. These are codes used to denote network commands that do the same thing as the CRUD functions discussed earlier. This is important since there is an added layer of difficulty when it comes to transmitting data over a network.

% CID PROGRESS STARTS HEREE FOR JUNE 6 2020

\subsection{Summary of the System}

The system is essentially a database. The database stores sequence data. It has the normal CRUD functions meaning Create, Read, Update, Delete. It can let anyone download sequence data. But only registered users can upload sequence data. When a user uploads data, the moderator will verify the data's authenticity before allowing it to be uploaded.

Then for the downloading of the data. The user will download normally, but under the hood the system will be doing a lot of work. It will check multiple "mirrors" of the download, by looking at the different available data nodes, then let the user choose the fastest or best option. File transfer will occur to bring the file to the user, then the system will be the one checking if the file were downloaded correctly. There is a simple organization of available files as well.

The system is described as \textit{community assisted}. Since ideally, institutions can contribute to sharing the data by hosting servers that will become the data nodes. This sharing allows a greater bandwidth, and hopefully speed in which this data can be transferred.

The database is a distributed database, which will aim to utilize the best features from Peer2Peer and Client/Servers. This will also aim to store sequences. Thus the "Distributed Database for Sequences" portion of the title.

\subsubsection{Community Aspect}
- community: define as set of people who generally collaborate together to do research
- focus on genome research community
- any users in a community can upload data
- users can become data nodes
- as data nodes increase, data speeds (bandwidth) should theoretically increase as well
- system is an opportunity to set up genomic databases for frequently used local data

- summary: community assists in adding data
- community assists in adding data storage (through data nodes)
- community assists in increasing the data speeds
- community is assisted since the data it uses frequently should theoretically get accessed faster

- addition: this is a framework to design and develop community-centered databases (can be modeled to be used beyond genome)

% CID PROGRESS ENDS HEREE FOR JUNE 4 2020

\begin{figure}[h]
\caption{App network diagram}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis1.png} 
\end{figure}

\begin{figure}[h]
\caption{POST: Uploading a sequence to the master node}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis3.png} 
\end{figure}

\begin{figure}[h]
\caption{POST: Uploading a sequence to the data node}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis2.png} 
\end{figure}

\begin{figure}[h]
\caption{GET: Downloading a sequence}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis4.png} 
\end{figure}

%CID: Explain the data model and data permissions

\begin{table}[h]
\caption{Data Model to be Used in the System}
\label{table:data_model_table}
\begin{tabular}{cl}
    \toprule
    key & description \\
    \midrule
    seq\textunderscore id & unique ID assigned to all sequences \\
    organism & organism where the sequence was extracted from \\
    quality & the rating of the data based on the number of downloads or another criteria \\
    uploader & user that uploaded the data \\
    institute & institute where the uploader or sequence was taken from \\
    upload\textunderscore date & date the data was uploaded to the database \\
    last\textunderscore modified & the date the metadata or data was edited \\
    data\textunderscore nodes & list of data nodes that this data is contained to \\
    \bottomrule
\end{tabular}
\end{table}

Table~\ref{table:data_model_table} on Page~\pageref{table:data_model_table} aims to show the metadata that will be initially stored in the database.

%CID: Data Permissions

\begin{table}[h]
\caption{Data Permissions and Users in the System}
\label{table:data_perm_table}
\begin{tabular}{cl}
    \toprule
    user & description \\
    \midrule
   Moderator & Has ability to upload, download, store data, verify \\
    Peer & Has ability to upload, download, store data \\
    Guest & Has ability to upload, download \\
   \bottomrule 
\end{tabular}
\end{table}

Table~\ref{table:data_perm_table} on Page~\pageref{table:data_perm_table} shows the different set of users and the permissions allowed to them in the system. The moderator is a special user who has the ability to not allow some data to be uploaded. This is to make sure there is a means to verifying the data being uploaded. All users can upload, download, and store (or host) data. All other users, which are called guests has an ability to download and upload data.

%HIII NEED HELP HEREE
\section{Proof of Correctness}
correct if
GET - gets data
POST - puts data
PUT - edits data
DELTE - deletes data

speed should be X

A proof of correctness for the implementation.

\section{Theoretical Analysis of Speed}

Theoretical analysis of the speed up by checking time complexity of the base method versus the distributed method.

in the base method
client sends data in X bytes per second. When X people are using. 
(insert math equation)
Each person gets the file at 1 byte per second.

in the proposed method
client has N data nodes and each node sends data in X bytes per second. When X people are using. Assuming perfect and equal distribution. 
(insert math equation)
Each person gets  at least N bytes per second

Since N is always greater than 1, the speed should always be greater in the proposed method

As user number increases, base method will reach a limit to its ability to handle data transfer.

As user number increases, proposed method will also reach a limit. But scalability is possibile to remove the limit.

\section{Comparison of base method vs distributed method}

base method (NCBI)
- houses a lot of biological data
- go to website
- one website
- one data source
- single point of failure
- implemented

proposed method
- ability to scale
- multiple (scalable data nodes) data sources
- multiple (scalable data nodes) data links
- not yet implemented
- can be done / implemented on a local scale (limit nodes to the needs)
- only have data for a certain data community
- only have users from a certain data community
- able to design the APIs for speed, security, consistency, tolerance, etc. depending on community needs

- increase in data speed transfer (if there are sufficient available data nodes)
- can reduce the data stored to house only locally used and needed data

\section{Recommendations}

Implement this. Describe the technologies used and why. And make the code public. Better yet create a template code to easily copy and implement own data centers.

Implementation problems. The network API will face problems in port access especially in UPD internet. Only 2 ports are currently available.

Have a set of testing APIs and have testing APIs embedded in each component. And the testing consistently checks which data have integrity.

Have security. The API structure is one way of having security in the future even if it isnt implemented yet. Restructure and redesigning the APIs are one way to achieve security. Adding user authentication is one way to achieve this. As well as having a separate layer between user and master. Where all interactions happen in that layer (even data transfer??). 
Abstract this project even further. Creating means for master nodes to communicate and collaborate with other master nodes. This will let different caDDS (assuming this is implemented) to communicate and collaborate (send data) to other caDDS.

Abstract this project for non-genomic data as well. To develop the idea of a distributed data community (where distributed databases are easy to scale, implement, and connect). Which can help share many kinds of scientific data more easily. A specific suggestion is for other fields with large data gathered, such as space image data.

\section{Slides}
1. Title
2. Abstract
3. Overview

CURRENT
BIO
1. Genes
- all organisms have genes
- genes from DNA: AGCT
- genes are instructions and blueprints for life

2. Genomics
- study of genes
- more complex organisms have more complex genes
- studying complex organisms needs a lot of storage

3. Sequencing
- reading genes from cells is sequencing
- replications(coverage) \& fragmentation of DNA to read
- store the many fragments

4. Human Genome Project
- finished in 2003, project wanted to read the entire human genome
- didn't have the tools at the time
- big cost, many institutions, 10 years

5. Next Generation Sequencing
- new technology to read sequences faster
- reading faster -> more reads -> more data

6. Growing problem of storage
- moore's law make faster processing power and better storage
- but moore's law is too slow for genomic data increase
- each has 60x coverage, around 200gb each
- NGS, More genomic researches, more files needed to be saved and transferred

TECH
7. Database
- store data with tech

8. Definition
- DBMS chuchu

9. Database Principles
- consistency
- availability
- partition tolerance

10. Database Types
- Centralized
- Distributed
- use Distributed becuz x

11. Application Architecture
- Client Server
- P2P
- Hybrid
- Use Hybrid because X

RRL
12.SeqTorr
- idea
- pros
- cons

13.BioTorrents
- idea
- pros
- cons

14.PeerDB
- idea
- pros
- cons

15.NCBI
- idea
- pros
- cons

16. PROBLEM
-classical databases
-distributed databases

17. OBJECTIVES
- make a dist database
- community assisted
- main APIs
- time complexity
- pros and cons

18. SCOPE
- no this and that

SYSTEM
19. Network
- five levels, but needs Level 4 and 5 only
20. Summary
- system is this
- distributed
- community assisted
- genomic research

21. User-System
- user talks to system (via master)

22. Master-Data
- master is the main interface
- data is what stores info

23. Internal Node Interface
- frontend talks to backend
- backend talks to other backend via APIs

24. Internal Node Structure
- front end
- back end
- database

COMMUNITY
25. Summary
- good for community
- easy to expand storage

COMPARISONS
26. Speed Comparisons
- simple time complexity chart
- faster si caDDS ofc

27. Pros and Cons Comparisons
-caDDS is scalable, easy for community, has docu hehe
-NCBI is one, very specific features,

28. Recommendations
-implement
-security
-public code
-stuff

% CID PROGRESS FOR JUNE 4 HAS A LOT OF ROUGH IDEAS
% CID PROGRESS ENDS HEREE FOR JUNE 4 2020

%\begin{figure}[h]
%  \centering
%  \includegraphics[width=\linewidth]{sample-franklin}
%  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
%    Ewing, Inc. [Public domain], via Wikimedia
%    Commons. (\url{https://goo.gl/VLCRBB}).}
%  \Description{A woman and a girl in white dresses sit in an open car.}
%\end{figure}

\section{Appendices}

%%%%%%%%%%%%%%%%%%%%%%%%%%%% ACKNOWLEDGEMENT & GRATITUDE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}

(Not yet serious) Our advisers have really helped us over the school year. The dificult pandemic has changed our thesis to the point it has been - Cid
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibfile}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{User-System Interface}
(check appendix for detailed pseudocode)

parang this section abstracts the user part from the system part

The user commmunicates with the system. User can do create, read, update, delete.

The system on its own delegates how things are stored or updated consistently ...

The user communicates with the master all the time???

\section{Master Node-Data Node Interface}
(check appendix for detailed pseudocode)

parang this section abstracts how master and data communicate and store data

GET
The user requests data from the system (via the master gate hehe)

The master holds a copy of all the data locations in the data nodes (can be called a directory). 
The master once needs a certain data, will check its directory which data nodes have it
The master once knows which of its data nodes have the data (multiple) will call them to inform it will request data
The data is informed that data will be requests
The master sends the directory of addresses of data nodes with the data and transfer key???? to the user

The user user receives the transfer key and uses it in the website to request data from data nodes

CREATE
The user wants to send data to the system (via the master gate hehe)

The master holds a copy of all the data locations (can be called a directory). 
The master will find 2+ data nodes to hold the data, and store them in a temp directory
The master will call the 2+ data nodes and check if they have enough space and are able to receive the data
The master will wait for an ack before sending the data to the 2+ data nodes
The master will ask the user to generate a hashcode
The master will send the user data nodes to send the data
The data nodes will receive the data and hash
The data nodes will complete receiving data, generate a hash and compare it with the user's hash. If different, repeat the process (may ask from a data node with complete data).
The data nodes if all have the complete, correct data will inform master node.

The user user receives the transfer key 
The user generates a hash of its data
The user uses the transfer key to send data to data nodes

\section{Master Node-Data Node Structure}
(check appendix for detailed pseudocode)

MASTER NODE
holds data:
directory of data nodes with verified data
directory of data nodes with unverified data (being transferred currently)
directory of data nodes space free

holds function:
which data nodes have data x?
does data node N have data x?
is user able to receive data X?
is user able to send data X?
is data node N able to send data X?
is data node N able to receive data X?

what is hash of data X on data node N?
what is hash of user data X to be sent?
Is hash of data X on data node N and user data X same?

is data X being transferred currently from data node N?
is data X being transferred currently to data node N?
is data X done transferring to data node N?
is data X across all data nodes consistent? (same hash)?

how much space does data node N have?
how much space does user data to send need?

send data node N location to user?
where is the location of data node N?

DATA NODES
holds data:
Metadata
Data
Directory of Metadata
Storage Space

holds function:
has storage space X?
how much storage space?
is data X across all data nodes consistent? (same hash)?
how much space does user data need?
able to receive data X?
able to send data X?
what is hash of data X?
what is hash of user data X to be sent?
Is hash of data X and user data X same?

\section{Node Internal Interface}
(check appendix for detailed pseudocode)
Frontend-Backend Interface
Via Data APIs
GET, POST, PUT, DELETE

Via Data Sync APIs
DATA SAME?

Via Data Availability APIs
DATA AVAILABLE?
DATA NODE ABLE TO RECEIVE?
DATA NODE ABLE TO SEND?

\section{Node Internal Structure}
(check appendix for detailed pseudocode)

parang this section abstracts how nodes function inside

Frontend - debugging mostly, users dont see this frontend???? or isang frontend lang ni master nakikita
Backend - storage before it goes to database? for functions? for APIs??

Inside Backend: Idea pala to categorize functions into these APIs idk if tama paggamit ko ng APIs
Network APIs - abstracted part to connect to the network
Database Sync APIs - abstracted part to keep database consistent
Database Availability APIs - abstracted part to keep database available

Database - stores shit

\subsection{Community Aspect}
- good for community since the system is expandable or customizable??
- community can see the whole framework and structure
- communities can set up their own master and data nodes
- individuals in communities can host their own data nodes (ideally master node is the institution with the data)
- ideally having multiple data nodes makes it download faster somehow
- communities can only store the data they need
- communities basically can self-expand storage, self-organize somehow the storage, and self-increase their bandwidth

\end{document}
\endinput
%%
%% End of file `sample-acmsmall.tex'.
%%

### 2020-06-18 Paper2020ACM/Presentation.tex
Initial Overleaf Import

--- 2020-06-18 11-25/Presentation.tex
@@
 %\usetheme{default}
-%\usetheme{AnnArbor}
+\usetheme{AnnArbor}
 %\usetheme{Antibes}
@@
 %\usetheme{Berlin}
-\usetheme{Boadilla}
 %\usetheme{CambridgeUS}
@@
 %\usecolortheme{whale}
+
+%\usetheme{Boadilla}
 \usecolortheme{wolverine}
@@
 
-\title[Short title]{Thesis Pre-final presentation} % The short title appears at the bottom of every slide, the full title is only on the title page
+\title[caDDS]{caDDS: Community Assisted Distributed Database for Sequences}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%----------------------------------------------------------------------------------------
+%	PRESENTATION SLIDES
+%----------------------------------------------------------------------------------------
+
 \begin{document}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OVERVIEW %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}
@@
 
-%----------------------------------------------------------------------------------------
-%	PRESENTATION SLIDES
-%----------------------------------------------------------------------------------------
-
-%------------------------------------------------
-\section{Introduction}
-
-% to do sa intro
-% list all jargon
-% explain most jargon
-% iklian bio part
-% pahabain tech part esp p2p
-% explain CDN
-
-    %%%%%%%%%%%%%%%%%%% BIO SLIDES %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-    \subsection{Definition} %keep daw
-    \begin{frame}{Human Karyotype}
-        \centering
-\includegraphics[scale=1]{karyotype.jpg}
-    \end{frame}
-    \begin{frame}{Definition of Genomics}
-        According to the National Institute of Health \textit{Specifically for the Human Genome Project}
-                \begin{block} {\textbf{Definition of Genomics}}
-        Genomics is the study of all of a person's genes (the genome), including interactions of those genes with each other and with the person's environment. \cite{genomics-definition}
-                \end{block}
-    The genomics definition we will use is for \textbf{all} organisms
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO + RRL %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{Abstract}
 \end{frame}
 
-    \subsection{Introduction}
-    \begin{frame}{Introduction}
-        \begin{itemize}
-            \item DNA comprises of the components A,T,G,C
-            \begin{itemize}
-                \item \textbf{A} - Adenine
-                \item \textbf{T} - Thymine
-                \item \textbf{G} - Guanine
-                \item \textbf{C} - Cytosine
-            \end{itemize}
-            \item There needs to be a lot of these components to describe complex organisms
-            \begin{itemize}
-                \item The full human genome contains around $3.2 x 10^9$ base pairs \cite[p.~4]{introgenomics}
-            \end{itemize}
-        \end{itemize}
-        \centering
-        \includegraphics[scale=0.1]{gene-base-pair.png}
+\begin{frame}{Genes}
 \end{frame}
 
-    \subsection{Genomics Data}
-    \begin{frame}{Genomics Data}
-            \textbf{Data Type}
-    \begin{itemize}
-        \item Store DNA sequences digitally
-        \item A simple data file format is called FASTA format
-            \begin{itemize}
-                \item starts with $>$
-                \item then can add keywords and attributes separated by $|$
-            \end{itemize}
-    \end{itemize}
-    \begin{block}{Sample.FASTA}
-    $>$AB000263 $|$acc=AB000263$|$descr=Homo sapiens mRNA for prepro cortistatin like peptide, complete cds.$|$len=70
-ACAAGATGCCATTGTCCCCCGGCCTCCTGCTGCTGCTGCTCTCCGGGGCCACGGCCACCGCTGCCCTGCC
-
-    \end{block}
+\begin{frame}{Genomics}
 \end{frame}
 
-    \subsection{Sequencing}
 \begin{frame}{Sequencing}
-        \textbf{Sequencing}
-    \begin{itemize}
-        \item Sequencers are machines used to look at the DNA of a specimen (A,G,C,T)
-        \begin{itemize}
-            \item Sequencers look at the exact order of a single strand \cite{genomics-definition}. 
-        \end{itemize}
-        \item DNA is \textbf{replicated} and fragmented before processing
-        \item Sequencers then reads (or sequences) them
-        \item Output of the read fragments is contained in a .FASTA file
-    \end{itemize}
-            \centering
-        \includegraphics[scale=0.2]{sequencing.png}
 \end{frame}
@@
 \begin{frame}{Next Generation Sequencing}
-    (2003) The first human genome took ten years of work and $3,000,000,000$ \$ to finish.
-    \begin{itemize}
-        \item (2019) Currently the technology today in Japan can make 10,000 human genomes per day (Timewise: 36.5 million times faster) \cite[p.~19]{introgenomics}
-    \end{itemize}
-    \textbf{Next Generation Sequencers}
-        \begin{itemize}
-        \item Can read more faster, and cheaper
-        \item More data
-        \item Technology will get faster, there needs to be systems in preparation for that
-    \end{itemize}
-    \end{frame}
-    %%%%%%%%%%%%%%%%%%% TECH SLIDES %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-    \subsection{Data Storage}
-    \begin{frame}{Data Storage}
-        A \textbf{database} (\textbf{DBMS}, Database Management System) is defined by 
-        \begin{block} {\textbf{Definition}}
-
-        DBMS contains information about a particular enterprise
-        \begin{itemize}
-            \item Collection of interrelated data
-            \item Set of programs to access the data 
-            \item An environment that is both convenient and efficient to use
-        \end{itemize}
-        \cite{Silberschatz2010}
-        \end{block}
 \end{frame}
 
-    \subsection{Database Principles}
-    \begin{frame}{Database Principles}
-        \begin{itemize}
-            \item \textbf{Principles} to remember when creating a database (CAP). Idea is that only 2 of the three can be focused on. \cite[Ch.~19]{Silberschatz2010}
-            \begin{itemize}
-                \item \textbf{Consistency}
-                \begin{itemize}
-                    \item Update one part, will all other (replicated) parts be updated quickly as well
-                    \item Data is accurate and copied faithfully
-                \end{itemize}
-                \item \textbf{Accessibility}
-                \begin{itemize}
-                    \item How easy it is to access data
-                    \item API protocols
-                \end{itemize}
-                \item  \textbf{Partition Tolerance}
-                \begin{itemize}
-                    \item If one portion breaks, the other portions are still usable
-                    \centering 
-                    \includegraphics[scale=0.2]{partition_tolerance.png}
-                    \end{itemize} 
-             \end{itemize}
-        \end{itemize}
+\begin{frame}{Growing problem of storage}
 \end{frame}
 
-    \subsection{Special Database}
-    \begin{frame}{Special Databases}
-        \textbf{Large Data}
-        \begin{itemize}
-            \item \textbf{Distributed Database}
-            \begin{itemize}
-                \item A distributed database system consists of loosely coupled sites that share no physical component             
-                \item Database systems that run on each site are independent of each other
-                \item Transactions may access data at one or more sites
-                \cite[Ch.~19]{Silberschatz2010}
-                \item Beneficial for large data with a certain speed requirement
-            \end{itemize}
-        \end{itemize}
+\begin{frame}{Database}
 \end{frame}
 
-    \begin{frame}{Special Databases}
-        \textbf{Genomics Data}
-        \texttt{needs a different set of features to easily conduct processes (alignment, assembly, comparison).}
-        \begin{itemize}
-            \item Need a special database to store all of these data
-            \item Designed for more efficient and optimal use
-            \item Has special needs for the users of genomic data
-            \item Discuss in the RRL the different ways researchers approached this
-        \end{itemize}
+\begin{frame}{Database Definition}
 \end{frame}
 
-        \begin{frame}{Network Layers}
-         \textbf{Networks Layers} how computers pass data to each other
-        \begin{itemize}
-            \item \begin{itemize}
-            \item Level 1: Medium used to send signals (physical, electrical signals of 0 and 1), through wire or by air
-            \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together different . Goal of sending a group of signals.
-            \item Level 3: Introduction of networks: IP Address to uniquely identify entities. 
-            \item Level 4: Management of error free data transmission: Sending a group of messages via TCP, UDP, FTP
-            \item \textbf{Level 5}: Abstraction of Level 4: Function calls programmers will use to do level 4 processes.
-            \end{itemize}
-        \end{itemize}
+\begin{frame}{Database Principles}
 \end{frame}
 
-    \subsection{Application Architecture}
-    \begin{frame}{Application Architecture}
-        \begin{itemize}
-            \item The \textbf{application architecture} is how a networked app is structured across the various end systems (i.e. computers, servers, etc.) \cite{kurose}
-            \item Falls into either client-server, peer-to-peer, or hybrid architecture
-        \end{itemize}
-    \end{frame}
-        \subsubsection{Client-Server}
-        \begin{frame}{Client-Server}
-            \begin{itemize}
-                \item A \textbf{client-server} architecture relies on a Web server that is always available and has a fixed IP address\\
-                \item A single server connects to many clients
-            \end{itemize}
-            \centering
-            \includegraphics[scale=0.5]{Server-based-network.png}
-        \end{frame}
-        \subsubsection{Peer-to-Peer \& Hybrid}
-        \begin{frame}{Peer-to-Peer \& Hybrid}
-            \begin{itemize}
-                \item A \textbf{peer-to-peer (P2P)} architecture has minimal to no reliance on always-on, dedicated servers
-                \item Peers are desktops/laptops that communicate to each other without passing through a dedicated server
-                \item A \textbf{hybrid} architecture combines elements of P2P and client-server architectures, attempting to utilize advantages of both
-            \end{itemize}
-            \centering
-            \includegraphics[scale=0.5]{P2P-network.png}
-        \end{frame}
-        \begin{frame}{BitTorrent terminology}
-            \begin{itemize} % siguro breeze through what these mean 
-                \item \textbf{Peers} are users that connect to the BitTorrent network.
-                \item \textbf{Torrent files} contain metadata about files \& folders, trackers, and hash values.
-                \item \textbf{Trackers} are servers that keep track of peers and availability of files
-                \item \textbf{Seeds} are peers which have downloaded the entire file and are making it available
-            \end{itemize}
+\begin{frame}{Database Types}
 \end{frame}
 
-\section{Review of Related Literature}
-    \subsection{BioTorrents}
-    \begin{frame}{BioTorrents}
-        % outline
-        \begin{itemize}
-        \item System for legally sharing scientific data that works on top of the BitTorrent protocol. It is much like a public tracker.
-        \item The main website hosts .torrent and metadata files
-        \item Info about each dataset includes categories, license, filenames, etc. that helps users in searching for relevant datasets
-        \item Torrent file has list of trackers, servers that "know" which peers are serving which files.
-        \item Torrent client downloads from multiple peers simultaneously.
-        \item To upload a torrent, the client makes a .torrent file using the torrenting software, then uploads it to the main website along with metadata.
-The client must have their computer continuously turned on to keep the file available as a peer.
-\cite{biotorrents}
-        \end{itemize}
+\begin{frame}{Application Architecture}
+\end{frame}
+
+\begin{frame}{SeqTorr}
 \end{frame}
@@
 \begin{frame}{BioTorrents}
-        \begin{itemize}
-            \item As with all public trackers, BioTorrents requires "good souls" that will keep many torrents available as often as possible.
-            \item The datasets must be moderated to prevent abuse.
-            \item The BioTorrents website is no longer up.
-        \end{itemize}
 \end{frame}
 
-    \subsection{PeerDB}
 \begin{frame}{PeerDB}
-        \begin{itemize}
-            \item Proposed P2P system for sharing general data
-            \item Each node is equipped with a database
-            \item Queries are passed around neighbors, with a maximum number of hops
-            \item Caches information about other peers to speed up queries
-            \item Large number of nodes & relatively small number of lookup servers
-            \cite{peerdb}
-        \end{itemize}
 \end{frame}
 
-    \subsection{SeqTorr}
-    \begin{frame}{SeqTorr}
-    \textbf{SeqTorr} \cite{seqtorr}
-    \begin{itemize}
-        \item Utilizes distributed database
-        \item Master node
-        \begin{itemize}
-            \item containing metadata of all the sequences
-            \item where user authentication is handled
-        \end{itemize}
-        \item Data node
-        \begin{itemize}
-            \item contain a portion of all the sequences
-            \item needs to replicate a sequence \texttt{r} times before a sequence is considered available
-            \item proximity of data nodes can increase download speed but isn't true P2P
-        \end{itemize}
-        \item Defined API Implementation
-        \begin{itemize}
-            \item POST
-            \item GET
-            \item PUT
-            \item DELETE
-        \end{itemize}
-
-    \end{itemize}
+\begin{frame}{base system: NCBI}
 \end{frame}
 
-    \begin{frame}{SeqTorr}
-        Problems:
-        \begin{itemize}
-            \item program not currently available
-            \begin{itemize}
-                \item cannot measure any statistics on this
-            \end{itemize}
-            \item not Peer2Peer
-        \end{itemize}
+\begin{frame}{Problem}
 \end{frame}
 
-    \begin{frame}{Table}
-   \begin{table}[]
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ \hline
-BioTorrents & P2P           & No         & Depends                    & Bio-related  \\ \hline
-SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ \hline
-PeerDB      & P2P           & No         & Depends                    & General      \\ \hline
-NCBI        & CS            & Yes        & Slow (US server)           & Bio-related  \\ \hline
-\end{tabular}
-\end{table} 
+\begin{frame}{Objectives}
 \end{frame}
 
-% NCBI Biosystems
-% pros
-% cons
-
-% BitTorrious
-% pros
-% \cite{bittorious}
-% cons
-
-\section{Problem}
-\begin{frame}{Problem}
-The different gaps or problems found with other implementations
-
-\begin{itemize}
-    \item Genome researchers download entire database of genomic data
-    \item Genome researchers only need a portion of the database, but end up downloading all
-    \item Client-Server approach can face bottlenecks (server speed)
-    \item Code has no documentation, thus harder to maintain
-%% add more
-
-\end{itemize}
+\begin{frame}{Scope}
 \end{frame}
 
-\section{Objectives}
-\begin{frame}{Objectives}
-Based on the problems we found earlier, we propose the following objectives
-    \begin{itemize}
-        \item Create information system that speeds up the download and upload of genomic data by at least 3x compared to existing systems
-        \item The system should have the main API's
-        \begin{itemize}
-            \item POST genome data
-            \item GET genome data (implement hybrid P2P)
-            \item PUT genome metadata (optional: PUT data)
-            \item DELETE genome data
-        \end{itemize}
-        \item Write documentation for the system
-    \end{itemize}
+\begin{frame}{System}
 \end{frame}
 
-\section{Scope}
-\begin{frame}{Scope}
-The current scope and coverage of the project is only
-\begin{itemize}
-    \item Security and Authentication is not part of the scope
-    \item Use only FASTA file format 
-    \begin{itemize}
-        \item DNA nucleotides only
-    \end{itemize}
-    \item UI will be purely for functional purposes (very basic design)
-\end{itemize}
+\begin{frame}{Network}
 \end{frame}
 
-\section{Theoretical Framework}
-
-\begin{frame}{Theoretical Framework: System Description}
-    \begin{itemize}
-        \item A \textbf{master node} stores metadata of all sequences, may or may not contain sequence files
-        \item \textbf{Data nodes} stores sequence files & metadata of all sequence files in its filesystem
-        \item Both master node & data nodes have user interface that allow users to upload, search, and download sequence files
-    \end{itemize}
-\end{frame}
-% DRAW.IO FILE: https://drive.google.com/file/d/1Nbp0Kmgcr2BZ13LBT3SGLMBoA6JA_Eth/view
-\begin{frame}{Theoretical Framework: Schematics}
-\textbf{App network diagram} \\ \medskip
-\centering
-\includegraphics[scale=0.5]{thesis1.png}
+\begin{frame}{Summary}
 \end{frame}
 
-\begin{frame}{Theoretical Framework: Schematics}
-\textbf{POST: Uploading a sequence to the master node} \\ \medskip
-\centering
-\includegraphics[scale=0.5]{thesis3.png}
+\begin{frame}{Community}
 \end{frame}
 
-\begin{frame}{Theoretical Framework: Schematics}
-\textbf{POST: Uploading a sequence to a data node} \\ \medskip
-\centering
-\includegraphics[scale=0.5]{thesis2.png}
+\begin{frame}{Speed Comparisons}
 \end{frame}
 
-\begin{frame}{Theoretical Framework: Schematics}
-\textbf{GET: Downloading a sequence} \\ \medskip
-\centering
-\includegraphics[scale=0.5]{thesis4.png}
+\begin{frame}{Comparisons}
 \end{frame}
 
-\section{Proposed Methodology}
-\begin{frame}{Proposed Methodology}
-\begin{enumerate}
-    \item Creating the system
-    \begin{enumerate}
-        \item Make prototype for UI
-        \item Code data node
-        \item Deploy \& test functionality of data node/s
-        \item Code master node
-        \item Deploy \& test functionality of master node/s
-        \item Test interoperability of nodes
-    \end{enumerate}
-    \item Benchmarking
-    \begin{enumerate}
-        \item Create test for testing original speed
-        \item Test original download speeds
-        \item Create test for testing new speed
-        \item Test new speeds
-    \end{enumerate}
-        \item Documentation
-    \begin{enumerate}
-        \item PDF & Github Style
-        \item How to use (all users)
-        \item How to set up
-        \item How to maintain & add features
-        \item How to test?
-    \end{enumerate}
-\end{enumerate}
-\end{frame}
-
-\subsection{Proposed Methodology Timeline}
-\begin{frame}{Proposed Methodology: Timeline 1}
-    \begin{itemize}
-        \item Mid January: Start
-        \item Mid April: End
-    \end{itemize}
-    \begin{enumerate}
-        \item End January Deliverable: 
-        \begin{itemize}
-            \item Wireframe Mockup
-            \item Pseudocode / Flowchart of Entire System (to be reviewed by the adviser)
-            \item Finalized feature list
-            \item Documentation Version 1 (Cid)
-        \end{itemize}
-        \item End of February Deliverable:
-            \begin{itemize}
-                \item Benchmarking specifications finalization \& Set up
-                \item Start of bench marking of existing tech
-                \item Start Coding
-                \item Documentation Version 2 (Cid)
-            \end{itemize}
-    \end{enumerate}            
-            \end{frame}
-
-    \begin{frame}{Proposed Methodology: Timeline 2}
-        \begin{enumerate}  
-        \item Mid March Deliverable:
-            \begin{itemize}
-                \item Finish Coding version 1 of system
-                \item Implement basic UI
-                \item Finish bench marking existing 
-                \item Start Alpha Testing (us and other undergrad thesis students)
-                \item Documentation Version 3 (Cid)
-            \end{itemize}
-        \item End of March Deliverable:
-            \begin{itemize}
-                \item Revise version 1 to version 2 of system based on Alpha Test
-                \item Revise basic UI
-                \item Testing of system for feature 
-                \item Start Beta Testing (public, DCS \& PGC community)
-                \item Documentation Version 4 (Cid)
-            \end{itemize}
-        \item Mid April Deliverable
-            \begin{itemize}
-                \item Revise version 2 to version 3 of system based on Beta test
-                \item Documentation Version 5 (Cid)
-            \end{itemize}
-    \end{enumerate}
+\begin{frame}{Recommendations}
 \end{frame}

### 2020-06-18 Paper2020ACM/meeting 1-21-2020.tex
Initial Overleaf Import

199-1-21 Meet Notes
Data sync bet servers

when a dataset gets uploaded:
- notify master node of new update
- master picks which other data nodes will host a copy of the dataset
- then asks those nodes to copy from the original data node
- data nodes will copy from the original
- after complete download, flag locally and notify master node about completion

upd or delete:
- inform everyone if metadata has been changed

only master node keeps a record of all metadata in the dist system
all data nodes have to ask master node if user wants a file it does not own

use nginx
http://w.x.y.z:80/api/ (exposed endpoint) -> http://w.x.y.z:5984 (couchdb)

assume data nodes have a public ip
assume all data is public

dev:
start coding NOW, then decide what features to change/add/remove

benchmarking:
refer to lit on factors to consider for benchmarking
look for fasta files

### 2020-06-19 Paper2020ACM/Presentation.tex
Edited presentation

@@
 \begin{frame}{Abstract}
+  Genomic data the past few years have grown exponentially with the technology barely keeping up.
+The problem faced by genome researchers is the large data set and difficulty transferring the files.
+Previous distributed databases are either not meant for genomic data, or difficult to replicate.
+This paper lays the groundwork for a distributed database that is designed to easily scale to accommodate bigger storage, and also reduces the data speed bottleneck faced by centralized databases.
+The proposed system has a master node and data nodes.
+Data nodes should theoretically increase data transfer speeds even as the number of users increase.
+(to be updated based on conclusion) While current databases(e.g.
+NCBI) have more data, more users, and a live platform.
+The proposed system is aimed towards a smaller community, with more frequent and localized data, which the data storage can easily be expanded as the need arises.
+A recommendation is to implement the system and make the code public for the use of other communities.
 \end{frame}

### 2020-06-19 Paper2020ACM/Paper2020.tex
modified presentation.tex and removed it from the main paper

@@
 
-\section{Slides}
-1. Title
-2. Abstract
-3. Overview
-
-CURRENT
-BIO
-1. Genes
-- all organisms have genes
-- genes from DNA: AGCT
-- genes are instructions and blueprints for life
-
-2. Genomics
-- study of genes
-- more complex organisms have more complex genes
-- studying complex organisms needs a lot of storage
-
-3. Sequencing
-- reading genes from cells is sequencing
-- replications(coverage) \& fragmentation of DNA to read
-- store the many fragments
-
-4. Human Genome Project
-- finished in 2003, project wanted to read the entire human genome
-- didn't have the tools at the time
-- big cost, many institutions, 10 years
-
-5. Next Generation Sequencing
-- new technology to read sequences faster
-- reading faster -> more reads -> more data
-
-6. Growing problem of storage
-- moore's law make faster processing power and better storage
-- but moore's law is too slow for genomic data increase
-- each has 60x coverage, around 200gb each
-- NGS, More genomic researches, more files needed to be saved and transferred
-
-TECH
-7. Database
-- store data with tech
-
-8. Definition
-- DBMS chuchu
-
-9. Database Principles
-- consistency
-- availability
-- partition tolerance
-
-10. Database Types
-- Centralized
-- Distributed
-- use Distributed becuz x
-
-11. Application Architecture
-- Client Server
-- P2P
-- Hybrid
-- Use Hybrid because X
-
-RRL
-12.SeqTorr
-- idea
-- pros
-- cons
-
-13.BioTorrents
-- idea
-- pros
-- cons
-
-14.PeerDB
-- idea
-- pros
-- cons
-
-15.NCBI
-- idea
-- pros
-- cons
-
-16. PROBLEM
--classical databases
--distributed databases
-
-17. OBJECTIVES
-- make a dist database
-- community assisted
-- main APIs
-- time complexity
-- pros and cons
-
-18. SCOPE
-- no this and that
-
-SYSTEM
-19. Network
-- five levels, but needs Level 4 and 5 only
-20. Summary
-- system is this
-- distributed
-- community assisted
-- genomic research
-
-21. User-System
-- user talks to system (via master)
-
-22. Master-Data
-- master is the main interface
-- data is what stores info
-
-23. Internal Node Interface
-- frontend talks to backend
-- backend talks to other backend via APIs
-
-24. Internal Node Structure
-- front end
-- back end
-- database
-
-COMMUNITY
-25. Summary
-- good for community
-- easy to expand storage
-
-COMPARISONS
-26. Speed Comparisons
-- simple time complexity chart
-- faster si caDDS ofc
-
-27. Pros and Cons Comparisons
--caDDS is scalable, easy for community, has docu hehe
--NCBI is one, very specific features,
-
-28. Recommendations
--implement
--security
--public code
--stuff
-
-% CID PROGRESS FOR JUNE 4 HAS A LOT OF ROUGH IDEAS
-% CID PROGRESS ENDS HEREE FOR JUNE 4 2020
-
 %\begin{figure}[h]

### 2020-06-19 Paper2020ACM/Presentation.tex
modified presentation.tex and removed it from the main paper

@@
 \end{frame}
-
+\section{Introduction}
+\subsection{Genomic Concepts}
 \begin{frame}{Genes}
+  \begin{itemize}   
+    \item all organisms have genes
+    \item genes from DNA: AGCT
+    \item genes are instructions and blueprints for life
+\end{itemize}
 \end{frame}
@@
 \begin{frame}{Genomics}
+  \begin{block} {\textbf{Definition of Genomics}}
+    Genomics is the study of all of a person's genes (the genome), including interactions of those genes with each other and with the person's environment. \cite{genomics-definition}
+  \end{block}
+  \begin{itemize}   
+    \item study of genes
+    \item more complex organisms have more complex genes
+    \item studying complex organisms needs a lot of storage
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Sequencing}
+  \begin{itemize}   
+    \item reading genes from cells is sequencing
+    \item replications(coverage) \& fragmentation of DNA to read
+    \item store the many fragments
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Human Genome Project}
+  \begin{itemize}   
+    \item finished in 2003, project wanted to read the entire human genome
+    \item didn't have the tools at the time
+    \item big cost, many institutions, 10 years
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Next Generation Sequencing}
+  \begin{itemize}   
+    \item new technology to read sequences faster
+    \item reading faster -> more reads -> more data
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Growing problem of storage}
+  \begin{itemize}   
+    \item moore's law make faster processing power and better storage
+    \item but moore's law is too slow for genomic data increase
+    \item each has 60x coverage, around 200gb each
+    \item NGS, More genomic researches, more files needed to be saved and transferred
+  \end{itemize}
 \end{frame}
 
+\subsection{Database Concepts}
 \begin{frame}{Database}
+  \begin{itemize}   
+    \item store data with tech
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Database Definition}
+  \begin{itemize}   
+    \item DBMS chuchu
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Database Principles}
+  \begin{itemize}   
+    \item consistency
+    \item availability
+    \item partition tolerance
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Database Types}
+  \begin{itemize}   
+    \item Centralized
+    \item Distributed
+    \item use Distributed becuz x
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Application Architecture}
+  \begin{itemize}   
+    \item Client Server
+    \item P2P
+    \item Hybrid
+    \item Use Hybrid because X
+  \end{itemize}
 \end{frame}
 
+\section{Review of Related technology}
+
 \begin{frame}{SeqTorr}
+  \begin{itemize}   
+    \item idea
+    \item pros
+    \item cons
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{BioTorrents}
+  \begin{itemize}   
+    \item idea
+    \item pros
+    \item cons
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{PeerDB}
+  \begin{itemize}   
+    \item idea
+    \item pros
+    \item cons
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{base system: NCBI}
+  \begin{itemize}   
+    \item why popular
+    \item idea
+    \item idea
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{Problem}
+\begin{frame}{base system: NCBI}
+  \begin{itemize}   
+    \item pros
+    \item cons
+    \item cons
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{Objectives}
+\section{Problem}
+\begin{frame}{Problem}
+  \begin{itemize}   
+    \item classical databases
+    \item distributed databases
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{Scope}
+\section{Objectives}
+\begin{frame}{Objectives}
+  \begin{itemize}   
+    \item make a dist database
+    \item community assisted
+    \item main APIs
+    \item time complexity
+    \item pros and cons
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{System}
+\section{Scope}
+\begin{frame}{Scope}
+  \begin{itemize}   
+    \item no this
+    \item and
+    \item that
+  \end{itemize}
 \end{frame}
 
+\section{Theoretical Framework}
 \begin{frame}{Network}
+  \begin{itemize}   
+    \item five levels
+    \item need to focus on Level 4 and 5 only
+    \item 
+  \end{itemize}
+\end{frame}
+
+\subsection{System Summay}
+\begin{frame}{System}
+  \begin{itemize}   
+    \item system is this
+    \item distributed
+    \item community assisted
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Summary}
+  \begin{itemize}   
+    \item 
+    \item 
+    \item 
+  \end{itemize}
 \end{frame}
 
+\subsection{Community}
 \begin{frame}{Community}
+  \begin{itemize}   
+    \item good for community
+    \item easy to expand storage
+    \item 
+  \end{itemize}
 \end{frame}
 
+\subsection{Comparisons}
 \begin{frame}{Speed Comparisons}
+  \begin{itemize}   
+    \item simple time complexity chart
+    \item faster si caDDS ofc 
+    \item 
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{Comparisons}
+\begin{frame}{Final Comparisons}
+  \begin{itemize}   
+    \item caDDS is scalable, easy for community, has docu hehe
+    \item NCBI is one, stable, widely used
+    \item 
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Recommendations}
+  \begin{itemize}   
+    \item implement
+    \item security
+    \item public code
+  \end{itemize}
 \end{frame}

### 2020-06-19 Paper2020ACM/Paper2020.tex
Fixed grammar for presentation and cleaned up abstract

@@
 \begin{abstract}
-300 words
-- purpose and problems:
-    - design a framework
-    - design a database
-- basic design:
-- system summary:
-    - master node and data node
-- conclusion:
-    - time complexity wise: 
-    - proposed system has this pros
-    - based system has this pros
-
 working draft (insert kulang when info is ready):

### 2020-06-19 Paper2020ACM/Presentation.tex
Fixed grammar for presentation and cleaned up abstract

@@
     \item all organisms have genes
-    \item genes from DNA: AGCT
+    \item Genes are made of a certain molecular structure called DNA which correspond to information denoted by: A,G,C,T
     \item genes are instructions and blueprints for life
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Human Genes}
+  \begin{itemize}   
+    \item There is a lot of information in one human cell
+    \item One human cell's entire genes laid out is 2 meters in length \cite{ency_sci_tech} and has $3.2x10^9$ base pairs \cite{introgenomics}
+    \item Human body has a lot of information, that when laid out spans the earth to the moon six thousand times
 \end{itemize}
@@
   \begin{itemize}   
-    \item study of genes
-    \item more complex organisms have more complex genes
-    \item studying complex organisms needs a lot of storage
+    \item \textbf{Genomics}: study of whole sets of genes and their interactions \cite{campbell}
+    \item More complex organisms have more complex genes (many parts and interactions)
+    \item Studying these complex organisms needs a lot of storage
+    \item The current online portal mainly used to study this is NCBI \cite{campbell}
   \end{itemize}
@@
   \begin{itemize}   
-    \item reading genes from cells is sequencing
-    \item replications(coverage) \& fragmentation of DNA to read
-    \item store the many fragments
+    \item Reading genes from cells is sequencing
+    \item Replications (coverage) \& Fragmentation of DNA is necessary to read
+    \item Every read stores the many fragments
+    \item Each sequence read is redundant and uses a lot of space
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Human Genome Project Successes}
+  \begin{itemize}   
+    \item 2 years after, the project led to a whole genome project for chimpanzees (human's closest relative) \cite{campbell}
+    \item These led to many whole genome analysis of organisms (around 4,300 in 2013) \cite{campbell}
+    \item Growing amount of research and data from genome research.
   \end{itemize}
@@
   \begin{itemize}   
-    \item finished in 2003, project wanted to read the entire human genome
-    \item didn't have the tools at the time
-    \item big cost, many institutions, 10 years
+    \item Finished in 2003, project wanted to read the entire human genome
+    \item Didn't have the mature tools at the time
+    \item Big cost, many institutions, 10 years
   \end{itemize}
@@
 
-\subsection{System Summay}
+\subsection{System Summary}
 \begin{frame}{System}

### 2020-06-22 Paper2020ACM/Paper2020.tex
Carl updated the project

@@
 
-\subsubsection{Principles of Databases} 
-There are 3 principles needed when creating a database.
-The idea is that only two of the three can be focused on at any given time.
-Only two since the current technology can only handle two.
-These principles and limitations need to be taken into consideration when designing the system.
-\cite{Silberschatz2010}
-
-\begin{itemize}
-    \item Consistency - Does the data have integrity?
-To update one part, all other (replicated) parts should be updated quickly as well; Data is accurate and copied faithfully.
-An example would be that genes for specimen A, when copied to another data node should contain the exact copy of the data, with no changes.
-    \item Accessibility - How easy it is to access data, API protocols? An example would be the number of commands  a user needs to do to access every data.
-    \item Partition Tolerance - How resilient is the system when it comes to failures?
-If one portion breaks, the other portions should still be entirely usable.
-An example would be a power failure for one system, how would this affect the other systems?
-\end{itemize}
-
 \subsubsection{Database Types} 
@@
 For genomic research, when the data storage needs to be increased often, distributed databases are most appropriate.
+
+\subsubsection{CAP theorem for Databases} 
+There are 3 principles needed when creating a database.
+The idea is that only two of the three can be focused on at any given time.
+Only two since the current technology can only handle two.
+These principles and limitations need to be taken into consideration when designing the system.
+
+\begin{itemize}
+    \item Consistency - Does the data have integrity?
+To update one part, all other (replicated) parts should be updated quickly as well; Data is accurate and copied faithfully.
+An example would be that genes for specimen A, when copied to another data node should contain the exact copy of the data, with no changes.
+    \item Availability - Is the database usable by clients all of the time? The entire system must be in downtime as infrequently as possible, despite failures.
+    \item Partition Tolerance - How resilient is the system when it comes to failures?
+If one portion breaks, the other portions should still be entirely usable.
+An example would be a power failure for one system, how would this affect the other systems?
+\end{itemize}
+
+Partition tolerance is a de facto requirement of distributed databases and cannot be avoided, so the main trade-off is between availablity and consistency.
+\cite{Silberschatz2010} Suppose a distributed system prefers consistency.
+A user updates an SQL row in one partition of the database.
+Beforehand, this row has already been replicated among many partitions.
+To ensure that all replicas of the row are consistent, the system must lock this part of the database to prevent other users from reading and editing it, so availability will be affected in the meantime.
+A distributed system preferring availability, such as a messenger app, will not lock parts of the database in the purpose of being open to serve its users, but cannot guarantee that the data it outputs is always up-to-date.
+There might be a delay before the data is fully consistent among geographic partitions.
 

### 2020-06-22 Paper2020ACM/Paper2020.tex
Smol edits

@@
 \subsubsection{Community Aspect}
+
 - community: define as set of people who generally collaborate together to do research

### 2020-06-22 Paper2020ACM/Presentation.tex
Slight changes to presentation
Carl updated the project
Smol edits

@@
     \item Didn't have the mature tools at the time
-    \item Big cost, many institutions, 10 years
+    \item Big cost, many institutions, 10 years \cite{introgenomics}
   \end{itemize}
@@
   \begin{itemize}   
-    \item new technology to read sequences faster
-    \item reading faster -> more reads -> more data
+    \item New sequencing technology to read faster
+    \item Faster reads corresponds to more data
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Moore's Law}
+  \begin{block}{\textbf{Moore's law}} 
+    states that the number of transistors on a chip doubles every two years.\cite{kurose}
+  \end{block}
+  \begin{itemize}   
+    \item Prediction of the speed new technology (faster processing power and better storage) is developed
+    \item But moore's law is too slow for the speed of genomic data increase
+    \item
   \end{itemize}
@@
   \begin{itemize}   
-    \item moore's law make faster processing power and better storage
-    \item but moore's law is too slow for genomic data increase
-    \item each has 60x coverage, around 200gb each
+    \item Each read\cite{seqtorr} has 60x coverage, around 200gb each
     \item NGS, More genomic researches, more files needed to be saved and transferred

### 2020-06-23 Paper2020ACM/Paper2020.tex
New changes 5:50pm

@@
 Partition tolerance is a de facto requirement of distributed databases and cannot be avoided, so the main trade-off is between availablity and consistency.
-\cite{Silberschatz2010} Suppose a distributed system prefers consistency.
+\cite{Silberschatz2010} Suppose a distributed system prefers consistency over availability.
 A user updates an SQL row in one partition of the database.
 Beforehand, this row has already been replicated among many partitions.
-To ensure that all replicas of the row are consistent, the system must lock this part of the database to prevent other users from reading and editing it, so availability will be affected in the meantime.
+To ensure that all replicas of the row are consistent, the system must lock this part of the database to prevent other users from reading and editing it, so availability will be negatively affected in the meantime.
 A distributed system preferring availability, such as a messenger app, will not lock parts of the database in the purpose of being open to serve its users, but cannot guarantee that the data it outputs is always up-to-date.
-There might be a delay before the data is fully consistent among geographic partitions.
+There might be a delay before the data is fully consistent throughout geographic locations.
 
@@
 
-BioTorrents was successful, many people had used it to share data.
+BioTorrents was successful; many people had used it to share data.
 It was active for a while.
@@
 An important aspect of the proposed system is the transfer of files to other computers.
-To achieve that, there are network technologies and systems developed.
+To achieve that, there are network technologies and systems that need to be understood.
 Each layer of the network has certain features, securities, and abstraction which contributes to the overall idea of sharing files across networks.
@@
 This later on will be important to understanding the scope of what the thesis will do.
-Level 1 and 2 can be vulnerable if anyone can read the data in the air, partly why HTTPS was developed to encrypt the data.
-
-Level 4 and 5 is what the system will mainly use.
+Level 1 and 2 can be vulnerable if anyone can read the data inside the network, partly why HTTPS was developed to encrypt the data.
+
+Level 4 and 5 is what the system will mainly use and operate on.
 In these systems, there are APIs commonly used to transfer files between computers on networks.
@@
 Making sure files transferred are error free will be needed to do data synchronization later on.
+Another important network concept is that computers closer on the network (when talking directly) should have a faster speed and lower latency than computers farther apart on a network.
 
@@
 \subsubsection{Community Aspect}
-
-- community: define as set of people who generally collaborate together to do research
-- focus on genome research community
-- any users in a community can upload data
-- users can become data nodes
-- as data nodes increase, data speeds (bandwidth) should theoretically increase as well
-- system is an opportunity to set up genomic databases for frequently used local data
-
-- summary: community assists in adding data
-- community assists in adding data storage (through data nodes)
-- community assists in increasing the data speeds
-- community is assisted since the data it uses frequently should theoretically get accessed faster
-
-- addition: this is a framework to design and develop community-centered databases (can be modeled to be used beyond genome)
+The sytem was also described as community assisted.
+The definition of \textbf{community} to be used here is a \textit{set of people who generally collaborate together to do research}.
+And the focus is the genome research community.
+Reemphasizing certain features from the database, that any users can upload data.
+This means all users in the community can contribute to the knowledge base, and contribute data to be used by the community.
+Another important community aspect is that users can become data nodes themselves.
+Meaning these users can increase the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later).
+Frequently used local data as well will be more easily accessible (since it's on local data nodes, and should have faster access speeds).
+The concept of closer networks having faster speeds (discussed in the network section) is at play here.
+Another community aspect will be this paper, publicly describing many aspects of the structure and code.
+So that this concept may be further developed and even implemented by genome research communities.
+
+To summarize, this helps the community since the community may add their own data.
+The community may increase storage (through data nodes) these may also assist in increasing data speeds.
+And frequently used data will have faster access.
 
@@
 
+Data aross Data Nodes is consistent
+
 A proof of correctness for the implementation.
@@
 
-Theoretical analysis of the speed up by checking time complexity of the base method versus the distributed method.
-
-in the base method
-client sends data in X bytes per second. When X people are using. 
-(insert math equation)
-Each person gets the file at 1 byte per second.
-
-in the proposed method
-client has N data nodes and each node sends data in X bytes per second. When X people are using. Assuming perfect and equal distribution. 
-(insert math equation)
-Each person gets  at least N bytes per second
-
-Since N is always greater than 1, the speed should always be greater in the proposed method
-
-As user number increases, base method will reach a limit to its ability to handle data transfer.
-
-As user number increases, proposed method will also reach a limit. But scalability is possibile to remove the limit.
+Theoretical analysis of the speed up by checking time complexity of the base method versus the proposed method.
+
+\subsection{Base Method}
+In the base method, represented by NCBI.
+We will abstract the transfer into this algorithm.
+Assuming as well, constant speed, equal division of bandwidth, all users are getting the same data all at once, and there is a constant overhead before downloading can occur.
+
+\subsubsection{Base Method Algorithm}
+
+1. There is a constant overhead H seconds to set up the communication between the user and the main server.
+1. The main server sends data in the speed of X bytes per second. X is the speed of data transfer from the server.
+2. Users will receive the data with an equal portion of the X bytes per second (bandwidth).
+
+\subsubsection{Base Case of the Base Method}
+1 total users will receive data in the speed of X/1 bytes per second.
+2 total users will receive data in the speed of X/2 bytes per second.
+
+With simple induction we can assume the general case to be:
+
+\subsubsection{General Case of the Base Method} 
+N total users will receive data in the speed of X/N bytes per second. N is the number of users sharing the bandwidth.
+
+\subsubsection{Notable Cases of the Base Method}
+If N=X, all users will receive data with 1 byte per second.
+If N>X, all users will receive data with less than 1 byte per second, or if we forego the assumption of total equal division: Some will receive 1 byte per second, and others will receive 0 bytes per second or no transfer at all.
+This a bottleneck scenario.
+
+\subsection{Proposed Method}
+For the proposed system, it is very similar to the base method, but the idea of data nodes (D) will be introduced.
+We will abstract the system into this algorithm.
+Assuming as well, constant speed, equal division of bandwidth, all users are getting the same data as once, and there is a constant overhead before downloading the data for each user, there is also an equal division of the data nodes' bandwidth to the users.
+
+\subsubsection{Proposed Method Algorithm for One Data Node}
+1. There is a constant overhead H seconds to set up the communication between the user and the relevant nodes.
+2. One data node will send data in the speed of X bytes per second. X is the speed of data transfer from 1 data node.
+3. Users will receive data with an equal portion of the X bytes per second.
+
+\subsubsection{Proposed Method Algorithm for Multiple Data Nodes}
+1. There is a constant overhead H seconds to set up the communication between the user and the relevant nodes.
+2. All data nodes (D number of data nodes) will send data in the speed of D times X bytes per second. X is the speed of data transfer from 1 data node.
+3. Users will receive data with an equal portion of the D times X bytes per second.
 
@@
 - one data source
-- single point of failure
+- single point of failure \cite{seqtorr}
 - implemented

### 2020-06-23 Paper2020ACM/Presentation.tex
New changes 5:50pm

@@
     \item Each read\cite{seqtorr} has 60x coverage, around 200gb each
-    \item NGS, More genomic researches, more files needed to be saved and transferred
+    \item with NGS, more genomic researches, more files needed to be saved and transferred
   \end{itemize}
@@
   \begin{itemize}   
-    \item store data with tech
+    \item Technology for data storage
+    \item Technology needed to be developed to resolve the growing storage concern
   \end{itemize}
@@
 \begin{frame}{Database Definition}
+A database or a DBMS (DatBase Management System) \cite{Silberschatz2010} is defined as:
 \begin{itemize}
-    \item DBMS chuchu
+    \item Collection of interrelated data
+    \item Set of programs to access the data
+    \item An environment that is both convenient and efficient to use
 \end{itemize}
@@
     \item Distributed
-    \item use Distributed becuz x
+    \item Distributed will be focused since it has the ability to speed up transfer speeds
   \end{itemize}
@@
     \item Hybrid
-    \item Use Hybrid because X
+    \item Hybrid architecture will be used because the Client-Server has bottlenects, and the P2P has no centralized management
   \end{itemize}

### 2020-06-23 Paper2020ACM/Paper2020.tex
9pm version

@@
 \subsubsection{Base Method Algorithm}
-
-1. There is a constant overhead H seconds to set up the communication between the user and the main server.
-1. The main server sends data in the speed of X bytes per second. X is the speed of data transfer from the server.
-2. Users will receive the data with an equal portion of the X bytes per second (bandwidth).
+\begin{enumerate}
+    \item There is a constant overhead H seconds to set up the communication between the user and the main server.
+    \item The main server sends data in the speed of X bytes per second. X is the speed of data transfer from the server.
+    \item Users will receive the data with an equal portion of the X bytes per second (bandwidth).
+\end{enumerate}
 
 \subsubsection{Base Case of the Base Method}
-1 total users will receive data in the speed of X/1 bytes per second.
-2 total users will receive data in the speed of X/2 bytes per second.
-
-With simple induction we can assume the general case to be:
+    \begin{itemize} 
+        \item 1 total users will receive data in the speed of X/1 bytes per second. Downloading data (size of X bytes) will take X/X + H seconds to transfer.
+        \item 2 total users will receive data in the speed of X/2 bytes per second. Downloading data (size of X bytes) will take X/(2X) + H seconds to transfer.
+    \end{itemize}
 
 \subsubsection{General Case of the Base Method} 
-N total users will receive data in the speed of X/N bytes per second. N is the number of users sharing the bandwidth.
+With simple induction we can assume the general case to be: N total users will receive data in the speed of X/N bytes per second, with H seconds of an initial overhead.
+N is the number of users sharing the bandwidth.
+Downloading data (size of X bytes) will take X/(NX) + H or (1/N) + H seconds to transfer.
+A notable aspect of this general case is that an increase in the number of users decreases the user download speed.
 
 \subsubsection{Notable Cases of the Base Method}
-If N=X, all users will receive data with 1 byte per second.
-If N>X, all users will receive data with less than 1 byte per second, or if we forego the assumption of total equal division: Some will receive 1 byte per second, and others will receive 0 bytes per second or no transfer at all.
+    \begin{itemize} 
+        \item If N = X, all users will receive data with 1 byte per second.
+        \item If N > X, all users will receive data with less than 1 byte per second, or if we forego the assumption of total equal division: Some will receive 1 byte per second, and others will receive 0 bytes per second or no transfer at all.
 This a bottleneck scenario.
+    \end{itemize}
 
@@
 We will abstract the system into this algorithm.
-Assuming as well, constant speed, equal division of bandwidth, all users are getting the same data as once, and there is a constant overhead before downloading the data for each user, there is also an equal division of the data nodes' bandwidth to the users.
+Assuming as well, constant speed, equal division of bandwidth, all users are getting the same data as once, there is a constant overhead before downloading the data for each user, there is also an equal division of the data nodes' bandwidth to the users, and all data nodes have an exact copy of the data.
 
 \subsubsection{Proposed Method Algorithm for One Data Node}
-1. There is a constant overhead H seconds to set up the communication between the user and the relevant nodes.
-2. One data node will send data in the speed of X bytes per second. X is the speed of data transfer from 1 data node.
-3. Users will receive data with an equal portion of the X bytes per second.
+
+\begin{enumerate}
+    \item There is a constant overhead H seconds to set up the communication between the user and the relevant nodes.
+    \item One data node will send data in the speed of X bytes per second. X is the speed of data transfer from 1 data node.
+    \item Users will receive data with an equal portion of the X bytes per second. 
+\end{enumerate}
+
+\subsubsection{Base Case of the Proposed Method for One Data Node}
+
+\begin{itemize} 
+    \item 1 total users (and 1 data node) will receive data in the speed of 1 times X/1 bytes per second. Downloading data (size of X bytes) will take X/X + H seconds to transfer.
+    \item 2 total users (and 1 data node) will receive data in the speed of 1 times X/2 bytes per second. Downloading data (size of X bytes) will take X/(2X) + H seconds to transfer.
+\end{itemize}
+
+\subsubsection{General Case of the Proposed Method for One Data Node} 
+Exactly the same as the  General Case of the Base Method, N total users and 1 data node will receive data in the speed of X/N bytes per second with H seconds of an initial overhead.
+N is the number of users sharing the bandwidth.
+Downloading data (size of X bytes)  will take X/(NX) + H or (1/N) + H seconds to transfer.
 
 \subsubsection{Proposed Method Algorithm for Multiple Data Nodes}
-1. There is a constant overhead H seconds to set up the communication between the user and the relevant nodes.
-2. All data nodes (D number of data nodes) will send data in the speed of D times X bytes per second. X is the speed of data transfer from 1 data node.
-3. Users will receive data with an equal portion of the D times X bytes per second.
+\begin{enumerate}
+    \item There is a constant overhead H seconds to set up the communication between the user and the relevant nodes.
+    \item All data nodes (D number of data nodes) will send data in the speed of D times X bytes per second. X is the speed of data transfer from 1 data node.
+    \item Users will receive data with an equal portion of the D times X bytes per second.
+\end{enumerate}
+
+\subsubsection{Base Case of the Proposed Method for Multiple Data Nodes}
+\begin{itemize}
+    \item 1 total users (and D data nodes) will receive data in the speed of D times X/1 bytes per second. Downloading data (size of X bytes) will take DX/X + H seconds to transfer.
+    \item 2 total users (and D data nodes) will receive data in the speed of D times X/2 bytes per second. Downloading data (size of X bytes) will take DX/(2X) + H seconds to transfer.
+\end{itemize}
+
+\subsubsection{General Case of the Proposed Method for Multiple Data Nodes} 
+This is different from the General Case of the Proposed Method for Multiple Data Nodes by a factor of D.
+N total users and D data nodes will receive data in the speed of DX/N bytes per second with H seconds of an initial overhead.
+N is the number of users sharing the bandwidth.
+Downloading data (size of X bytes)  will take DX/(NX) + H or (D/N) + H seconds to transfer.
+A notable aspect of this general case is that as the number of data nodes increases, the user's speed also increases.
+Even if the previous feature from the general case of the base method is still at play, (an increase in the number of users decreases the user download speed).
+
+\subsubsection{Notable Cases of the Proposed Method}
+\begin{itemize}
+    \item  If D = 1, N = X, all users will receive data with 1 byte per second (base method and proposed method for one data node is excactly the same).
+    \item If D = 1, N > X, all users will receive data with less than 1 byte per second, or if we forego the assumption of total equal division: Some will receive 1 byte per second, and others will receive 0 bytes per second or no transfer at all.
+This a bottleneck scenario.
+(also the same).
+    \item If D = 0, there is no data transfer at all. This is an extreme case.
+    \item If D = 1, the proposed method will match the base case, general cases, and notable cases for the base method
+    \item If D > 1, the proposed method will be faster than the base method (as long as H, X, N are the same).
+\end{itemize}
+
+\subsection{Sumamry of the Theoretical Analysis of Speed}
+Generally as the number of users increases, the data transfer speed decreases by a constant factor for all methods.
+There is also a concept of bottleneck when the number of users is much greater than the data speed put out by all methods.
+
+The proposed method (with D > 1, and same H, X, N) will be faster than the base method.
+A notable extreme for the proposed method is that once D = 0, no transfer will occur and will be slower than the bottleneck base method.
+With this analysis, it is suggested the proposed method is faster than the base method on most cases (D > 1).
 
@@
 
-- increase in data speed transfer (if there are sufficient available data nodes)
+- the section on theoretical speed increase in data speed transfer (data nodes > 1)
 - can reduce the data stored to house only locally used and needed data

### 2020-06-24 Paper2020ACM/Paper2020.tex
June 24 8am
Revisions June 24 10:30

@@
 
-\subsubsection{CAP theorem for Databases} 
+\subsubsection{CAP theorem for Databases} \label{cap}
 There are 3 principles needed when creating a database.
@@
 
-\subsection{NCBI}
+\subsection{NCBI} \label{NCBI}
 The National Center for Biotechnology Information or more commonly known as NCBI, which discussed briefly in the section on \textit{BioTorrents} and in the \textit{Introduction}.
@@
 \begin{enumerate}
-    \item No public existing usable public framework on distributed databases for genomes
+    \item No existing usable public framework on distributed databases for genomes
     \item Existing distributed database system for genomes is not replicable
@@
     \begin{itemize}
-        \item To resolve 3 \& 4 of the Problems of Classical Databases. 
+        \item To resolve 3 \& 4 of the Problems of Classical Databases, and 1 \& 2 of the Problems of Distributed Databases. 
     \end{itemize}
@@
 
-\subsubsection{Community Aspect}
+\subsubsection{Community Aspect} \label{community}
 The sytem was also described as community assisted.
@@
 
-%HIII NEED HELP HEREE
 \section{Proof of Correctness}
-correct if
-GET - gets data
-POST - puts data
-PUT - edits data
-DELTE - deletes data
-
-speed should be X
-
-Data aross Data Nodes is consistent
+Hii i need help fixing this section di ko pa alam ano ilalagay
+
+The distributed database will function correctly if:
+
+\begin{enumerate}
+    \item GET transfers the data to the user with no error
+    \item POST transfers the data to the data node with no error
+    \item PUT edits the data
+    \item DELETE removes the data
+    \item all DATA NODES with DATA X have the exact copy of DATA X
+\end{enumerate}
 
@@
 \subsection{Base Method}
-In the base method, represented by NCBI.
+In the base method, represented by NCBI (see Section \ref{NCBI}).
 We will abstract the transfer into this algorithm.
@@
 \subsubsection{General Case of the Proposed Method for Multiple Data Nodes} 
-This is different from the General Case of the Proposed Method for Multiple Data Nodes by a factor of D.
+This is different from the General Case of the Proposed Method for a Single Data Node by a factor of D.
 N total users and D data nodes will receive data in the speed of DX/N bytes per second with H seconds of an initial overhead.
@@
 
-\subsection{Sumamry of the Theoretical Analysis of Speed}
+\subsection{Summary of the Theoretical Analysis of Speed} \label{theoreticalspeed}
 Generally as the number of users increases, the data transfer speed decreases by a constant factor for all methods.
@@
 
-\section{Comparison of base method vs distributed method}
-
-base method (NCBI)
-- houses a lot of biological data
-- go to website
-- one website
-- one data source
-- single point of failure \cite{seqtorr}
-- implemented
-
-proposed method
-- ability to scale
-- multiple (scalable data nodes) data sources
-- multiple (scalable data nodes) data links
-- not yet implemented
-- can be done / implemented on a local scale (limit nodes to the needs)
-- only have data for a certain data community
-- only have users from a certain data community
-- able to design the APIs for speed, security, consistency, tolerance, etc. depending on community needs
-
-- the section on theoretical speed increase in data speed transfer (data nodes > 1)
-- can reduce the data stored to house only locally used and needed data
+\section{Comparison of base method vs proposed method}
+
+For the final comparisons of the base method the proposed method, the current base method (NCBI in Section \ref{NCBI}) is the go-to website for biological data and genomic data.
+It is currently used by and is one website and one server.
+This points to a single point of failure\cite{seqtorr}.
+The section \ref{theoreticalspeed} on theoretical analysis of speed described the problem of having a bottleneck when the number of users are too many for the data bandwidth of the server.
+
+The proposed method introduces the idea of data nodes and in section \ref{theoreticalspeed} explains how having at least one data node should theoretically rival the speed of the base method.
+And having multiple data nodes will be faster in theory than the base method.
+Another advantage of having a distributed database is the community aspect (see Section \ref{community}).
+The may be implemented on a local scale, meaning it can have the necessary data for the community.
+The most frequently used data, may be put on more data nodes increasing the speed of the data transfer (see Section \ref{theoreticalspeed}).
+Another is that the features on the designs of the speed, consistency, and fault tolerance (see Section \ref{cap} ) may be calibrated to the needs of the community.
 
 \section{Recommendations}
-
-Implement this. Describe the technologies used and why. And make the code public. Better yet create a template code to easily copy and implement own data centers.
-
-Implementation problems. The network API will face problems in port access especially in UPD internet. Only 2 ports are currently available.
-
-Have a set of testing APIs and have testing APIs embedded in each component. And the testing consistently checks which data have integrity.
-
-Have security.
-The API structure is one way of having security in the future even if it isnt implemented yet.
-Restructure and redesigning the APIs are one way to achieve security.
-Adding user authentication is one way to achieve this.
-As well as having a separate layer between user and master.
-Where all interactions happen in that layer (even data transfer??).
-Abstract this project even further.
-Creating means for master nodes to communicate and collaborate with other master nodes.
-This will let different caDDS (assuming this is implemented) to communicate and collaborate (send data) to other caDDS.
-
-Abstract this project for non-genomic data as well.
-To develop the idea of a distributed data community (where distributed databases are easy to scale, implement, and connect).
-Which can help share many kinds of scientific data more easily.
-A specific suggestion is for other fields with large data gathered, such as space image data.
+A recommendation is to implement this proposed system. It is advised to have a public code with an easy-to-start documentation.
+
+A small but important technical step is to pay attention to port access especially with university internet, some universities only limit the usable ports to two and this would be an important hurdle to overcome in managing multiple data nodes.
+
+Another is to look into security, which aspects of the code or data must be private or moderated.
+How does data privacy work?
+And how does security affect the accessibility of the system?
+What are the different risks in distributed databases and how to prevent them?
+Many questions arise in security but these will only be answered better once implementation starts.
+
+A recommendation is to look into a way for master nodes to communicate with other master nodes, this will start the idea of communication and collaboration between caDDS to other caDDS.
+
+The last recommendation is to abstract this project for non-genomic data as well.
+This is good especially for other scientific communities with its own growing data set such as cosmology with its space image data.
 

### 2020-06-24 Paper2020ACM/Paper2020.tex
June 24 2pm

@@
 
-(Not yet serious) Our advisers have really helped us over the school year. The dificult pandemic has changed our thesis to the point it has been - Cid
+We want to acknowledge our advisers for giving the idea for out thesis.
+And for answering our many questions about genomics.
+Writing this paper overlapped with the covid19 pandemic which changed a big portion of the scope of this thesis.
+We thank our advisers for making this work manageable during this pandemic crisis.
 \end{acks}

### 2020-06-24 Paper2020ACM/Paper2020.tex
Revision June 24 5pm

@@
 \begin{abstract}
-working draft (insert kulang when info is ready):
 Genomic data the past few years have grown exponentially with the technology barely keeping up.
@@
 The proposed system is aimed towards a smaller community, with more frequent and localized data, which the data storage can easily be expanded as the need arises.
+The proposed system is also theoretically faster as long as there are multiple working data nodes.
 A recommendation is to implement the system and make the code public for the use of other communities.
@@
 
-We want to acknowledge our advisers for giving the idea for out thesis.
+We want to acknowledge our advisers for giving the idea for our thesis.
 And for answering our many questions about genomics.

### 2020-06-24 Paper2020ACM/Presentation.tex
June 24 2pm
Revision June 24 5pm

@@
   \begin{itemize}   
-    \item Technology for data storage
+    \item Database is the technology for data storage
     \item Technology needed to be developed to resolve the growing storage concern
@@
 \begin{frame}{Database Principles}
+
   \begin{itemize}   
@@
   \end{itemize}
+\end{frame}
+
+\begin{frame}{Table}
+% Bio-related -> biological
+% What's the difference between biological vs general vs FASTA
+% What does general mean?
+% What does biological mean? -> explain pa to
+% May 3 types of data ...
+   \begin{table}[]
+        \begin{tabular}{|l|l|l|l|l|}
+        \hline
+        System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ \hline
+        BioTorrents & P2P           & No         & Depends                    & Biological  \\ \hline
+        SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ \hline
+        PeerDB      & P2P           & No         & Depends                    & General      \\ \hline
+        NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\ \hline
+        \end{tabular}
+    \end{table} 
 \end{frame}
@@
   \begin{itemize}   
-    \item simple time complexity chart
-    \item faster si caDDS ofc 
+    \item base method
+    \item proposed method
     \item 
@@
   \begin{itemize}   
-    \item caDDS is scalable, easy for community, has docu hehe
-    \item NCBI is one, stable, widely used
+    \item base method
+    \item proposed method
     \item 

### 2020-06-24 Paper2020ACM/Presentation.tex
Added pictures and old content from old slides

--- 2020-06-22 Paper2020ACM/Presentation.tex
@@
 \subsection{Genomic Concepts}
+
+\begin{frame}{Human DNA Length}
+    \centering
+    \includegraphics[scale=0.3]{dna-length.jpg}
+\end{frame}
+
 \begin{frame}{Genes}
@@
     \item Genes are made of a certain molecular structure called DNA which correspond to information denoted by: A,G,C,T
-    \item genes are instructions and blueprints for life
+    \item Genes are instructions and blueprints for life
   \end{itemize}
+\centering
+\includegraphics[scale=0.1]{gene-base-pair.png}
 \end{frame}
@@
   \end{itemize}
+  \centering
+  \includegraphics[scale=0.2]{sequencing.png}
 \end{frame}
@@
   \begin{itemize}
-    \item New sequencing technology to read faster
+    \item original sequencing for HGP took ten years, now (2019) one laboratory can do 10,000 a day (36.5 million times faster)  \cite[p.~19]{introgenomics}   
+    \item New sequencing technology to read faster and cheaper
     \item Faster reads corresponds to more data
   \end{itemize}
+\end{frame}
+
+\begin{frame}{Human genome cost over time}
+Graph of human genome cost over time \cite{genomics-cost} \\
+\centering
+\includegraphics[scale=0.3]{human-gen-cost.jpg}
+\end{frame}
+
+\begin{frame}{Sequence data cost over time}
+Graph of a megabase sequence cost over time \cite{genomics-cost}
+\begin{itemize}
+    \item Megabase means one million bases
+\end{itemize}
+\centering
+\includegraphics[scale=0.1]{seq-cost.jpeg}
+\end{frame}
+
+\begin{frame}{NGS vs Hard Disk Storage}
+        Computational sense, there is a need to store, upload, and retrieve sequence files that are growing larger. 
+        \begin{itemize}
+        \item Slope of orange line (NGS) is larger than slope of yellow line (pre-NGS)
+        \item Hard disk storage (blue) graph is greater than yellow (pre-NGS) but is now less than orange (NGS)
+        \end{itemize}
+        \centering
+            \includegraphics[scale=0.2]{seq-data.jpeg}
 \end{frame}
@@
 
-\subsection{Database Concepts}
+\subsection{Data Storage Concepts}
 \begin{frame}{Database}
@@
 \end{frame}
+  \begin{frame}{Data Storage}
+        A \textbf{database} (\textbf{DBMS}, Database Management System) is defined by 
+        \begin{block} {\textbf{Definition}}
 
-\begin{frame}{Database Definition}
+        DBMS contains information about a particular enterprise
         \begin{itemize}
-    \item DBMS chuchu
+            \item Collection of interrelated data
+            \item Set of programs to access the data 
+            \item An environment that is both convenient and efficient to use
+        \end{itemize}
+        \cite{Silberschatz2010}
+        \end{block}
+    \end{frame}
+
+    \subsubsection{Database Principles}
+    \begin{frame}{Database Principles}
+        \begin{itemize}
+            \item \textbf{Principles} to remember when creating a database (CAP). Idea is that only 2 of the three can be focused on. \cite[Ch.~19]{Silberschatz2010}
+            \begin{itemize}
+                \item \textbf{Consistency}
+                \begin{itemize}
+                    \item Update one part, will all other (replicated) parts be updated quickly as well
+                    \item Data is accurate and copied faithfully
+                \end{itemize}
+                \item \textbf{Accessibility}
+                \begin{itemize}
+                    \item How easy it is to access data
+                    \item API protocols
+                \end{itemize}
+                \item  \textbf{Partition Tolerance}
+                \begin{itemize}
+                    \item If one portion breaks, the other portions are still usable
+                    \centering 
+                    \includegraphics[scale=0.2]{partition_tolerance.png}
+                    \end{itemize} 
+             \end{itemize}
+        \end{itemize}
+    \end{frame}
+  \begin{frame}{Normal: NCBI}
+        \begin{itemize}
+            \item NCBI has a large collection of genomic datasets in its website, which can be searched and downloaded.
+        \end{itemize}
+    \end{frame}
+    \begin{frame}{Normal: NCBI}
+        \begin{itemize}
+            \item Search page https://www.ncbi.nlm.nih.gov/nucleotide
+            \includegraphics[scale=0.5]{ncbi1.png}
+        \end{itemize}
+    \end{frame}
+    \begin{frame}{Normal: NCBI}
+        \begin{itemize}
+            \item List of results
+            \includegraphics[scale=0.5]{ncbi2.png}
+        \end{itemize}
+    \end{frame}
+    \begin{frame}{Normal: NCBI}
+        \begin{itemize}
+            \item Download page
+            \includegraphics[scale=0.5]{ncbi3.png}
         \end{itemize}
@@
 
-\begin{frame}{Database Principles}
+    % cinomment out ko kasi meron naman explanation sa taas
+    %\subsection{P2P Database}
+    %\begin{frame}{P2P Databases}
+    %    something explaining briefly p2p databases
+    %\end{frame}
+
+        \subsection{BioTorrents}
+    \begin{frame}{P2P: BioTorrents}
         \begin{itemize}
-    \item consistency
-    \item availability
-    \item partition tolerance
+            \item Scientific data continues to grow, and so does the demand for easier accessibility
+            \item Centralized servers using HTTP or FTP cannot keep up with concurrent requests
+            \item Peer-to-peer protocols does not scale well for large files
+            \item BitTorrent handles both these problems
         \end{itemize}
@@
 
-\begin{frame}{Database Types}
+    \begin{frame}{P2P: BioTorrents}
         \begin{itemize}
-    \item Centralized
-    \item Distributed
-    \item use Distributed becuz x
+        \item System for legally sharing scientific data that works on top of the BitTorrent protocol. It is much like a public tracker.
+        \item The main website hosts .torrent and metadata files
+        \item Info about each dataset includes categories, license, filenames, etc. that helps users in searching for relevant datasets
+        \item Torrent file has list of trackers, servers that "know" which peers are serving which files.
+        \item Torrent client downloads from multiple peers simultaneously.
+        \item To upload a torrent, the client makes a .torrent file using the torrenting software, then uploads it to the main website along with metadata.
+The client must have their computer continuously turned on to keep the file available as a peer.
+\cite{biotorrents}
         \end{itemize}
@@
 
-\begin{frame}{Application Architecture}
+    \begin{frame}{P2P: BioTorrents}
+        \includegraphics[scale=0.33]{biotorrents-page.png}
+    \end{frame}
+
+    \begin{frame}{P2P: BioTorrents}
         \begin{itemize}
-    \item Client Server
-    \item P2P
-    \item Hybrid
-    \item Use Hybrid because X
+            \item As with all public trackers, BioTorrents requires "good souls" that will keep many torrents available as often as possible.
+            \item The datasets must be moderated to prevent abuse.
+            \item The BioTorrents website is no longer up.
         \end{itemize}
@@
 
-\section{Review of Related technology}
-
-\begin{frame}{SeqTorr}
+    \subsection{PeerDB}
+    \begin{frame}{P2P: PeerDB}
+    % try adding screenshots
         \begin{itemize}
-    \item idea
-    \item pros
-    \item cons
+            \item Proposed P2P system for sharing general data
+            \item Each node is equipped with a database
+            \item Queries are passed around neighbors, with a maximum number of hops
+            \item Caches information about other peers to speed up queries
+            \item Large number of nodes \& relatively small number of lookup servers
+            \cite{peerdb}
         \end{itemize}
@@
 
-\begin{frame}{BioTorrents}
+    \subsection{Distributed Database}
+    \begin{frame}{Distributed Databases}
         \begin{itemize}
-    \item idea
-    \item pros
-    \item cons
+            \item \textbf{Distributed Database}
+            \begin{itemize}
+                \item A distributed database system consists of loosely coupled sites that share no physical component             
+                \item Database systems that run on each site are independent of each other
+                \item Transactions may access data at one or more sites
+                \cite[Ch.~19]{Silberschatz2010}
+                \item Beneficial for large data with a certain speed requirement
+            \end{itemize}
         \end{itemize}
@@
 
-\begin{frame}{PeerDB}
+    \subsection{SeqTorr}
+    \begin{frame}{Distributed: SeqTorr}
+    \textbf{SeqTorr} \cite{seqtorr}
     \begin{itemize}
-    \item idea
-    \item pros
-    \item cons
+        \item Standard genomics workflow consists of uploading and downloading data on an international database like NCBI.
+        \item Having a distributed scalable local infrastructure to store genomic data instead of relying on NCBI would be beneficial since
+        \begin{itemize}
+            \item only certain sequences in NCBI are relevant to the institutions (e.g. Asian sequences are needed more than Caucasian ones)
+            \item researchers within the country can share data with fast speeds
+            \item institutions can share in the hosting of data
+        \end{itemize}
+    \end{itemize}
+    \end{frame}
+    \begin{frame}{Distributed: SeqTorr}
+    \begin{itemize}
+        \item Utilizes distributed database
+        \item Master node
+        \begin{itemize}
+            \item containing metadata of all the sequences
+            \item where user authentication is handled
+        \end{itemize}
+        \item Data node
+        \begin{itemize}
+            \item contain a portion of all the sequences
+            \item needs to replicate a sequence \texttt{r} times before a sequence is considered available
+            \item proximity of data nodes can increase download speed
+        \end{itemize}
+        \item Defined API Implementation
+        \begin{itemize}
+            \item POST
+            \item GET
+            \item PUT
+            \item DELETE
+        \end{itemize}
+
     \end{itemize}
@@
 
-\begin{frame}{base system: NCBI}
+    \begin{frame}{Distributed: SeqTorr}
+    % have a unified way of discussing each database
+    % what aspects of the database to discuss each
+    % extra features will come after
+        Problems: % have one for each database system
         \begin{itemize}
-    \item why popular
-    \item idea
-    \item idea
+            \item program not currently available
+            \begin{itemize}
+                \item cannot measure any statistics on this
+            \end{itemize}
         \end{itemize}
@@
 
-\begin{frame}{base system: NCBI}
-  \begin{itemize}   
-    \item pros
-    \item cons
-    \item cons
-  \end{itemize}
+    \begin{frame}{Table}
+% Bio-related -> biological
+% What's the difference between biological vs general vs FASTA
+% What does general mean?
+% What does biological mean? -> explain pa to
+% May 3 types of data ...
+   \begin{table}[]
+        \begin{tabular}{|l|l|l|l|l|}
+        \hline
+        System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ \hline
+        BioTorrents & P2P           & No         & Depends                    & Biological  \\ \hline
+        SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ \hline
+        PeerDB      & P2P           & No         & Depends                    & General      \\ \hline
+        NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\ \hline
+        \end{tabular}
+    \end{table} 
 \end{frame}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PROBLEM %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Problem}
+% after discussion of dist. databases
 \begin{frame}{Problem}
-  \begin{itemize}   
-    \item classical databases
-    \item distributed databases
-  \end{itemize}
+The different gaps or problems found with other implementations
+\textbf{Problems of Classical Databases}
+\begin{enumerate}
+%PROBLEMS OF CLASSICAL DATABASES (ALL FOUR)
+    \item Genome researchers download entire database of genomic data
+    \item Genome researchers only need a portion of the database, but end up downloading everything  (gigabytes size)
+    \item Client-Server approach can face bottlenecks (server speed) %is resolved by making it more distributed
+    \end{enumerate}
+% or having a hard time getting a portion of the database
+% easier to download parts vs whole, whole vs parts
+
+% not easy to navigate the database? 
+% what is the big problem? too big data to download all at once?
+% why is it a problem? speed of download? or size?
+
+\textbf{Problems of Distributed Databases}
+
+\begin{enumerate}
+%PROBLEM OF DIST. DATABASES
+    \item Only one framework on distributed database for genomes
+    \item Existing distributed database for genome code has little documentation % concept of APIs and documentation
+
+\end{enumerate}
 \end{frame}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\titlepage % Print the title page as the first slide
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OBJECTIVES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Objectives}
+% to tackle those problems
+% should be written in connection with the problems said:
+% problem x : objective x
+% para mas clear yung objectives
+
 \begin{frame}{Objectives}
+Based on the problems we found earlier, we propose the following objectives
  \begin{itemize}
-    \item make a dist database
-    \item community assisted
-    \item main APIs
-    \item time complexity
-    \item pros and cons
+    \item Create an information system that speeds up the download and upload of genomic data
+    \begin{itemize}
+        \item To resolve 1 \& 2 of the Problems of Classical Databases. 
+    \end{itemize}
+\end{itemize}
+
+\begin{itemize}
+    \item The system should have the main API’s
+    \begin{itemize}
+        \item POST genome data
+        \item GET genome data (implement hybrid P2P)
+        \begin{itemize}
+            \item This will aim to resolve number 3 in the Problems of Classical Databases
+        \end{itemize}
+        \item PUT genome metadata (optional: PUT data)
+        \item DELETE genome data
+    \end{itemize}
+    \item Write documentation for the system
+    \begin{itemize}
+        \item This will be used to resolve the Problems of Distributed Databases
+    \end{itemize}
 \end{itemize}
@@
 \begin{frame}{Scope}
+The current scope and coverage of the project is only
 \begin{itemize}
-    \item no this
-    \item and
-    \item that
+    \item Security and Authentication is not part of the scope
+    \item Use only FASTA file format 
+    \begin{itemize}
+        \item DNA nucleotides only
+    \end{itemize}
+    \item Only a proof of concept, UI will be purely for functional purposes (very basic design)
 \end{itemize}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% THEORETICAL FRAMEWORK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Theoretical Framework}
-\begin{frame}{Network}
+
+        \begin{frame}{Network Layers}
+
+         \textbf{Networks Layers} how computers pass data to each other
         \begin{itemize}
-    \item five levels
-    \item need to focus on Level 4 and 5 only
-    \item 
+            \item \begin{itemize}
+            \item Level 1: Medium used to send signals (physical, electrical signals of 0 and 1), through wire or by air
+            \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together different . Goal of sending a group of signals.
+            \item Level 3: Introduction of networks: IP Address to uniquely identify entities. 
+            \item Level 4: Management of error free data transmission: Sending a group of messages via TCP, UDP, FTP
+            \item \textbf{Level 5}: Abstraction of Level 4: Function calls programmers will use to do level 4 processes.
             \end{itemize}
+        \end{itemize}
+    \end{frame}
+
+    \subsection{Application Architecture}
+    % part ng theoretical framework
+    % pwedeng walang slide of c/s
+    % pwedeng ibanggit nalang ang client-server architecture
+
+    % methods: ito framework
+    % implemens client-server...
+    \begin{frame}{Application Architecture}
+        \begin{itemize}
+            \item The \textbf{application architecture} is how a networked app is structured across the various end systems (i.e. computers, servers, etc.) \cite{kurose}
+            \item Falls into either client-server, peer-to-peer, or hybrid architecture
+        \end{itemize}
+    \end{frame}
+
+        \subsubsection{Client-Server}
+        \begin{frame}{Client-Server}
+            \begin{itemize}
+                \item A \textbf{client-server} architecture relies on a Web server that is always available and has a fixed IP address\\
+                \item A single server connects to many clients
+            \end{itemize}
+            \centering
+            \includegraphics[scale=0.5]{Server-based-network.png}
+        \end{frame}
+
+        \subsubsection{Peer-to-Peer \& Hybrid}
+        \begin{frame}{Peer-to-Peer \& Hybrid}
+            \begin{itemize}
+                \item A \textbf{peer-to-peer (P2P)} architecture has minimal to no reliance on always-on, dedicated servers
+                \item Peers are desktops/laptops that communicate to each other without passing through a dedicated server
+                \item A \textbf{hybrid} architecture combines elements of P2P and client-server architectures, attempting to utilize advantages of both
+            \end{itemize}
+            \centering
+            \includegraphics[scale=0.5]{P2P-network.png}
+        \end{frame}
+
+        \begin{frame}{Bittorent terminology}
+            % ADD SCREENSHOT
+            Peers \\ \\ 
+             \centering
+            \includegraphics[scale=0.5]{bit-2.PNG}
+        \end{frame}
+
+        \begin{frame}{Bittorent terminology}
+            % ADD SCREENSHOT
+            Trackers \\ \\
+             \centering
+            \includegraphics[scale=0.5]{bit-3.PNG}
+        \end{frame}
+
+        \begin{frame}{BitTorrent terminology}
+            % visually show each term din
+            % screenshot of torrent
+            \begin{itemize} % siguro breeze through what these mean 
+                \item \textbf{Peers} are users that connect to the BitTorrent network.
+                \item \textbf{Torrent files} contain metadata about files \& folders, trackers, and hash values.
+                \item \textbf{Trackers} are servers that keep track of peers and availability of files
+                \item \textbf{Seeds} are peers which have downloaded the entire file and are making it available
+            \end{itemize}
+        \end{frame}
+
+\begin{frame}{Theoretical Framework: System Description}
+    \begin{itemize}
+        \item A \textbf{master node} stores metadata of all sequences, may or may not contain sequence files
+        \item \textbf{Data nodes} stores sequence files \& metadata of all sequence files in its filesystem
+        \item Both master node \& data nodes have user interface that allow users to upload, search, and download sequence files
+    \end{itemize}
+\end{frame}
+
+% DRAW.IO FILE: https://drive.google.com/file/d/1Nbp0Kmgcr2BZ13LBT3SGLMBoA6JA_Eth/view
+\begin{frame}{Theoretical Framework: Schematics}
+\textbf{App network diagram} \\ \medskip
+\centering
+\includegraphics[scale=0.5]{thesis1.png}
+\end{frame}
+
+\begin{frame}{Theoretical Framework: Schematics}
+\textbf{POST: Uploading a sequence to the master node} \\ \medskip
+\centering
+\includegraphics[scale=0.5]{thesis3.png}
+\end{frame}
+
+\begin{frame}{Theoretical Framework: Schematics}
+\textbf{POST: Uploading a sequence to a data node} \\ \medskip
+\centering
+\includegraphics[scale=0.5]{thesis2.png}
+\end{frame}
+
+\begin{frame}{Theoretical Framework: Schematics}
+\textbf{GET: Downloading a sequence} \\ \medskip
+\centering
+\includegraphics[scale=0.5]{thesis4.png}
+\end{frame}
+
+\subsection{Data Model}
+\begin{frame}{Data Model}
+This data model is lifted from SeqTorr \cite{seqtorr}
+
+\small
+\begin{center}
+
+\begin{tabular}{c|c}
+\hline
+    key & description \\
+\hline
+\hline
+    seq\textunderscore id & unique ID assigned to all sequences \\
+    organism & organism where the sequence was extracted from \\
+    quality & the rating of the data based on the number of downloads or another criteria \\
+    uploader & user that uploaded the data \\
+    institute & institute where the uploader or sequence was taken from \\
+    upload\textunderscore date & date the data was uploaded to the database \\
+    last\textunderscore modified & the date the metadata or data was edited \\
+    data\textunderscore nodes & list of data nodes that this data contains
+\end{tabular}
+\end{center}
+
+\end{frame}
+
+\subsection{Data Permissions}
+\begin{frame}{Data Permissions}
+\begin{center}
+\begin{tabular}{c|c}
+\hline
+    user & description \\
+\hline
+\hline
+   Moderator & Has ability to upload, download, store data, verify \\
+    Peer & Has ability to upload, download, store data \\
+    Guest & Has ability to upload, download \\
+
+\end{tabular}
+\end{center}
+
 \end{frame}

### 2020-06-24 Paper2020ACM/Presentation.tex
Revision June 24 630pm

@@
     \textbf{Problems of Classical Databases}
+    %PROBLEMS OF CLASSICAL DATABASES (ALL FOUR)
     \begin{enumerate}
-%PROBLEMS OF CLASSICAL DATABASES (ALL FOUR)
-    \item Genome researchers download entire database of genomic data
-    \item Genome researchers only need a portion of the database, but end up downloading everything  (gigabytes size)
-    \item Client-Server approach can face bottlenecks (server speed) %is resolved by making it more distributed
+        \item Genome researchers download an entire database of genomic data, especially when uncertain how to update their data
+        \item Genome researchers only need a portion of the database, but end up downloading the entire database
+        \item Client-Server approach can face bottlenecks (server speed)
+        \item P2P approach has a need of a large unmoderated active user base to work
     \end{enumerate}
-% or having a hard time getting a portion of the database
-% easier to download parts vs whole, whole vs parts
-
-% not easy to navigate the database? 
-% what is the big problem? too big data to download all at once?
-% why is it a problem? speed of download? or size?
-
     \textbf{Problems of Distributed Databases}
@@
     %PROBLEM OF DIST. DATABASES
-    \item Only one framework on distributed database for genomes
-    \item Existing distributed database for genome code has little documentation % concept of APIs and documentation
-
+        \item No existing usable public framework on distributed databases for genomes
+        \item Existing distributed database system for genomes is not replicable
     \end{enumerate}
@@
 \begin{itemize}
-    \item Create an information system that speeds up the download and upload of genomic data
+    \item Design the framework for an information system that speeds up the download and upload of genomic data
     \begin{itemize} 
         \item To resolve 1 \& 2 of the Problems of Classical Databases. 
+    \end{itemize}
+    \item  Design a distributed database system
+    \begin{itemize}
+        \item To resolve 3 \& 4 of the Problems of Classical Databases, and 1 \& 2 of the Problems of Distributed Databases. 
     \end{itemize}
@@
 \begin{itemize}
-    \item The system should have the main API’s
+    \item For the framework, the system should have the main API’s 
     \begin{itemize}
         \item POST genome data
-        \item GET genome data (implement hybrid P2P)
+        \item GET genome data (this will implement a hybrid P2P transfer)
         \begin{itemize}
-            \item This will aim to resolve number 3 in the Problems of Classical Databases
+            \item This will aim to speed up number 3 in the Problems of Classical Databases
         \end{itemize}
@@
         \item DELETE genome data
-    \end{itemize}
-    \item Write documentation for the system
-    \begin{itemize}
-        \item This will be used to resolve the Problems of Distributed Databases
     \end{itemize}
@@
 \begin{itemize}
-    \item Security and Authentication is not part of the scope
-    \item Use only FASTA file format 
+    \item To create a framework for a proposed system
+    \item To design the correctness of a system
+    \item To look at the basic time complexity of the base system (NCBI) and the proposed system
+    \item To compare the proposed system to the base system
+\end{itemize}
+
+The current scope and coverage of the project will NOT consider:
 \begin{itemize}
-        \item DNA nucleotides only
+    \item Security and Authentication 
+    \item Specific technologies to create the system
+    \item All the data stored in the system (but focus on main data only)
+    \item Implementing the actual system (but pseudocodes to the author's best ability will be provided)
 \end{itemize}
-    \item Only a proof of concept, UI will be purely for functional purposes (very basic design)
-\end{itemize}
+
 \end{frame}
@@
          \textbf{Networks Layers} how computers pass data to each other
-        \begin{itemize}
-            \item \begin{itemize}
+        \begin{enumerate}
                     \item Level 1: Medium used to send signals (physical, electrical signals of 0 and 1), through wire or by air
-            \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together different . Goal of sending a group of signals.
+                    \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together different interfaces. Goal of sending a group of signals.
                     \item Level 3: Introduction of networks: IP Address to uniquely identify entities.
                     \item Level 4: Management of error free data transmission: Sending a group of messages via TCP, UDP, FTP
-            \item \textbf{Level 5}: Abstraction of Level 4: Function calls programmers will use to do level 4 processes.
-            \end{itemize}
-        \end{itemize}
+                    \item Level 5: Abstraction of Level 4: Function calls programmers will use to do level 4 processes.
+        \end{enumerate}
         \end{frame}
@@
   \begin{itemize}   
-    \item system is this
-    \item distributed
-    \item community assisted
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Summary}
-  \begin{itemize}   
     \item 
@@
   \begin{itemize}   
-    \item good for community
-    \item easy to expand storage
-    \item 
+    \item Community may add their own data
+    \item Community may increase their data storage capacity
+    \item Community may increase data transfer speeds (through data nodes)
+    \item Frequently used data can become more accessible
   \end{itemize}
@@
 \begin{frame}{Speed Comparisons}
+
+Variables used:
   \begin{itemize}   
-    \item simple time complexity chart
-    \item faster si caDDS ofc 
-    \item 
+    \item \textbf{H} - Constant overhead before downloading may occur
+    \item \textbf{X} - Constant data transfer speed by the server in bytes per second
+    \item \textbf{N} - Number of users simultaneously using the system at the same time
   \end{itemize}
+
+Assumptions used:
+\begin{itemize}
+    \item constant data transfer
+    \item equal division of bandwidth
+    \item all users are getting the same data as once
+    \item there is a constant overhead before downloading the data for each user
+    \item there is also an equal division of the data nodes' bandwidth to the users
+    \item and all data nodes have an exact copy of the data.
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Speed Comparisons Algorithm}
+Algorithm for the base method:
+    \begin{enumerate}
+    \item There is a constant overhead H seconds to set up the communication between the user and the main server.
+    \item The main server sends data in the speed of X bytes per second. X is the speed of data transfer from the server.
+    \item Users will receive the data with an equal portion of the X bytes per second (bandwidth).
+    \end{enumerate}
+
+Algorithm for the proposed method:
+    \begin{enumerate}
+    \item There is a constant overhead H seconds to set up the communication between the user and the relevant nodes.
+    \item All data nodes (D number of data nodes) will send data in the speed of D times X bytes per second. X is the speed of data transfer from 1 data node.
+    \item Users will receive data with an equal portion of the D times X bytes per second.
+    \end{enumerate}
+\end{frame}
+
+\begin{frame}{Speed Comparisons Conclusions}
+
+Generally as the number of users increases, the data transfer speed decreases by a constant factor for all methods.
+There is also a concept of \textbf{bottleneck} when the number of users is much greater than the data speed put out by all methods.
+
+The proposed method (with D $>$ 1, and same H, X, N) will be faster than the base method.
+\textit{With this analysis, it is suggested the proposed method is faster than the base method on most cases (D $>$ 1).
+}
 \end{frame}
@@
 \begin{frame}{Final Comparisons}
+  \begin{block} {\textbf{Base Method: NCBI}}
     \begin{itemize}
-    \item caDDS is scalable, easy for community, has docu hehe
-    \item NCBI is one, stable, widely used
-    \item 
+        \item The current base method (NCBI) is the go-to website for biological data and genomic data. 
+        \item One website and one server. This points to a single point of failure\cite{seqtorr}.
+        \item The section on theoretical analysis of speed described the problem of having a bottleneck when the number of users are too many for the data bandwidth of the server. 
     \end{itemize}
+
+  \end{block}
+
+    \begin{block} {\textbf{Proposed Method: caDDS}}
+
+    \begin{itemize}
+        \item With multiple data nodes, this is faster in theory than the base method.
+        \item  Community: \begin{itemize}
+            \item Database only contains community used data
+            \item Frequently used data can be made more accessible
+            \item The design and principles of the database may be calibrated to the needs of a community
+        \end{itemize}
+    \end{itemize}
+  \end{block}
 \end{frame}
@@
   \begin{itemize}   
-    \item implement
-    \item security
-    \item public code
+    \item Implement the proposed system
+    \item Pay attention to port problems with university internet
+    \item Discuss and mitigate different security risks
+    \item Find a way to connect master nodes to other master nodes
+    \item Abstract the data to non-genomic data, for use in other scientific fields
   \end{itemize}

### 2020-06-24 Paper2020ACM/Paper2020.tex
Revision June 24 630pm
Revision June 24 1030pm

@@
     \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together
-different . Goal of sending a group of signals.
+different interfaces. Goal of sending a group of signals.
     \item Level 3: Introduction of networks: IP Address to uniquely identify
@@
 
-The database is a distributed database, which will aim to utilize the best features from Peer2Peer and Client/Servers.
+The database is a distributed database, which will aim to utilize the best features from peer-to-peer and client-server architectures.
 This will also aim to store sequences.
@@
 This means all users in the community can contribute to the knowledge base, and contribute data to be used by the community.
-Another important community aspect is that users can become data nodes themselves.
-Meaning these users can increase the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later).
+Another important community aspect is that institutions can increase the number of data nodes.
+Meaning there will be an increase in the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later).
 Frequently used local data as well will be more easily accessible (since it's on local data nodes, and should have faster access speeds).
@@
 For the final comparisons of the base method the proposed method, the current base method (NCBI in Section \ref{NCBI}) is the go-to website for biological data and genomic data.
-It is currently used by and is one website and one server.
+It is currently used and is one website and one server.
 This points to a single point of failure\cite{seqtorr}.
@@
 
-\section{User-System Interface}
-(check appendix for detailed pseudocode)
-
-parang this section abstracts the user part from the system part
-
-The user commmunicates with the system. User can do create, read, update, delete.
-
-The system on its own delegates how things are stored or updated consistently ...
-
-The user communicates with the master all the time???
-
-\section{Master Node-Data Node Interface}
-(check appendix for detailed pseudocode)
-
-parang this section abstracts how master and data communicate and store data
-
-GET
-The user requests data from the system (via the master gate hehe)
-
-The master holds a copy of all the data locations in the data nodes (can be called a directory). 
-The master once needs a certain data, will check its directory which data nodes have it
-The master once knows which of its data nodes have the data (multiple) will call them to inform it will request data
-The data is informed that data will be requests
-The master sends the directory of addresses of data nodes with the data and transfer key???? to the user
-
-The user user receives the transfer key and uses it in the website to request data from data nodes
-
-CREATE
-The user wants to send data to the system (via the master gate hehe)
-
-The master holds a copy of all the data locations (can be called a directory). 
-The master will find 2+ data nodes to hold the data, and store them in a temp directory
-The master will call the 2+ data nodes and check if they have enough space and are able to receive the data
-The master will wait for an ack before sending the data to the 2+ data nodes
-The master will ask the user to generate a hashcode
-The master will send the user data nodes to send the data
-The data nodes will receive the data and hash
-The data nodes will complete receiving data, generate a hash and compare it with the user's hash. If different, repeat the process (may ask from a data node with complete data).
-The data nodes if all have the complete, correct data will inform master node.
-
-The user user receives the transfer key 
-The user generates a hash of its data
-The user uses the transfer key to send data to data nodes
-
-\section{Master Node-Data Node Structure}
-(check appendix for detailed pseudocode)
-
-MASTER NODE
-holds data:
-directory of data nodes with verified data
-directory of data nodes with unverified data (being transferred currently)
-directory of data nodes space free
-
-holds function:
-which data nodes have data x?
-does data node N have data x?
-is user able to receive data X?
-is user able to send data X?
-is data node N able to send data X?
-is data node N able to receive data X?
-
-what is hash of data X on data node N?
-what is hash of user data X to be sent?
-Is hash of data X on data node N and user data X same?
-
-is data X being transferred currently from data node N?
-is data X being transferred currently to data node N?
-is data X done transferring to data node N?
-is data X across all data nodes consistent? (same hash)?
-
-how much space does data node N have?
-how much space does user data to send need?
-
-send data node N location to user?
-where is the location of data node N?
-
-DATA NODES
-holds data:
-Metadata
-Data
-Directory of Metadata
-Storage Space
-
-holds function:
-has storage space X?
-how much storage space?
-is data X across all data nodes consistent? (same hash)?
-how much space does user data need?
-able to receive data X?
-able to send data X?
-what is hash of data X?
-what is hash of user data X to be sent?
-Is hash of data X and user data X same?
-
-\section{Node Internal Interface}
-(check appendix for detailed pseudocode)
-Frontend-Backend Interface
-Via Data APIs
-GET, POST, PUT, DELETE
-
-Via Data Sync APIs
-DATA SAME?
-
-Via Data Availability APIs
-DATA AVAILABLE?
-DATA NODE ABLE TO RECEIVE?
-DATA NODE ABLE TO SEND?
-
-\section{Node Internal Structure}
-(check appendix for detailed pseudocode)
-
-parang this section abstracts how nodes function inside
-
-Frontend - debugging mostly, users dont see this frontend???? or isang frontend lang ni master nakikita
-Backend - storage before it goes to database? for functions? for APIs??
-
-Inside Backend: Idea pala to categorize functions into these APIs idk if tama paggamit ko ng APIs
-Network APIs - abstracted part to connect to the network
-Database Sync APIs - abstracted part to keep database consistent
-Database Availability APIs - abstracted part to keep database available
-
-Database - stores shit
-
-\subsection{Community Aspect}
-- good for community since the system is expandable or customizable??
-- community can see the whole framework and structure
-- communities can set up their own master and data nodes
-- individuals in communities can host their own data nodes (ideally master node is the institution with the data)
-- ideally having multiple data nodes makes it download faster somehow
-- communities can only store the data they need
-- communities basically can self-expand storage, self-organize somehow the storage, and self-increase their bandwidth
-
 \end{document}

### 2020-06-25 Paper2020ACM/Paper2020.tex
Revision June 25 1am

@@
     \midrule
-   Moderator & Has ability to upload, download, store data, verify \\
-    Peer & Has ability to upload, download, store data \\
-    Guest & Has ability to upload, download \\
+   Moderator & Has ability to upload, download, verify \\
+    Registered User & Has ability to upload, download \\
+    Guest & Has ability to download \\
    \bottomrule 
@@
 This is to make sure there is a means to verifying the data being uploaded.
-All users can upload, download, and store (or host) data.
-All other users, which are called guests has an ability to download and upload data.
+All registered users can upload and download.
+All other users, which are called guests, have an ability to download data.
 

### 2020-06-25 Paper2020ACM/Paper2020.tex
Revision June 25 5:30pm

@@
 The problem faced by genome researchers is the large data set and difficulty transferring the files.
-Previous distributed databases are either not meant for genomic data, or difficult to replicate.
+Previous distributed databases are either not meant for genome data, or difficult to replicate.
 This paper lays the groundwork for a distributed database that is designed to easily scale to accommodate bigger storage, and also reduces the data speed bottleneck faced by centralized databases.
 The proposed system has a master node and data nodes.
-Data nodes should theoretically increase data transfer speeds even as the number of users increase.
-(to be updated based on conclusion) While current databases(e.g.
+Master node handles administrative work while the Data nods contain all sequence data.
+Data nodes should theoretically increase data transfer speeds.
+While current databases(e.g.
 NCBI) have more data, more users, and a live platform.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% START OF PROGRESS FOR CID MAY 22 2020
 \section{Introduction}
@@
 
-Defining \textit{genome}
-% CID PROGRESS ENDS HEREE FOR MAY 22 2020
-%------------------------------------------
-% CID PROGRESS STARTS HEREE FOR MAY 29 2020
-
 \subsection{Definition of Genomics}
-Genomics is is the study of wholes sets of genes and their interactions\cite{campbell}.
+Genomics is is the study of wholes sets of genes and their interactions\cite[~p.437]{campbell}.
 The efforts to study this field has led to volumes of data that need to be analyzed.
@@
 Institutions such as the National Institute of Health and National Library of Medicine, have led this field and built a website to encourage collaboration online through their online portal called NCBI (National Center for Biotechnology Information).
-\cite{campbell} This website will be discussed later in more detail.
+\cite[~p.438]{campbell} This website will be discussed later in more detail in Section \ref{NCBI}.
 
@@
 DNA's structure is composed of chemical compounds (or nucleotides) that encode information.
-These nucleotides are commonly written down as A,G,C,T corresponding to their chemical makeup.
+These nucleotides are commonly written down as A,G,C,T corresponding to their chemical makeup\cite{alberts_mole}.
 Another important feature of genes is that these genes are the instructions passed on from one generation to another.
@@
 Genomics play a role in understanding different components of an organism by paying attention to the genes and nucleotide bases.
-Different organisms have certain differences in genes and this has been a focal point of many genomic studies, which sometimes is aimed as well as establishing surprising similarities among distinct organisms.
+Different organisms have certain differences in genes and this has been a focal point of many genomics studies, which sometimes is aimed as well as establishing surprising similarities among distinct organisms.
 
@@
 This also means more complex organisms are likely to have more complicated genes and therefore have a lot more information.
-Even if this general idea is true, there are some exceptions, some plants such as the Japanese canopy plant \textit{Paris japonica} have 50 times the genome size compared to humans\cite{campbell}.
+Even if this general idea is true, there are some exceptions, some plants such as the Japanese canopy plant \textit{Paris japonica} have 50 times the genome size compared to humans\cite[~p.442]{campbell}.
 These may be due to repeating non-essential DNA.
@@
 These paved the way for more and more whole genome studies of animals much different from humans.
-By April 2013, around 4,300 genomes had been studied\cite{campbell}.
+By April 2013, around 4,300 genomes had been studied\cite[~p.442]{campbell}.
 These contribute to the growing amount of research, analysis, and most importantly data around genomes of different organisms.
@@
 These are other aspects needed by a genomics researcher as well.
-And this is the reason why a specific class of Databases called \textit{Distributed Databases} will be needed in this paper.
+And this is the reason why a specific class of Databases called \textit{Distributed Databases} will be discussed in section \ref{db_types}.
 
@@
 These form a system that is made for accessing and updating data.
-The next section discusses the key points when considering how to  create a good database system (or DBMS).
-The key points are important to discuss many important features and dilemmas of the proposed system later.
+The next section (\ref{cap}) discusses the key points to consider in creating a good database system (or DBMS).
+The key points are important to discuss many important features and dilemmas of the proposed system.
 Certain database principles are needed to make sure the data is available, secure, and consistent.
 
-\subsubsection{Database Types} 
+\subsubsection{Database Types} \label{db_types}
 Two types of databases will be discussed in this paper.
@@
 One is more efficient and easier to make, another is better suited for collaborative work.
-Note some technical jargon needs to be used to explain some concepts.
+Note the technical jargon that needs to be used to explain some concepts.
 These are namely: \cite{centralizedvsdistributed}
@@
 \subsubsection{CAP theorem for Databases} \label{cap}
-There are 3 principles needed when creating a database.
+There are 3 principles needed when creating a database\cite{Silberschatz2010}.
+It is given that it is impossible to do all 3 principles at the same time perfectly\cite{gilbert2012perspectives}.
 The idea is that only two of the three can be focused on at any given time.
-Only two since the current technology can only handle two.
 These principles and limitations need to be taken into consideration when designing the system.
@@
 
-\subsubsection{Client-Server}
+\subsubsection{Client-Server} \label{client-server}
 A client-server architecture relies on a Web server that is always available and has a fixed IP address.
@@
 
-\subsubsection{P2P}
+\subsubsection{P2P} \label{p2p}
 A peer-to-peer (P2P) architecture has minimal to no reliance on always-on, dedicated servers.
@@
 There was also a focus on being able to choose information to share and use.
-The standard genomics workflow consists of uploading and downloading data on an international database like NCBI.
+The standard genomics workflow consists of uploading and downloading data on an international database like NCBI in Section \ref{NCBI}.
 Having a distributed scalable local infrastructure to store genomic data instead of relying on NCBI would be beneficial since: only certain sequences in NCBI are relevant to the institutions (e.g.
@@
 BioTorrents is a system for legally sharing scientific data that works on top of the BitTorrent protocol (or P2P).
-It is much like a public tracker.
+It is much like a public tracker \cite{biotorrents}.
 The main website hosts .torrent and metadata files.
@@
 % CID PROGRESS ENDS HEREE FOR MAY 29 2020
+
+\subsection{Dat Protocol}
+
+The Dat Protocol is a newly-developed, open source peer-to-peer protocol for sharing data\cite{dat}.
+With a single command, anyone can start hosting a directory of files for others to download, as long as they have a publicly accessible IP address and port.
+Other users can download and then host their own copy of the same directory.
+Downloads are distributed, meaning if a file is shared across multiple active peers, a new user will be downloading that file from those peers, so bandwidth is shared, much like in BitTorrent.
+A file's data is synced across peers, so any changes made to the original copy of the file will be sent to peers hosting the same file.
+And like Git, the version history is also stored and made accessible to peers.
+Unlike Git though, Dat works well for any kind of file, not just source code.
+Dat also has security built in.
+Only users with the directory's read key can download it, and only users with the write key can make changes.
+In short, Dat has "taken the best parts of Git, BitTorrent, and Dropbox".
+Another important feature is that files and directories are stored as-is, meaning no modifications are done by Dat when storing files.
+
+Dat's development began in 2013, so it is a relatively new protocol.
+There may still be large changes in the development and management of Dat, so the project is relatively volatile.
+Development is ongoing, and the community is active, yet there may still exist a few bugs and other difficulties when using Dat to build an application.
+Dat is built on NodeJS, and support for other programming languages and frameworks is quite limited.
+It also contains the same problems as other P2P protocols, namely, a lack of data moderation and incentives for users to serve as peers.
+
 % CID PROGRESS STARTS HEREE FOR JUNE 2 2020
@@
 \begin{table}[h]
-    \caption{Comparisons of Different Databases}
+    \caption{Comparisons of Different Databases and Technologies}
     \label{table:database_comparison_table}
@@
     NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\
+    Dat Protocol    & P2P      & Yes        & Depends           & General  \\ 
     \bottomrule
@@
 
-Table~\ref{table:database_comparison_table} on page ~\pageref{table:database_comparison_table} describes the different technologies studied.
+\subsection{Comparisons of Database Technologies} \label{compdb}
+
+Table~\ref{table:database_comparison_table} on page ~\pageref{table:database_comparison_table} compares the different technologies studied.
+An important point to note is that a majority of the databases are not available.
 This will also be used later on to describe what are the main problems of the different databases.
@@
 \begin{enumerate}
-\item Genome researchers download an entire database of genomic data, especially when uncertain how to update their data
+\item Genome researchers download an entire database of genomics data, especially when uncertain how to update their data
 \item Genome researchers only need a portion of the database, but end up downloading the entire database
-\item Client-Server approach can face bottlenecks (server speed)
-\item P2P approach has a need of a large unmoderated active user base to work
+\item Client-Server approach can face bottlenecks (server speed) (Section \ref{client-server})
+\item P2P approach has a need of a large unmoderated active user base to work (Section \ref{p2p})
 
@@
 \begin{enumerate}
-    \item No existing usable public framework on distributed databases for genomes
+    \item No existing usable public framework on distributed databases for genomes (Section \ref{compdb})
     \item Existing distributed database system for genomes is not replicable 
@@
 \begin{itemize}
-    \item Design the framework for an information system that speeds up the download and upload of genomic data
+    \item Design the framework for an information system that speeds up the download and upload of genomics data
     \begin{itemize} 
@@
     \item All the data stored in the system (but focus on main data only)
-    \item Implementing the actual system (but pseudocodes to the author's best ability will be provided)
+    \item Implementing the actual system
 \end{itemize}
@@
 and 1), through wire or by air
-    \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together
+    \item Level 2: Introduction of packets: WIFI, Ethernet, connecting together
 different interfaces. Goal of sending a group of signals.
@@
     \item Level 4: Management of error free data transmission: Sending a group
-of messages via TCP, UDP, FTP
+of messages via TCP, UDP, FTP (File Transfer Protocols)
     \item Level 5: Abstraction of Level 4: Function calls programmers will use
@@
 
-% CID PROGRESS STARTS HEREE FOR JUNE 6 2020
+\subsection{Summary of the System}
 
-\subsection{Summary of the System}
+\subsubsection{Overview}
+The two parts of the system infrastructure are the master node and data nodes. These are also inspired from SeqTorr.
+
+The master node is a server that facilitates all authentication, uploading, and moderation.
+It also catalogs the metadata of all files.
+It will have a web application on top, with front-end and back-end that users wanting to upload a sequence to the system will have to access.
+The master node is also in charge of distributing sequence data to the data nodes.
+
+The data nodes are servers that contain the sequence files themselves.
+Users will be downloading the sequences from these nodes.
+These nodes will also support the Dat protocol or a similar P2P network to enable instant syncing of data between data nodes and distributed download.
+
+One institute will host the master node and be responsible for all system administration.
+This institute may also host a data node.
+Partner institutes will have their own data node/s that will share in the storage and distribution of sequences.
+Data nodes provide redundancy.
+The same sequence will be uploaded to multiple data nodes, so when a single data node becomes unavailable for whatever reason, no information is actually lost because copies have been made in other nodes.
+Data nodes also have the feature of enabling distributed download.
+Utilizing the P2P functionality provided by Dat, a researcher can download from multiple data nodes simultaneously, easing bandwidth requirements and speeding up transmission rate.
 
@@
 When a user uploads data, the moderator will verify the data's authenticity before allowing it to be uploaded.
+The master node will then contact data nodes and see which are available to host.
 
 Then for the downloading of the data.
-The user will download normally, but under the hood the system will be doing a lot of work.
-It will check multiple "mirrors" of the download, by looking at the different available data nodes, then let the user choose the fastest or best option.
-File transfer will occur to bring the file to the user, then the system will be the one checking if the file were downloaded correctly.
-There is a simple organization of available files as well.
+In the master node, the user will search for a sequence by keywords from its metadata.
+The user is then informed which currently available data nodes host the wanted sequence.
+The user will download normally from a single host using HTTP, or they can choose to download from multiple nodes simultaneously, with the P2P protocol described above.
+
+% END OF ROUGH DRAFTS
 
@@
 \subsubsection{Community Aspect} \label{community}
-The sytem was also described as community assisted.
-The definition of \textbf{community} to be used here is a \textit{set of people who generally collaborate together to do research}.
+The system was also described as community assisted.
+The definition of \textbf{community} to be used here is a \textit{set of people and institutions who generally collaborate together to do research}.
 And the focus is the genome research community.
-Reemphasizing certain features from the database, that any users can upload data.
+Reemphasizing certain features from the database, that any registered user can upload data.
 This means all users in the community can contribute to the knowledge base, and contribute data to be used by the community.
@@
 Meaning there will be an increase in the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later).
-Frequently used local data as well will be more easily accessible (since it's on local data nodes, and should have faster access speeds).
+Frequently used local data as well will be more easily accessible (since data nodes should have faster access speeds).
 The concept of closer networks having faster speeds (discussed in the network section) is at play here.
@@
 \includegraphics[width=0.35\textwidth]{images/thesis3.png} 
-\end{figure}
-
-\begin{figure}[h]
-\caption{POST: Uploading a sequence to the data node}
-\centering
-\includegraphics[width=0.35\textwidth]{images/thesis2.png} 
 \end{figure}
@@
 \section{Proof of Correctness}
-Hii i need help fixing this section di ko pa alam ano ilalagay
-
 The distributed database will function correctly if:
@@
 
+The system will function correctly if it fulfills the main APIs without error, and that it is able to be partition-tolerant.
+Meaning the data is synchronized between the nodes.
 A proof of correctness for the implementation.
@@
 \begin{itemize} 
-    \item 1 total users (and 1 data node) will receive data in the speed of 1 times X/1 bytes per second. Downloading data (size of X bytes) will take X/X + H seconds to transfer.
+    \item 1 total user (and 1 data node) will receive data in the speed of 1 times X/1 bytes per second. Downloading data (size of X bytes) will take X/X + H seconds to transfer.
     \item 2 total users (and 1 data node) will receive data in the speed of 1 times X/2 bytes per second. Downloading data (size of X bytes) will take X/(2X) + H seconds to transfer.
@@
 Another advantage of having a distributed database is the community aspect (see Section \ref{community}).
-The may be implemented on a local scale, meaning it can have the necessary data for the community.
+The proposed method may be implemented on a local scale, meaning it can have the necessary data for the community.
 The most frequently used data, may be put on more data nodes increasing the speed of the data transfer (see Section \ref{theoreticalspeed}).
@@
 
-Another is to look into security, which aspects of the code or data must be private or moderated.
+Another is to look into security.
 How does data privacy work?
@@
 Many questions arise in security but these will only be answered better once implementation starts.
-
-A recommendation is to look into a way for master nodes to communicate with other master nodes, this will start the idea of communication and collaboration between caDDS to other caDDS.
 
@@
 
-We want to acknowledge our advisers for giving the idea for our thesis.
-And for answering our many questions about genomics.
+We want to acknowledge our advisers Dr.
+Jan Michael Yap and Sir El King Morado for giving the idea for our thesis.
+They have answered our many questions about genomics.
 Writing this paper overlapped with the covid19 pandemic which changed a big portion of the scope of this thesis.
-We thank our advisers for making this work manageable during this pandemic crisis.
+We have unexpressable gratitude to our advisers for being understanding and trusting with this.
+We also thank the people from Philipping Genome Center who welcomed us and gave us coffee and a workstation.
 \end{acks}

### 2020-06-25 Paper2020ACM/Presentation.tex
Revision June 24 1030pm
Revision June 25 5:30pm

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO + RRL %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{Abstract}
-Genomic data the past few years have grown exponentially with the technology barely keeping up.
-The problem faced by genome researchers is the large data set and difficulty transferring the files.
-Previous distributed databases are either not meant for genomic data, or difficult to replicate.
-This paper lays the groundwork for a distributed database that is designed to easily scale to accommodate bigger storage, and also reduces the data speed bottleneck faced by centralized databases.
-The proposed system has a master node and data nodes.
-Data nodes should theoretically increase data transfer speeds even as the number of users increase.
-(to be updated based on conclusion) While current databases(e.g.
-NCBI) have more data, more users, and a live platform.
-The proposed system is aimed towards a smaller community, with more frequent and localized data, which the data storage can easily be expanded as the need arises.
-The proposed system is also theoretically faster as long as there are multiple working data nodes.
-A recommendation is to implement the system and make the code public for the use of other communities.
-\end{frame}
+
 \section{Introduction}
-\subsection{Genomic Concepts}
+\subsection{Genomics Concepts}
 
@@
     \item Every read stores the many fragments
-    \item Each sequence read is redundant and uses a lot of space
+    \item Each sequence read has redundant information and uses a lot of space
   \end{itemize}
@@
     \item Prediction of the speed new technology (faster processing power and better storage) is developed
-    \item But moore's law is too slow for the speed of genomic data increase
+    \item But moore's law is too slow for the speed of genomics data increase
     \item
@@
   \begin{itemize}   
-    \item Each read\cite{seqtorr} has 60x coverage, around 200gb each
+    \item One sample read\cite{seqtorr} has 60x coverage, around 200gb each
     \item NGS, More genomic researches, more files needed to be saved and transferred
@@
   \begin{itemize}   
-    \item store data with tech
+    \item Database is a technology used to store and manage data
   \end{itemize}
@@
   \begin{frame}{Data Storage}
-        A \textbf{database} (\textbf{DBMS}, Database Management System) is defined by 
+        More formally a \textbf{database} (\textbf{DBMS}, Database Management System) is defined by 
         \begin{block} {\textbf{Definition}}
@@
         \begin{itemize}
-            \item \textbf{Principles} to remember when creating a database (CAP). Idea is that only 2 of the three can be focused on. \cite[Ch.~19]{Silberschatz2010}
+            \item \textbf{Principles} to remember when creating a database (CAP).
+Note that it is impossible to achieve perfectly all 3 principles.
+\cite{cap}Idea is that only 2 of the three can be focused on.
+\cite[Ch.~19]{Silberschatz2010}
             \begin{itemize}
@@
             \item NCBI has a large collection of genomic datasets in its website, which can be searched and downloaded.
+            \item Single source and a single point of failure
         \end{itemize}
@@
     \end{frame}
-
-    % cinomment out ko kasi meron naman explanation sa taas
-    %\subsection{P2P Database}
-    %\begin{frame}{P2P Databases}
-    %    something explaining briefly p2p databases
-    %\end{frame}
 
@@
 
+    \subsection{Dat Protocol}
+    \begin{frame}{Distributed: Dat Protocol}
+        \begin{itemize}
+            \item New open source protocol (not a database) developed in 2013, still volatile
+            \item Users can download and host files
+            \item Downloads are distributed among users
+            \item Version history exists
+            \item Works for any kind of file
+        \end{itemize}
+    \end{frame}
+
     \subsection{SeqTorr}
@@
     \begin{frame}{Distributed: SeqTorr}
-    % have a unified way of discussing each database
-    % what aspects of the database to discuss each
-    % extra features will come after
-        Problems: % have one for each database system
+        Problems: 
         \begin{itemize}
-            \item program not currently available
+            \item System not currently available
             \begin{itemize}
-                \item cannot measure any statistics on this
+                \item cOnly exists documentation of their APIs
             \end{itemize}
@@
     \begin{frame}{Table}
-% Bio-related -> biological
-% What's the difference between biological vs general vs FASTA
-% What does general mean?
-% What does biological mean? -> explain pa to
-% May 3 types of data ...
-   \begin{table}[]
-        \begin{tabular}{|l|l|l|l|l|}
-        \hline
-        System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ \hline
-        BioTorrents & P2P           & No         & Depends                    & Biological  \\ \hline
-        SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ \hline
-        PeerDB      & P2P           & No         & Depends                    & General      \\ \hline
-        NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\ \hline
+
+\begin{table}[h]
+    \caption{Comparisons of Different Databases and Technologies}
+    \label{table:database_comparison_table}
+    \begin{tabular}{ccccc}
+    \toprule
+    System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ 
+    \midrule
+    BioTorrents & P2P           & No         & Depends                    & Biological  \\ 
+    SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ 
+    PeerDB      & P2P           & No         & Depends                    & General      \\ 
+    NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\
+    Dat Protocol    & P2P      & Yes        & Depends           & General  \\ 
+    \bottomrule
     \end{tabular}
@@
     \begin{enumerate}
-        \item Genome researchers download an entire database of genomic data, especially when uncertain how to update their data
+        \item Genome researchers download an entire database of genomics data, especially when uncertain how to update their data
         \item Genome researchers only need a portion of the database, but end up downloading the entire database
-        \item Client-Server approach can face bottlenecks (server speed)
-        \item P2P approach has a need of a large unmoderated active user base to work
+        \item Client-Server approach can face bottlenecks (server speed) (Section \ref{client-server})
+        \item P2P approach has a need of a large unmoderated active user base to work (Section \ref{p2p})
     \end{enumerate}
@@
 
+\begin{frame}{Abstract}
+Genomic data the past few years have grown exponentially with the technology barely keeping up.
+The problem faced by genome researchers is the large data set and difficulty transferring the files.
+Previous distributed databases are either not meant for genomic data, or difficult to replicate.
+This paper lays the groundwork for a distributed database that is designed to easily scale to accommodate bigger storage, and also reduces the data speed bottleneck faced by centralized databases.
+The proposed system has a master node and data nodes.
+Data nodes should theoretically increase data transfer speeds even as the number of users increase.
+(to be updated based on conclusion) While current databases(e.g.
+NCBI) have more data, more users, and a live platform.
+The proposed system is aimed towards a smaller community, with more frequent and localized data, which the data storage can easily be expanded as the need arises.
+The proposed system is also theoretically faster as long as there are multiple working data nodes.
+A recommendation is to implement the system and make the code public for the use of other communities.
+\end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OBJECTIVES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \item All the data stored in the system (but focus on main data only)
-    \item Implementing the actual system (but pseudocodes to the author's best ability will be provided)
+    \item Implementing the actual system 
 \end{itemize}
@@
                     \item Level 1: Medium used to send signals (physical, electrical signals of 0 and 1), through wire or by air
-                    \item Level 2: Introduction of packets: Wifi, Ethernet, connecting together different interfaces. Goal of sending a group of signals.
+                    \item Level 2: Introduction of packets: WIFI, Ethernet, connecting together different interfaces. Goal of sending a group of signals.
                     \item Level 3: Introduction of networks: IP Address to uniquely identify entities.
@@
     \subsection{Application Architecture}
-    % part ng theoretical framework
-    % pwedeng walang slide of c/s
-    % pwedeng ibanggit nalang ang client-server architecture
 
-    % methods: ito framework
-    % implemens client-server...
     \begin{frame}{Application Architecture}
@@
 \begin{center}
-
-\begin{tabular}{c|c}
-\hline
+\begin{tabular}{cl}
+    \toprule
     key & description \\
-\hline
-\hline
+    \midrule
     seq\textunderscore id & unique ID assigned to all sequences \\
@@
     last\textunderscore modified & the date the metadata or data was edited \\
-    data\textunderscore nodes & list of data nodes that this data contains
+    data\textunderscore nodes & list of data nodes that this data is contained to \\
+    \bottomrule
 \end{tabular}
@@
 \begin{center}
-\begin{tabular}{c|c}
-\hline
+\begin{tabular}{cl}
+    \toprule
     user & description \\
-\hline
-\hline
-   Moderator & Has ability to upload, download, store data, verify \\
-    Peer & Has ability to upload, download, store data \\
-    Guest & Has ability to upload, download \\
-
+    \midrule
+   Moderator & Has ability to upload, download, verify \\
+    Registered User & Has ability to upload, download \\
+    Guest & Has ability to download \\
+   \bottomrule 
 \end{tabular}
@@
     \item Community may add their own data
-    \item Community may increase their data storage capacity
+    \item Community may increase their data storage capacity (via adding institution data nodes)
     \item Community may increase data transfer speeds (through data nodes)
@@
   \begin{itemize}   
-    \item Implement the proposed system
+    \item Implement the proposed system and make the code public
     \item Pay attention to port problems with university internet
     \item Discuss and mitigate different security risks
-    \item Find a way to connect master nodes to other master nodes
     \item Abstract the data to non-genomic data, for use in other scientific fields

### 2020-06-25 Paper2020ACM/Presentation.tex
Thesis submission June 25 6pm

@@
             \begin{itemize}
-                \item cOnly exists documentation of their APIs
+                \item Only exists documentation of their APIs
             \end{itemize}
@@
         \item Genome researchers only need a portion of the database, but end up downloading the entire database
-        \item Client-Server approach can face bottlenecks (server speed) (Section \ref{client-server})
-        \item P2P approach has a need of a large unmoderated active user base to work (Section \ref{p2p})
+        \item Client-Server approach can face bottlenecks (server speed)
+        \item P2P approach has a need of a large unmoderated active user base to work
     \end{enumerate}
@@
     \begin{itemize}
-        \item A \textbf{master node} stores metadata of all sequences, may or may not contain sequence files
+        \item The system is inspired by SeqTorr's design
+        \item A \textbf{Master node} does administrative functions: authentication, uploading, moderation. It also contains all metadata.
         \item \textbf{Data nodes} stores sequence files \& metadata of all sequence files in its filesystem
-        \item Both master node \& data nodes have user interface that allow users to upload, search, and download sequence files
+    \end{itemize}
+\end{frame}
+\begin{frame}{Theoretical Framework: System Description}
+    When a user downloads a sequence:
+    \begin{itemize}
+        \item The user searches for keywords in the system
+        \item The user is then given the results and the available data nodes
+        \item The user then chooses to download from one data node or from multiple data nodes
     \end{itemize}
@@
 \centering
-\includegraphics[scale=0.5]{thesis1.png}
+\includegraphics[scale=0.5]{images/thesis1.png}
 \end{frame}
 
-\begin{frame}{Theoretical Framework: Schematics}
-\textbf{POST: Uploading a sequence to the master node} \\ \medskip
+\begin{frame}{Theoretical Framework: POST}
+\textbf{POST: Uploading a sequence} \\ \medskip
 \centering
-\includegraphics[scale=0.5]{thesis3.png}
+\includegraphics[scale=0.50]{images/thesis1-Page-3.png} 
 \end{frame}
 
-\begin{frame}{Theoretical Framework: Schematics}
-\textbf{POST: Uploading a sequence to a data node} \\ \medskip
+\begin{frame}{Theoretical Framework: GET}
+\textbf{GET: Downloading a sequence} \\ \medskip
 \centering
-\includegraphics[scale=0.5]{thesis2.png}
+\includegraphics[scale=0.5]{images/thesis4.png}
 \end{frame}
 
-\begin{frame}{Theoretical Framework: Schematics}
-\textbf{GET: Downloading a sequence} \\ \medskip
+\begin{frame}{Theoretical Framework: PUT}
+\textbf{PUT: Updating a sequence} \\ \medskip
 \centering
-\includegraphics[scale=0.5]{thesis4.png}
+\includegraphics[scale=0.5]{images/thesis1-Page-5.png}
+\end{frame}
+
+\begin{frame}{Theoretical Framework: DELETE}
+\textbf{DELETE: Deleting a sequence} \\ \medskip
+\centering
+\includegraphics[scale=0.5]{images/thesis1-zzzzz.png}
 \end{frame}
@@
 
-\end{frame}
-
-\subsection{System Summary}
-\begin{frame}{System}
-  \begin{itemize}   
-    \item 
-    \item 
-    \item 
-  \end{itemize}
 \end{frame}

### 2020-06-26 Paper2020ACM/MeetingJune25.tex
June 25 9pm meeting notes

Marami to hahahaha get ready mostly sa proof of correctness

Whole Paper:
1) Make it 2 column format
2) Add sir El King Morado - UP Philippine Genome Center
3) Remove references to page numbers
4) Put figures on top of a column (see what sir el king sent)

Abstract
- okay content
- fix typoL Data nods

Intro Pg 1
- Remove sentence: [the goal of this intro]
- remove subsections in inntro
- Genomics study of blank. To generate data you need sequencing. But the data is big & needs effort. To sequence a lot, tech evolved and NGS gave high throughput
- Intro should be 2 columns lang. Make each paragraph a few sentences instead.
- Intro goal: layoutht background of the work
- Flow: Idea of A lot of data -> Problems -> Proposed Solution
-

-Intro Pg 2
- 3,000,000,000 -> 3 billion
- Shorten HGP section. Combine HGP + NGS to one paragraph.
- Fix typo: Genomics is is

-Intro: Database
- NGS -> high voluminous data -> need to store it in a digital format -> retrieve it for analysis

Overall info guide: 
- tama na yung flow
- lessen info
- compress database to 1/2 sentences. Then mention centralized/distributed database.

Into: CAP
- make it 1 paragraph
- only describe the most important principle and why

Application Architecture: (Move to Theoretical Framework)
- Discuss Client Server in 1 sentence
- Discuss P2P in 1 sentence
-Describe Hybrid

-Sample: 
- First we need to discuss 2 basic architecture: CS, P2P. 
- CS is blank. 
- P2P is blank. 
- Our proposed system will be hybrid getting these aspects from CS and P2P

RRL:
- first paragraph say: We review blank. We would like to focus on blank and blank. In the end we would like to make a comparitive analysis. 
- Table Comparing the diff Databass: Remove Dat, Describe each column narratively, and describe some conclusions from the table
- Answer the question: anong halaga ng columns? 
- Discuss each column CS/P2P/H, and the data on each column
- FASTA -> Biological
- Move dat to Methodology.

Review of Related Literature:
- the review tells you you are meant to critique the work you have seen
- here summarize gaps (mej copy the problems to here)

Problem:
- summarize the gaps. To review the current implementation has problems in size.
- To summarize, centralized databases suffer from these problems. Therefore the research aims to focus on developing a database with these features
- Problem Statement: Central thesis, one unifying question. May certain level of measurability.

Objectives:
- specify metrics, activities, na once you do this succeed.
- the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
- For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
- keep first 2 objectives, the last is: to evaluate the theoretical properties of the system based on the correctness and its speed

Methodology:
- dito mapupunta yung framework and APIs
- say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.

Scope
- 2 things: 
- 1) create a theoretical design for a proposed system
- 2) evaluate the theoretical running time and speed of the operations.
- define here what you mean by proposed system and base system (FTP protocol of NCBI be specific that this is what base system means)
- base system: FTP retrieval from NCBI

Limitation
- no implement
- be transparent, time constraints and pandemic
- no security and authentication

Theoretical Framework
- discusses abstractly anong paglalaruan na ideas
- Discuss Hybrid,CS, P2P in short sentences
- Network, forego layers 1-3. Briefly discuss layers 4 and 5 and their importance.
- Discuss how retrieval operations rely on layers 4 and 5 (FTP for NCBI, torrent??)
- good flow, but too many info, need to shorten it.

Summary of the System
- call this section caDDS
- start with an outline for the entire section
- The retrieval mehanism is inspired by SeqTorr. This section describes the proposed system. This is a hybrid focused or distributed system

Community
- this section is good
- move to 7.2 or after

System Description
- for each photo (GET, POST...) , need one paragraph explaining each one. 1 paragraph per operation (GET, POST...)
- for each photo outline what happens in 1) 2) 3)

Proof of Correctness (BIGTIME CHANGE)
- how do you guarantee that a get does what it is intended to do:
- cite preconditions, post conditions,all the paths from pre conditions to post conditions and how it would always be true
- like an if else
- shows the path from pre condition to post condition
- set pre conditions, assumptions, then show how it becomes postconditions (state after)

Sample:
Assume do not factor in physical state of the network, it is error free
for GET
input of this operation: search key
output of this operation: data transferred to the user

Master node during search has 2 possible output
1) at least 1 hit
2) not in the database

After having at least 1 hit
-> sends the link to user
-> user clicks to link if he wants
-> user presses link
-> user makes a download request to master
-> master redirects user to data node
-> data node sends data to user

Precondition: User chooses to download
Postcondition: User retrieves files

User inputs:search key
master outputs: results report

How to know master will send the correct file?
1) each metadata has a unique identifier
2) guarantees exact file since each file has a unique identifier

How to know it will reach the correct user
1) Each user has a unique identifier (IP Address)

Theoretical Analysis
- Mention base system means FTP from NCBI...
- Mention why only get/ retrieval is used. Since all operations can be broken up into small retrieval operations
- abstract size of a payload
- cite that network is assumed to be error free
- articulate assumptions better
- move the final comparisons here

Conclusion Template:
- say what you did
- summarize analysis of results
- 2 to 3 sentences of what you did/ problem
- 2 sentences correctness and speed
- correctness: system operations
- speed: faster/slower
- shown to do what is expected based on preconditions

Reccomendations
- okay

Acknowledgements -> Remove from Appendix

Biblio
- Make it APA format

### 2020-06-26 Paper2020ACM/MeetingJune25.tex
Revision June 26 3pm

@@
 Whole Paper:
-1) Make it 2 column format
-2) Add sir El King Morado - UP Philippine Genome Center
-3) Remove references to page numbers
-4) Put figures on top of a column (see what sir el king sent)
+check 1) Make it 2 column format
+check 2) Add sir El King Morado - UP Philippine Genome Center
+check 3) Remove references to page numbers
+check 4) Put figures on top of a column (see what sir el king sent)
 
@@
 - okay content
-- fix typoL Data nods
+check - fix typoL Data nods
 
@@
 -Intro Pg 2
-- 3,000,000,000 -> 3 billion
+check - 3,000,000,000 -> 3 billion
 - Shorten HGP section. Combine HGP + NGS to one paragraph.
-- Fix typo: Genomics is is
+check - Fix typo: Genomics is is
 
@@
 
-Acknowledgements -> Remove from Appendix
+check Acknowledgements -> Remove from Appendix
 
 Biblio
-- Make it APA format
+check - Make it APA format

### 2020-06-26 Paper2020ACM/Paper2020.tex
Thesis submission June 25 6pm
Revision June 26 3pm

@@
 \documentclass[acmsmall]{acmart}
+%\documentclass[manuscript,screen]{acmart}
+\usepackage{multicol, caption}
+\usepackage{placeins}
+
+\newenvironment{Figure}
+  {\par\medskip\noindent\minipage{\linewidth}}
+  {\endminipage\par\medskip}
 
@@
 %% the next command will enable that style.
-%%\citestyle{acmauthoryear}
+%\citestyle{acmauthoryear}
 
@@
 %% used to denote shared contribution to the research.
+\author{El King Morado}
+\authornote{Adviser}
+\email{elkingmoradoking@gmail.com}
+\affiliation{%
+  \institution{University of the Philippines - Philippine Genome Center}
+  \city{Quezon}
+  \country{Philippines}
+}
+
 \author{Carl Araya}
-\authornote{Both authors contributed equally to this research.}
-\email{jjazcarraga@up.edu.ph}
+\email{carl.araya256@gmail.com}
 \affiliation{%
@@
 \author{Jose Carlos Rodrigo J. Azcarraga}
-\authornotemark[1]
+\email{jjazcarraga@up.edu.ph}
 \affiliation{%
@@
 }
-\email{jjazcarraga@up.edu.ph}
 
@@
 The proposed system has a master node and data nodes.
-Master node handles administrative work while the Data nods contain all sequence data.
+Master node handles administrative work while the Data nodes contain all sequence data.
 Data nodes should theoretically increase data transfer speeds.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{multicols}{2}
+
 \section{Introduction}
-The goal of this introduction is to explain the idea of genomes, their size and magnitude.
 Inside every human body are the instructions to create and maintain it.
@@
 \subsection{Definition of Genomics}
-Genomics is is the study of wholes sets of genes and their interactions\cite[~p.437]{campbell}.
+Genomics is the study of whole sets of genes and their interactions\cite[~p.437]{campbell}.
 The efforts to study this field has led to volumes of data that need to be analyzed.
@@
 Back in 2003 it was a milestone because the first human genome \cite{introgenomics} had been first completely read.
-It took the project ten years of work, a dozen institutions, and 3,000,000,000 dollars to finish.
+It took the project ten years of work, a dozen institutions, and 3 billion dollars to finish.
 The technology for sequencing and genomics was still at its infancy.
@@
 
-\begin{figure}[h]
+\begin{Figure}
 \centering
-\includegraphics[width=0.65\textwidth]{images/human-gen-cost.jpg} 
-\caption{Cost per Genome Data Over Time}
+\includegraphics[width=0.80\linewidth]{images/human-gen-cost.jpg} 
+\captionof{figure}{Cost per Genome Data Over Time}
 \label{fig:human_gen_cost_fig}
-\end{figure}
+\end{Figure}
 
-To view these technology developments over time, we look at the Figure~\ref{fig:human_gen_cost_fig} on Page~\pageref{fig:human_gen_cost_fig} \cite{genomics-cost}, we note here the expensive cost at the initial stages of sequencing the human genome.
+To view these technology developments over time, we look at the Figure~\ref{fig:human_gen_cost_fig} \cite{genomics-cost}, we note here the expensive cost at the initial stages of sequencing the human genome.
 But as time progresses, the cost substantially went down to what it is now.
@@
 
-\begin{figure}[h]
-\caption{Megabase Cost Over Time}
+\begin{Figure}
+\captionof{figure}{Megabase Cost Over Time}
 \centering
-\includegraphics[width=0.65\textwidth]{images/seq-cost.jpeg} 
+\includegraphics[width=0.80\linewidth]{images/seq-cost.jpeg} 
 \label{fig:megabase_cost_fig}
-\end{figure}
+\end{Figure}
 
-Similar to Figure~\ref{fig:human_gen_cost_fig} on Page~\pageref{fig:human_gen_cost_fig}, we also note the same pattern for the cost of sequencing (not just humans) in Figure~\ref{fig:megabase_cost_fig}\cite{genomics-cost}.
+Similar to Figure~\ref{fig:human_gen_cost_fig}, we also note the same pattern for the cost of sequencing (not just humans) in Figure~\ref{fig:megabase_cost_fig}\cite{genomics-cost}.
 After the human genome project, many more whole genome projects have been completed.
@@
 
-% CID PROGRESS ENDS HEREE FOR MAY 29 2020
-
 \subsection{Dat Protocol}
@@
 
-% CID PROGRESS STARTS HEREE FOR JUNE 2 2020
-
-\begin{table}[h]
+\begin{table*}[ht]
     \caption{Comparisons of Different Databases and Technologies}
@@
     \end{tabular}
-\end{table}
+\end{table*}
 
@@
 
-Table~\ref{table:database_comparison_table} on page ~\pageref{table:database_comparison_table} compares the different technologies studied.
+Table~\ref{table:database_comparison_table} compares the different technologies studied.
 An important point to note is that a majority of the databases are not available.
@@
 \end{itemize}
-
-% CID PROGRESS ENDS HEREE FOR JUNE 2 2020
-
-% CID PROGRESS STARTS HEREE FOR JUNE 3 2020
-% REDID FORMAT
-% CID PROGRESS ENDS HEREE FOR JUNE 3 2020
-
-% CID PROGRESS STARTS HEREE FOR JUNE 4 2020
 
@@
 
-% CID PROGRESS ENDS HEREE FOR JUNE 4 2020
+\begin{Figure}
+\captionof{figure}{App network diagram}
+\centering
+\includegraphics[width=0.8\linewidth]{images/thesis1.png} 
+\end{Figure}
 
-\begin{figure}[h]
-\caption{App network diagram}
+\begin{Figure}
+\captionof{figure}{POST: Uploading a sequence}
 \centering
-\includegraphics[width=0.35\textwidth]{images/thesis1.png} 
-\end{figure}
+\includegraphics[width=0.80\linewidth]{images/thesis1-Page-3.png} 
+\end{Figure}
 
-\begin{figure}[h]
-\caption{POST: Uploading a sequence to the master node}
+\begin{Figure}
+\captionof{figure}{GET: Downloading a sequence}
 \centering
-\includegraphics[width=0.35\textwidth]{images/thesis3.png} 
-\end{figure}
+\includegraphics[width=0.80\linewidth]{images/thesis4.png} 
+\end{Figure}
 
-\begin{figure}[h]
-\caption{GET: Downloading a sequence}
+\begin{Figure}
+\captionof{figure}{PUT: Updating a sequence}
 \centering
-\includegraphics[width=0.35\textwidth]{images/thesis4.png} 
-\end{figure}
+\includegraphics[width=0.80\linewidth]{images/thesis1-Page-5.png} 
+\end{Figure}
 
-%CID: Explain the data model and data permissions
+\begin{Figure}
+\captionof{figure}{DELETE: Deleting a sequence}
+\centering
+\includegraphics[width=0.80\linewidth]{images/thesis1-zzzzz.png} 
+\end{Figure}
 
-\begin{table}[h]
+\begin{table*}[ht]
 \caption{Data Model to be Used in the System}
@@
 \end{tabular}
-\end{table}
+\end{table*}
 
-Table~\ref{table:data_model_table} on Page~\pageref{table:data_model_table} aims to show the metadata that will be initially stored in the database.
+Table~\ref{table:data_model_table} aims to show the metadata that will be initially stored in the database.
 
 %CID: Data Permissions
-
-\begin{table}[h]
+\begin{table*}[ht]
 \caption{Data Permissions and Users in the System}
@@
 \end{tabular}
-\end{table}
+\end{table*}
 
-Table~\ref{table:data_perm_table} on Page~\pageref{table:data_perm_table} shows the different set of users and the permissions allowed to them in the system.
+Table~\ref{table:data_perm_table} shows the different set of users and the permissions allowed to them in the system.
 The moderator is a special user who has the ability to not allow some data to be uploaded.
@@
 
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=\linewidth]{sample-franklin}
-%  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
-%    Ewing, Inc. [Public domain], via Wikimedia
-%    Commons. (\url{https://goo.gl/VLCRBB}).}
-%  \Description{A woman and a girl in white dresses sit in an open car.}
-%\end{figure}
-
-\section{Appendices}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%% ACKNOWLEDGEMENT & GRATITUDE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
+%\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{newapa}
 \bibliography{bibfile}
 
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
+\end{multicols}{2}
 

### 2020-06-27 Paper2020ACM/MeetingJune25.tex
June 27 5pm

--- 2020-06-26 Paper2020ACM/MeetingJune25.tex
@@
 Whole Paper:
-1) Make it 2 column format
-2) Add sir El King Morado - UP Philippine Genome Center
-3) Remove references to page numbers
-4) Put figures on top of a column (see what sir el king sent)
+✓ 1) Make it 2 column format
+✓ 2) Add sir El King Morado - UP Philippine Genome Center
+✓ 3) Remove references to page numbers
+✓ 4) Put figures on top of a column (see what sir el king sent)
 
-Abstract
-- okay content
-- fix typoL Data nods
+✓ Abstract
+✓ - okay content
+✓ - fix typo Data nods
 
 Intro Pg 1
-- Remove sentence: [the goal of this intro]
-- remove subsections in inntro
+✓ - Remove sentence: [the goal of this intro]
 - Genomics study of blank. To generate data you need sequencing. But the data is big & needs effort. To sequence a lot, tech evolved and NGS gave high throughput
@@
 - Flow: Idea of A lot of data -> Problems -> Proposed Solution
--
+- remove subsections in inntro (do this last)
 
 -Intro Pg 2
-- 3,000,000,000 -> 3 billion
+✓ - 3,000,000,000 -> 3 billion
 - Shorten HGP section. Combine HGP + NGS to one paragraph.
-- Fix typo: Genomics is is
+✓ - Fix typo: Genomics is is
 
@@
 Overall info guide: 
-- tama na yung flow
+✓ - tama na yung flow
 - lessen info
-- compress database to 1/2 sentences. Then mention centralized/distributed database.
+- compress database to 1 or 2 sentences. Then mention centralized/distributed database.
 
 Into: CAP
-- make it 1 paragraph
+- shorten and make it 1 paragraph
 - only describe the most important principle and why
@@
 - Discuss P2P in 1 sentence
--Describe Hybrid
-
--Sample: 
+- Describe Hybrid 
 - First we need to discuss 2 basic architecture: CS, P2P. 
@@
 RRL:
-- first paragraph say: We review blank. We would like to focus on blank and blank. In the end we would like to make a comparitive analysis. 
-- Table Comparing the diff Databass: Remove Dat, Describe each column narratively, and describe some conclusions from the table
+- make the first paragraph say: We review blank. We would like to focus on blank and blank. In the end we would like to make a comparitive analysis. 
+- Move Dat (move to theoretical framework)
+- Table Comparing the diff Databass, Describe each column narratively, and describe some conclusions from the table
 - Answer the question: anong halaga ng columns? 
 - Discuss each column CS/P2P/H, and the data on each column
-- FASTA -> Biological
-- Move dat to Methodology.
+✓ - FASTA rename to Biological
 
@@
 - the review tells you you are meant to critique the work you have seen
-- here summarize gaps (mej copy the problems to here)
+- here summarize gaps 
+✓ - copy the problems to here
 
-Problem:
-- summarize the gaps. To review the current implementation has problems in size.
+80 percent ✓ Problem:
+✓ - move the problems bullet points to RRL. Replace problem bullet points with a paragraph summary 
+✓ - summarize the gaps. To review the current implementation has problems in size.
 - To summarize, centralized databases suffer from these problems. Therefore the research aims to focus on developing a database with these features
-- Problem Statement: Central thesis, one unifying question. May certain level of measurability.
+✓ - Problem Statement: Central thesis, one unifying question. May certain level of measurability.
 
@@
 - specify metrics, activities, na once you do this succeed.
-- the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
-- For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
-- keep first 2 objectives, the last is: to evaluate the theoretical properties of the system based on the correctness and its speed
-
-Methodology:
-- dito mapupunta yung framework and APIs
-- say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.
+✓- the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
+✓- For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
+✓- keep first 2 objectives, the last is: to evaluate the theoretical properties of the system based on the correctness and its speed
+✓ - framework and APIs move to theoretical framework.
 
@@
 - 2) evaluate the theoretical running time and speed of the operations.
-- define here what you mean by proposed system and base system (FTP protocol of NCBI be specific that this is what base system means)
-- base system: FTP retrieval from NCBI
+- define here what you mean by proposed system and base system (
+- base system means FTP retrieval from NCBI
 
-Limitation
-- no implement
-- be transparent, time constraints and pandemic
-- no security and authentication
+✓ Limitation
+✓- make limitations section
+✓- no implement
+✓- be transparent, time constraints and pandemic
+✓- no security and authentication
 
 Theoretical Framework
+- should have a very basic flow
 - discusses abstractly anong paglalaruan na ideas
@@
 - good flow, but too many info, need to shorten it.
+- (half check, since this still needs to be moved to theoretical framework) say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.
 
 Summary of the System
-- call this section caDDS
-- start with an outline for the entire section
-- The retrieval mehanism is inspired by SeqTorr. This section describes the proposed system. This is a hybrid focused or distributed system
+✓ - call this section caDDS
+- start with an outline for the entire section (This section will describe the different operations in ...)
+- say: The retrieval mechanism is inspired by SeqTorr. This section describes the proposed system. This is a hybrid focused or distributed system
 
-Community
-- this section is good
-- move to 7.2 or after
+✓ Community
+✓ - this section is good
+✓- move to 7.2 or after
 
@@
 
-Proof of Correctness (BIGTIME CHANGE)
+Proof of Correctness (BIGTIME CHANGE + STUFF TO LEARN)
 - how do you guarantee that a get does what it is intended to do:
@@
 Theoretical Analysis
-- Mention base system means FTP from NCBI...
+- Mention base system means FTP from NCBI (will be stated in scope but good to state again here) ...
 - Mention why only get/ retrieval is used. Since all operations can be broken up into small retrieval operations
@@
 Conclusion Template:
+✓ - make a conclusion section
 - say what you did
@@
 
-Reccomendations
+✓ Reccomendations
 - okay
 
-Acknowledgements -> Remove from Appendix
+✓ Acknowledgements -> Remove from Appendix
 
-Biblio
-- Make it APA format
+✓ Biblio
+✓ - Make it APA format
+
+Presentation - after naaa

### 2020-06-27 Paper2020ACM/Paper2020.tex
June 27 5pm

@@
 
+\begin{table*}[ht]
+    \caption{Comparisons of Different Databases and Technologies}
+    \label{table:database_comparison_table}
+    \begin{tabular}{ccccc}
+    \toprule
+    System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ 
+    \midrule
+    BioTorrents & P2P           & No         & Depends                    & Biological  \\ 
+    SeqTorr     & H             & No         & Fast (PH server)           & Biological       \\ 
+    PeerDB      & P2P           & No         & Depends                    & General      \\ 
+    NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\
+    Dat Protocol    & P2P      & Yes        & Depends           & General  \\ 
+    \bottomrule
+    \end{tabular}
+\end{table*}
+
+\subsection{Comparisons of Database Technologies} \label{compdb}
+
+Table~\ref{table:database_comparison_table} compares the different technologies studied.
+An important point to note is that a majority of the databases are not available.
+This will also be used later on to describe what are the main problems of the different databases.
+
+To restate the different gaps or problems found with other implementations of Genomic Databases: 
+\subsection{Problems of Centralized Databases}
+\begin{enumerate}
+\item Genome researchers download an entire database of genomics data, especially when uncertain how to update their data
+\item Genome researchers only need a portion of the database, but end up downloading the entire database
+\item Client-Server approach can face bottlenecks (server speed) (Section \ref{client-server})
+\item P2P approach has a need of a large unmoderated active user base to work (Section \ref{p2p})
+
+\end{enumerate}
+
+\subsection{Problems of Current Distributed Databases for Genomes}
+\begin{enumerate}
+    \item No existing usable public framework on distributed databases for genomes (Section \ref{compdb})
+    \item Existing distributed database system for genomes is not replicable 
+\end{enumerate}
+
+\section{Problem}
+To summarize the gaps described earlier, there are problems in implementations of centralized databases, specifically with bottlenecks and moderation.
+Other problems are specific to genome researchers with updating of the data.
+The last problem is the gap of usable distributed databases targeted towards genomics researchers.
+The main problem will be: "How to design a distributed database that is useful for genomics researchers?"
+
+\section{Objectives}
+Based on the problems we found described, we propose to:
+
+\begin{itemize}
+    \item Design the framework for an information system that speeds up the download and upload of genomics data
+    \item  Design a distributed database system
+    \item Evaluate the theoretical properties of the system based on the correctness and speed
+\end{itemize}
+
+The first objective addresses the genome researchers problems of centralized databases.
+The second objective aims to resolve the limitations of the client-server and P2P approaches, and the existing problems of distributed databases.
+The third objective aims to show the theoretical rationale for using these distributed databases.
+
+\section{Scope}
+
+The project will focus on creating a framework, and outlining key API features of the framework. The current scope of the project is as follows:
+
+\begin{itemize}
+    \item To create a framework for a proposed system
+    \item To design the correctness of a system
+    \item To look at the basic time complexity of the base system (NCBI) and the proposed system
+    \item To compare the proposed system to the base system
+
+\end{itemize}
+
+\section{Limitations}
+The research had time constraints and physical constraints because of the unexpected Covid-19 pandemic.
+The research will not implement the system or explain the technology stack to be used.
+It will not include security and authentication.
+
+\section{Theoretical Framework}
+
 \subsection{Dat Protocol}
@@
 It also contains the same problems as other P2P protocols, namely, a lack of data moderation and incentives for users to serve as peers.
-
-\begin{table*}[ht]
-    \caption{Comparisons of Different Databases and Technologies}
-    \label{table:database_comparison_table}
-    \begin{tabular}{ccccc}
-    \toprule
-    System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ 
-    \midrule
-    BioTorrents & P2P           & No         & Depends                    & Biological  \\ 
-    SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ 
-    PeerDB      & P2P           & No         & Depends                    & General      \\ 
-    NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\
-    Dat Protocol    & P2P      & Yes        & Depends           & General  \\ 
-    \bottomrule
-    \end{tabular}
-\end{table*}
-
-\subsection{Comparisons of Database Technologies} \label{compdb}
-
-Table~\ref{table:database_comparison_table} compares the different technologies studied.
-An important point to note is that a majority of the databases are not available.
-This will also be used later on to describe what are the main problems of the different databases.
-
-\section{Problem}
-To restate the different gaps or problems found with other implementations of Genomic Databases
-\subsection{Problems of Classical Databases}
-\begin{enumerate}
-\item Genome researchers download an entire database of genomics data, especially when uncertain how to update their data
-\item Genome researchers only need a portion of the database, but end up downloading the entire database
-\item Client-Server approach can face bottlenecks (server speed) (Section \ref{client-server})
-\item P2P approach has a need of a large unmoderated active user base to work (Section \ref{p2p})
-
-\end{enumerate}
-
-\subsection{Problems of Current Distributed Databases for Genomes}
-\begin{enumerate}
-    \item No existing usable public framework on distributed databases for genomes (Section \ref{compdb})
-    \item Existing distributed database system for genomes is not replicable 
-\end{enumerate}
-
-\section{Objectives}
-
-Based on the problems we found described, we propose to:
-
-\begin{itemize}
-    \item Design the framework for an information system that speeds up the download and upload of genomics data
-    \begin{itemize} 
-        \item To resolve 1 \& 2 of the Problems of Classical Databases. 
-    \end{itemize}
-    \item  Design a distributed database system
-    \begin{itemize}
-        \item To resolve 3 \& 4 of the Problems of Classical Databases, and 1 \& 2 of the Problems of Distributed Databases. 
-    \end{itemize}
-\end{itemize}
-
-\begin{itemize}
-    \item For the framework, the system should have the main API’s 
-    \begin{itemize}
-        \item POST genome data
-        \item GET genome data (this will implement a hybrid P2P transfer)
-        \begin{itemize}
-            \item This will aim to speed up number 3 in the Problems of Classical Databases
-        \end{itemize}
-        \item PUT genome metadata (optional: PUT data)
-        \item DELETE genome data
-    \end{itemize}
-\end{itemize}
-
-\section{Scope}
-
-The project will focus on creating a framework, and outlining key API features of the framework. The current scope of the project is as follows:
-
-\begin{itemize}
-    \item To create a framework for a proposed system
-    \item To design the correctness of a system
-    \item To look at the basic time complexity of the base system (NCBI) and the proposed system
-    \item To compare the proposed system to the base system
-
-\end{itemize}
-
-The current scope and coverage of the project will NOT consider:
-
-\begin{itemize}
-    \item Security and Authentication 
-    \item Specific technologies to create the system
-    \item All the data stored in the system (but focus on main data only)
-    \item Implementing the actual system
-\end{itemize}
-
-\section{Theoretical Framework}
 
@@
 
-\subsection{Summary of the System}
+\begin{itemize}
+    \item (Move to Theoretical Framework/caDDS) The system is an information system, and all information systems must have the main API’s 
+    \begin{itemize}
+        \item POST genome data
+        \item GET genome data (this will implement a hybrid P2P transfer)
+        \begin{itemize}
+            \item This will aim to speed up number 3 in the Problems of Classical Databases
+        \end{itemize}
+        \item PUT genome metadata (optional: PUT data)
+        \item DELETE genome data
+    \end{itemize}
+\end{itemize}
+
+\subsection{caDDS: Community Assisted Distributed Database for Sequences}
 
@@
 Thus the "Distributed Database for Sequences" portion of the title.
-
-\subsubsection{Community Aspect} \label{community}
-The system was also described as community assisted.
-The definition of \textbf{community} to be used here is a \textit{set of people and institutions who generally collaborate together to do research}.
-And the focus is the genome research community.
-Reemphasizing certain features from the database, that any registered user can upload data.
-This means all users in the community can contribute to the knowledge base, and contribute data to be used by the community.
-Another important community aspect is that institutions can increase the number of data nodes.
-Meaning there will be an increase in the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later).
-Frequently used local data as well will be more easily accessible (since data nodes should have faster access speeds).
-The concept of closer networks having faster speeds (discussed in the network section) is at play here.
-Another community aspect will be this paper, publicly describing many aspects of the structure and code.
-So that this concept may be further developed and even implemented by genome research communities.
-
-To summarize, this helps the community since the community may add their own data.
-The community may increase storage (through data nodes) these may also assist in increasing data speeds.
-And frequently used data will have faster access.
 
@@
 All other users, which are called guests, have an ability to download data.
+
+\subsection{Community Aspect} \label{community}
+The system was also described as community assisted.
+The definition of \textbf{community} to be used here is a \textit{set of people and institutions who generally collaborate together to do research}.
+And the focus is the genome research community.
+Reemphasizing certain features from the database, that any registered user can upload data.
+This means all users in the community can contribute to the knowledge base, and contribute data to be used by the community.
+Another important community aspect is that institutions can increase the number of data nodes.
+Meaning there will be an increase in the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later).
+Frequently used local data as well will be more easily accessible (since data nodes should have faster access speeds).
+The concept of closer networks having faster speeds (discussed in the network section) is at play here.
+Another community aspect will be this paper, publicly describing many aspects of the structure and code.
+So that this concept may be further developed and even implemented by genome research communities.
+
+To summarize, this helps the community since the community may add their own data.
+The community may increase storage (through data nodes) these may also assist in increasing data speeds.
+And frequently used data will have faster access.
 
@@
 
+\section{Conclusion}
+
 \section{Recommendations}

### 2020-06-29 Paper2020ACM/Paper2020.tex
Updates from Overleaf

@@
 \section{Introduction}
-Inside every human body are the instructions to create and maintain it.
-These instructions are found inside the genome.
-The instructions inside the human genome are incredibly long.
-For example, entire human genome, inside a singular cell, when laid out is around 2 meters in length.\cite{ency_sci_tech}
 
-Combining  and laying out all the information for all the cells in the human body.
-The length is incredibly long that it can span the earth to moon 6000 times.
-This describes as well the vast amount of DNA information in the body.
-These kinds of information will help scientists explain many of the understood and unknown processes inside all organisms.
-These may also be used to compare different organisms to each other.
-And to compare humans to other humans.
+There is a vast amount of information inside the human body.
+One of these key biological information storage designs are in the genes.
+These tell each cell how to organize, how to build, maintain the organism\cite{alberts_mole}.
+These genes inside a single cell (humans) when laid out is 2 meters in length.\cite{ency_sci_tech} And combining all of the cells, the information laid out would reach the moon, and back, 6000 times.
 
-\subsection{Definition of Genomics}
-Genomics is the study of whole sets of genes and their interactions\cite[~p.437]{campbell}.
-The efforts to study this field has led to volumes of data that need to be analyzed.
-Which has also spurred the field bioinformatics, which aims to use computers to aid the analysis and storage of biological data.
-Institutions such as the National Institute of Health and National Library of Medicine, have led this field and built a website to encourage collaboration online through their online portal called NCBI (National Center for Biotechnology Information).
-\cite[~p.438]{campbell} This website will be discussed later in more detail in Section \ref{NCBI}.
+These biological information storage designs or genes are studied in the field of \textit{Genomics}.
+Genomics studies the whole sets of genes and their interactions \cite[~p.437]{campbell}.
+The definition by the NIH (National Institute of Health) adds to that definition by including the study of how these genes interact with the environment.\cite{genomics-definition} Genes are large, and the human genome is filled with 3.2 billion pairs of information (or bases) \cite{introgenomics}.
 
-National Institute of Health \cite{genomics-definition} formally defines genomics as \\ \\
-\textit{the study of all of a person's genes (the genome), including interactions of those genes with each other and with the person's environment.} \\ \\
+Before these information is analyzed digitally, they must be read into a machine first.
+This is called \textit{Sequencing}.
+Here a machine takes fragments of genes and reads them into a digital format.
+These machines take in a lot of information from different but redundant fragments of genes to get a complete reading.
+Another set of computational processes are needed to combine these information to form the gene.
+Back in the 1980s the speed of sequencing was still slow at around 1000 bases a day\cite{campbell}.
 
-In the definition, one main component of genomics are genes.
-Genes comprise of DNA.
-DNA, deoxyribonucleic acid, is a chemical compound that works like a computer bytes to store information for organisms.
-DNA's structure is composed of chemical compounds (or nucleotides) that encode information.
-These nucleotides are commonly written down as A,G,C,T corresponding to their chemical makeup\cite{alberts_mole}.
-Another important feature of genes is that these genes are the instructions passed on from one generation to another.
-\cite[p.~2]{hartl2018}
+In 1993, when sequencing technology was still at its infancy.
+There was an initiative to complete the entire human genome\cite{introgenomics}.
+It was a slow process which took 10 years, a lot of collaborations, 3 billion dollars.
+In 2000 when the human genome project had been ongoing, the speed of sequencing grew to 1000 base pairs every second, 24 hours a day, 7 days a week \cite{campbell}.
+The technology became faster and more reliable.
 
-These nucleotides form sets of base pairs (otherwise known as \textit{bases}).
-These bases encode the information stated earlier.
-They are pairs since, there is always a corresponding complement pair for every nucleotide.
-To give a bigger picture, when these bases are combined and organized, these act as a blueprint to describe and instruct the body(or organism) on how to create it and how to maintain it.\cite{alberts_mole}
-
-If all the bases, genes, genomics combined form the blueprint of the organism.
-There is useful knowledge there and there will be a need to understand all the complexity and interactions to further understand how the blueprint of an organism works.
-Genomics play a role in understanding different components of an organism by paying attention to the genes and nucleotide bases.
-Different organisms have certain differences in genes and this has been a focal point of many genomics studies, which sometimes is aimed as well as establishing surprising similarities among distinct organisms.
-
-The instructions inside the genes explain the complexity of an organism.
-This also means more complex organisms are likely to have more complicated genes and therefore have a lot more information.
-Even if this general idea is true, there are some exceptions, some plants such as the Japanese canopy plant \textit{Paris japonica} have 50 times the genome size compared to humans\cite[~p.442]{campbell}.
-These may be due to repeating non-essential DNA.
-But the idea of having many unknown components and interactions in a complex organism means more information will be needed and gathered.
-Humans, being complex, are extensively studied in order to detail the complex processes inside the body.
-The presence of many studies for the human genome has increased the amount of information generated.
-To give an idea of the amount of information, for humans, the full human genome contain around $3.2x10^9$ base pairs \cite{introgenomics}.
-
-Storing information is one aspect, but before analyzing that information, these have to be physically taken from the organism's cells and read.
-Reading these information from genes will need a certain technology.
-The technology is called \textit{Sequencing}.
-
-\subsection{Sequencing}
-
-The machines to generate the sequences are aptly called \textbf{Sequencers} that look (or more appropriately, \textit{read}) the DNA of a specimen.
-The machine outputs A,G,C,T, sometimes with uncertainties.
-Sequencers look at the exact order of a single strand of DNA.
-The process of sequencing DNA seems straightforward that a machine will read only one strand of DNA.
-In actuality, one strand is too little for one machine to read.
-So before reading, the DNA is \textit{replicated} multiple times in the laboratory.
-Here there is a concept of repeated reading or coverage which will be discussed later.
-
-Sequencers also can't read the entire length of DNA.
-With this, the DNA is also \textit{fragmented} before processing.
-By fragmented meaning it will be cut up into pieces.
-Sequencers can only read it in parts.
-
-The idea of \textit{replication} and \textit{fragmentation} are important because these increase the data taken up when sequencing.
-Replication causes a lot of redundant data in the sequence.
-Fragmentation causes there to be a need to piece back the information together.
-Now with those concepts in mind sequencers then reads (or sequences) the replicated and fragmented data.
-
-These reads or sequences taken will need an extra step before the data becomes useful.
-It will then be needed to be put back together computationally to form the \textit{correct or reference} DNA.
-These utilize many of the technology developed by bioinformatice.
-These are one of the reasons why genomic data is large because even the fragmented replicated data is still stored, not just the reference DNA.
-
-\subsection{Human Genome Project}
-
-Another way to see the way genomic data is large is by looking at the \textit{Human Genome Project (HGP)}.
-The project's aim was to sequence the entire human genome.
-At the time it needed a lot of collaborations, a lot of financing, and new computational tools (piecing everything together).
-Back in 2003 it was a milestone because the first human genome \cite{introgenomics} had been first completely read.
-It took the project ten years of work, a dozen institutions, and 3 billion dollars to finish.
-The technology for sequencing and genomics was still at its infancy.
-
-Once the Human Genome Project was completed in 2003 and the studies fully published in 2006.
-The research then paved the way to study chimpanzees since their DNA is the closest to humans.
-And the whole genome structure of humans can easily be used to recreate the whole genome sequence of the chimpanzee.
-This was completed 2 years after the Human Genome Project\cite{campbell}.
-These paved the way for more and more whole genome studies of animals much different from humans.
-By April 2013, around 4,300 genomes had been studied\cite[~p.442]{campbell}.
-These contribute to the growing amount of research, analysis, and most importantly data around genomes of different organisms.
-
-\subsubsection{Next Generation Sequencing}
-
-The introduction of \textit{Next Generation Sequencers or (NGS)} can be seen as a another significant event in the history of genomics.
-This is a new technology which has allowed the machines to read faster and cheaper.
-These utilize a variety of technologies, some make use of light, some use ion semiconductors, others use a special DNA marker with a fluorescent probe \cite[~p.67]{paulselzer2018}.
-Back in 1980 the speed in which sequencing can be done was 1000 base pairs a day.
-In 2000 when the human genome project had been ongoing, the speed grew to 1000 base pairs every second, 24 hours a day, 7 days a week \cite{campbell}.
-In 2000 the technology had been growing faster and more reliable.
-
-NGS has contributed to this speedup even further.
-Many countries that research have started employing NGS.
-The NGS technology in Japan can now read 10,000 human genomes per day.
-\cite[p.~19]{introgenomics} Compare this to the many years and institutions it took to study the first human genome back in 2003.
-Now a singular laboratory can achieve this with speed.
-
-These show how the technology for reading genomes has been developed over the years.
-This tech has increased the speed of reading, and as well as research.
-Which directly affects the speed in which genome data is produced.
-Since each sequence read requires storage.
-This poses a problem to projects involving sequences, causing the data volumes to be usually very large\cite{bon_compression}.
-
-There exists a concept of \textit{coverage} as well.
-Referring back to the \textit{replication} and \textit{fragmentation} concept earlier.
-There needs to be a specific number of reads a machine needs in order to produce a good quality reference sequence (the correct, standard sequence).
-Meaning the sequence isn't read only once, where multiple reads need multiple reads.
-
-This contributes to the large data size of sequence reads.
-Currently, for a coverage of around 60, the file size for a compressed human genome can easily reach 200 GB.
-And a project of 10-20 genomes would need approximately 4 TB of storage.
-Transferring these data across research groups would indeed be a challenge.
-\cite[~p.68]{paulselzer2018} These highlight the important and growing problem of storage and file transfers in genome research.
+After the human genome project finished, there was a lot of data to be analyzed and many papers were published in the next few years.
+The research led to many advancements and many more genomes to be studied.
+From the whole genome of the closest relative to humans (chimpanzee) in 2006 to 4300 different whole genomes in 2013\cite[~p.442]{campbell}.
+Part of the reason why research was capable of looking at more whole genomes was because of next generation sequencing (NGS).
+NGS used a variety of technologies to make sequencing faster, cheaper, and more reliable\cite[~p.67]{paulselzer2018}.
+The technology speedup can be noticed in Japan, where today 10,000 human whole genomes can be studied in a single day.
+This has greatly increased the speed of research and data production in the field of genomics.
 
@@
 
-To view these technology developments over time, we look at the Figure~\ref{fig:human_gen_cost_fig} \cite{genomics-cost}, we note here the expensive cost at the initial stages of sequencing the human genome.
-But as time progresses, the cost substantially went down to what it is now.
-The big fall of the prices around 2004 - 2005 corresponds to the introduction of the Next Generation Sequencers.
-
-\subsection{Database}
-
 \begin{Figure}
@@
 
-Similar to Figure~\ref{fig:human_gen_cost_fig}, we also note the same pattern for the cost of sequencing (not just humans) in Figure~\ref{fig:megabase_cost_fig}\cite{genomics-cost}.
-After the human genome project, many more whole genome projects have been completed.
-And the cost of sequencing dropped 1000 fold from 2008 to 2012 alone \cite{bon_compression}.
-This lowering of cost, again corresponds to the idea of sequencing becoming more available.
-Increasing the number of research projects, therefore contributing to the growing number again of genomic data.
+In the Figure~\ref{fig:human_gen_cost_fig} \cite{genomics-cost}, it can be noticed that the cost per genome data goes down over time (note the year 2004-2005) where NGS was introduced.
+In this next Figure~\ref{fig:megabase_cost_fig}\cite{genomics-cost} we also note the drop of sequencing cost from 2008 to 2012 by 1000 fold \cite{bon_compression}.
+These graphs show how the cost has been reduced which should increase the data production by a huge amount as well.
 
-\begin{center}
-\textit{Moore's law states that the number of transistors on a chip doubles every two years.}\cite{kurose}
-\end{center}
-
-The idea of Moore's law should give relief to genomic researchers since this means, that as time passes by, technology for processing and storage will improve.
-And the overall capacity for what technology can do will increase as well.
-
-But, by paying attention to the speed in which the number of sequence data increases seems to exceed the speed of Moore's Law.
-And this will pose an additional problem since the technology doesn't seem to be catching up.
-The immediate problem is the storage of all the data being produced.
-A certain technology needs to be made for the purpose of holding this data.
-This technology is called \textit{Databases}.
-But there is a greater need as well in making sure the data stays secure, consistent, easily accessible, and fast.
-These are other aspects needed by a genomics researcher as well.
-And this is the reason why a specific class of Databases called \textit{Distributed Databases} will be discussed in section \ref{db_types}.
-
-\subsubsection{Databases Definition}
-A database (DBMS, Database Management System) is defined by containing information about a particular enterprise. Functionalities include: \cite{Silberschatz2010}
-\begin{itemize}
-    \item Collection of interrelated data
-    \item Set of programs to access the data
-    \item An environment that is both convenient and efficient to use
-\end{itemize}
-
-These form a system that is made for accessing and updating data.
-The next section (\ref{cap}) discusses the key points to consider in creating a good database system (or DBMS).
-The key points are important to discuss many important features and dilemmas of the proposed system.
-Certain database principles are needed to make sure the data is available, secure, and consistent.
-
-\subsubsection{Database Types} \label{db_types}
-Two types of databases will be discussed in this paper.
-Each have their specific niche and use.
-One is more efficient and easier to make, another is better suited for collaborative work.
-Note the technical jargon that needs to be used to explain some concepts.
-These are namely: \cite{centralizedvsdistributed}
-
-\begin{itemize}
-    \item Centralized - In a centralized database, all data are managed by a single DBMS in a single node to be distributed to users.
-As a result, transactions are easily manageable because everything can be processed in the single node.
-Here, the cost of communication is high and reliability is low, as an error in the database will mean disaster to the network as a whole.
-These are easier to create.
-
-    \item Distributed - In a distributed database, data are stored in multiple servers in different geographical locations.
-Each node may contain a part or the entirety of the data.
-The difficulty of managing transactions will go up, but the result is higher availability, faster response, and modular growth, among other advantages.
-These databases are better creating spread out databases, and are better for collaborative work.
-\end{itemize}
-
+The huge data must be stored in a form of digital information storage design as well.
+These are where the technologies of databases come in.
+Databases or (DBMS: Database Management Systems) manage the digital information, and the access to it.
+\cite{Silberschatz2010}.
+The two types of databases are the centralized and the distributed databases\cite{centralizedvsdistributed}.
+The centralized databases have a single node or a single access point.
+While distributed databases have multiple nodes, and each node contains a portion or a replica of the data.
+Distributed databases are more complex but have higher availability, faster response, and is better for collaborative work.
 For genomic research, when the data storage needs to be increased often, distributed databases are most appropriate.
@@
 
+Figure 3 shows the way in which the different components of the network interact with each other. The master node is connected to all data nodes for syncing purposes.
+
 \begin{Figure}
@@
 %%
+
+\bibliography{bibfile}

### 2020-07-07 Paper2020ACM/MeetingJune25.tex
Latest edits

@@
 ✓ - 3,000,000,000 -> 3 billion
-- Shorten HGP section. Combine HGP + NGS to one paragraph.
+✓- Shorten HGP section. Combine HGP + NGS to one paragraph.
 ✓ - Fix typo: Genomics is is
 
--Intro: Database
-- NGS -> high voluminous data -> need to store it in a digital format -> retrieve it for analysis
+✓-Intro: Database
+✓- NGS -> high voluminous data -> need to store it in a digital format -> retrieve it for analysis
 
@@
 
-Scope
-- 2 things: 
-- 1) create a theoretical design for a proposed system
-- 2) evaluate the theoretical running time and speed of the operations.
-- define here what you mean by proposed system and base system (
-- base system means FTP retrieval from NCBI
+✓Scope
+✓- 2 things: 
+✓- 1) create a theoretical design for a proposed system
+✓- 2) evaluate the theoretical running time and speed of the operations.
+✓- define here what you mean by proposed system and base system (
+✓- base system means FTP retrieval from NCBI
 

### 2020-07-07 Paper2020ACM/Paper2020.tex
Latest edits

@@
 These are where the technologies of databases come in.
-Databases or (DBMS: Database Management Systems) manage the digital information, and the access to it.
+Databases or DBMS (Database Management Systems) manage the digital information, and the access to it.
 \cite{Silberschatz2010}.
@@
 While distributed databases have multiple nodes, and each node contains a portion or a replica of the data.
-Distributed databases are more complex but have higher availability, faster response, and is better for collaborative work.
+Distributed databases are more complex but have higher availability, faster response, and are better for collaborative work.
 For genomic research, when the data storage needs to be increased often, distributed databases are most appropriate.
@@
 
-The project will focus on creating a framework, and outlining key API features of the framework. The current scope of the project is as follows:
-
-\begin{itemize}
-    \item To create a framework for a proposed system
-    \item To design the correctness of a system
-    \item To look at the basic time complexity of the base system (NCBI) and the proposed system
-    \item To compare the proposed system to the base system
-
-\end{itemize}
+The research will focus on creating a theoretical design for the proposed system.
+And to evaluate the theoretical properties of the proposed system and the base system.
+The proposed system is the caDDS (community assisted Distributed Databases for Sequences), specifically the GET operation.
+While the base system, is the FTP (File Transfer Protocol) of NCBI for getting sequences.
+NCBI is currently a complex platform, but the scope of this research will only be the FTP access route.
 

### 2020-07-20 Paper2020ACM/MeetingJune25.tex
July 2020 edits

@@
 
-Intro Pg 1
+✓Intro Pg 1
 ✓ - Remove sentence: [the goal of this intro]
-- Genomics study of blank. To generate data you need sequencing. But the data is big & needs effort. To sequence a lot, tech evolved and NGS gave high throughput
-- Intro should be 2 columns lang. Make each paragraph a few sentences instead.
-- Intro goal: layoutht background of the work
-- Flow: Idea of A lot of data -> Problems -> Proposed Solution
-- remove subsections in inntro (do this last)
+✓- Genomics study of blank. To generate data you need sequencing. But the data is big & needs effort. To sequence a lot, tech evolved and NGS gave high throughput
+✓- Intro should be 2 columns lang. Make each paragraph a few sentences instead.
+✓- Intro goal: layoutht background of the work
+✓- Flow: Idea of A lot of data -> Problems -> Proposed Solution
+✓- remove subsections in inntro (do this last)
 
--Intro Pg 2
+✓Intro Pg 2
 ✓ - 3,000,000,000 -> 3 billion
@@
 
-Overall info guide: 
+✓ Overall info guide: 
 ✓ - tama na yung flow
-- lessen info
-- compress database to 1 or 2 sentences. Then mention centralized/distributed database.
+✓ - lessen info
+✓ - compress database to 1 or 2 sentences. Then mention centralized/distributed database.
 
-Into: CAP
-- shorten and make it 1 paragraph
-- only describe the most important principle and why
+✓ Into: CAP
+✓ - shorten and make it 1 paragraph
+✓ - only describe the most important principle and why
+✓- CID: Carlll paread through nalang ng nagawa ko
 
 Application Architecture: (Move to Theoretical Framework)
-- Discuss Client Server in 1 sentence
-- Discuss P2P in 1 sentence
-- Describe Hybrid 
-- First we need to discuss 2 basic architecture: CS, P2P. 
-- CS is blank. 
-- P2P is blank. 
-- Our proposed system will be hybrid getting these aspects from CS and P2P
+✓ - Discuss Client Server in 1 sentence
+✓- Discuss P2P in 1 sentence
+✓- Describe Hybrid 
+✓- First we need to discuss 2 basic architecture: CS, P2P. 
+✓- CS is blank. 
+✓- P2P is blank. 
+✓- Our proposed system will be hybrid getting these aspects from CS and P2P
 
@@
 - Move Dat (move to theoretical framework)
-- Table Comparing the diff Databass, Describe each column narratively, and describe some conclusions from the table
+(Todo monday) - Table Comparing the diff Databass, Describe each column narratively, and describe some conclusions from the table
 - Answer the question: anong halaga ng columns? 
@@
 - specify metrics, activities, na once you do this succeed.
-✓- the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
-✓- For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
+✓ - the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
+✓ - For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
 ✓- keep first 2 objectives, the last is: to evaluate the theoretical properties of the system based on the correctness and its speed
@@
 Theoretical Framework
-- should have a very basic flow
-- discusses abstractly anong paglalaruan na ideas
-- Discuss Hybrid,CS, P2P in short sentences
-- Network, forego layers 1-3. Briefly discuss layers 4 and 5 and their importance.
-- Discuss how retrieval operations rely on layers 4 and 5 (FTP for NCBI, torrent??)
-- good flow, but too many info, need to shorten it.
-- (half check, since this still needs to be moved to theoretical framework) say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.
+✓ - should have a very basic flow
+✓ - discusses abstractly anong paglalaruan na ideas
+✓ - Discuss Hybrid,CS, P2P in short sentences
+✓ - Network, forego layers 1-3. Briefly discuss layers 4 and 5 and their importance.
+✓ - Discuss how retrieval operations rely on layers 4 and 5 (FTP for NCBI, torrent??)
+✓ - good flow, but too many info, need to shorten it.
+✓ - (half check, since this still needs to be moved to theoretical framework) say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.
 

### 2020-07-20 Paper2020ACM/Paper2020.tex
July 2020 edits

@@
 \begin{abstract}
-Genomic data the past few years have grown exponentially with the technology barely keeping up.
+Genomics data the past few years have grown exponentially with the technology barely keeping up.
 The problem faced by genome researchers is the large data set and difficulty transferring the files.
@@
 The proposed system has a master node and data nodes.
-Master node handles administrative work while the Data nodes contain all sequence data.
+The master node handles administrative work while the Data nodes contain all sequence data.
 Data nodes should theoretically increase data transfer speeds.
@@
 One of these key biological information storage designs are in the genes.
-These tell each cell how to organize, how to build, maintain the organism\cite{alberts_mole}.
-These genes inside a single cell (humans) when laid out is 2 meters in length.\cite{ency_sci_tech} And combining all of the cells, the information laid out would reach the moon, and back, 6000 times.
+These tell each cell how to organize, how to build, and how to maintain the organism\cite{alberts_mole}.
+The genes inside a single cell (humans) when laid out is 2 meters in length.\cite{ency_sci_tech} And combining all of the cells, the information laid out would reach the moon, and back, 6000 times.
 
@@
 Genomics studies the whole sets of genes and their interactions \cite[~p.437]{campbell}.
-The definition by the NIH (National Institute of Health) adds to that definition by including the study of how these genes interact with the environment.\cite{genomics-definition} Genes are large, and the human genome is filled with 3.2 billion pairs of information (or bases) \cite{introgenomics}.
+The definition by the NIH (National Institute of Health) adds to that definition by including the study of how these genes interact with the environment.\cite{genomics-definition} Genes are large, the human genome specifically is filled with 3.2 billion pairs of information (or bases) \cite{introgenomics}.
 
-Before these information is analyzed digitally, they must be read into a machine first.
+Before these information are analyzed digitally, they must be read into a machine first.
 This is called \textit{Sequencing}.
@@
 Another set of computational processes are needed to combine these information to form the gene.
-Back in the 1980s the speed of sequencing was still slow at around 1000 bases a day\cite{campbell}.
+Back in the 1980s the speed of sequencing was still slow and the top laboratory could only process around 1000 bases a day\cite{campbell}.
 
@@
 There was an initiative to complete the entire human genome\cite{introgenomics}.
-It was a slow process which took 10 years, a lot of collaborations, 3 billion dollars.
-In 2000 when the human genome project had been ongoing, the speed of sequencing grew to 1000 base pairs every second, 24 hours a day, 7 days a week \cite{campbell}.
+It was a slow process which took 10 years, a lot of collaborations, and 3 billion dollars.
+In 2000, when the human genome project had been ongoing, the speed of sequencing grew to processing 1000 base pairs every second, 24 hours a day, 7 days a week \cite{campbell}.
 The technology became faster and more reliable.
@@
 Distributed databases are more complex but have higher availability, faster response, and are better for collaborative work.
-For genomic research, when the data storage needs to be increased often, distributed databases are most appropriate.
+For genomics research, when the data storage needs to be increased often, distributed databases are most appropriate.
 
-\subsubsection{CAP theorem for Databases} \label{cap}
-There are 3 principles needed when creating a database\cite{Silberschatz2010}.
-It is given that it is impossible to do all 3 principles at the same time perfectly\cite{gilbert2012perspectives}.
-The idea is that only two of the three can be focused on at any given time.
-These principles and limitations need to be taken into consideration when designing the system.
+Databases also have three principles \cite{Silberschatz2010}, these 3 principles form the ideal database.
+But these 3 principles, practically cannot be done all at the same time \cite{Silberschatz2010}.
+These are the considerations and limitations that need to be addressed.
+One principle is consistency, data should be accurate and faithfully replicated.
+Another is availability, is the database able to serve users all of the time?
+The down times should be as infrequent as possible.
+The last and which will be the focus is Partition Tolerance.
+Partition Tolerance is the resiliency of a system.
+What happens when one portion malfunctions, will the whole system fail or will the rest easily adjust?
+This is necessary for a distributed database.
+And since Partition Tolerance \cite{Silberschatz2010} is chosen, then the trade off will be between availability and consistency.
 
-\begin{itemize}
-    \item Consistency - Does the data have integrity?
-To update one part, all other (replicated) parts should be updated quickly as well; Data is accurate and copied faithfully.
-An example would be that genes for specimen A, when copied to another data node should contain the exact copy of the data, with no changes.
-    \item Availability - Is the database usable by clients all of the time? The entire system must be in downtime as infrequently as possible, despite failures.
-    \item Partition Tolerance - How resilient is the system when it comes to failures?
-If one portion breaks, the other portions should still be entirely usable.
-An example would be a power failure for one system, how would this affect the other systems?
-\end{itemize}
+The trade off for this system is between availability and consistency.
+For a consistency-priority-database, a change in one portion of the database, a create SQL command, will lock the entire database preventing other changes (reducing availability), then it will make sure that change is replicated in all relevant portions of the database, before making the database available again.
 
-Partition tolerance is a de facto requirement of distributed databases and cannot be avoided, so the main trade-off is between availablity and consistency.
-\cite{Silberschatz2010} Suppose a distributed system prefers consistency over availability.
-A user updates an SQL row in one partition of the database.
-Beforehand, this row has already been replicated among many partitions.
-To ensure that all replicas of the row are consistent, the system must lock this part of the database to prevent other users from reading and editing it, so availability will be negatively affected in the meantime.
-A distributed system preferring availability, such as a messenger app, will not lock parts of the database in the purpose of being open to serve its users, but cannot guarantee that the data it outputs is always up-to-date.
-There might be a delay before the data is fully consistent throughout geographic locations.
-
-\subsection{Application Architecture}
-Another important consideration is how a system is built.
-These describe the blueprint and interaction within a database system.
-These provide insights similar to the questions asked when looking between database types.
-Database types describe how the data is handled inside the system.
-But application architecture describes how the users interact with the system.
-The application architecture is how a networked app is structured across the various end systems (i.e.
-computers, servers, etc.) \cite{kurose} The end systems are another way to call users or clients.
-The application architecture types fall into either client-server, peer-to-peer, or hybrid architecture.
-
-\subsubsection{Client-Server} \label{client-server}
-A client-server architecture relies on a Web server that is always available and has a fixed IP address.
-When a client (or user) sends a request for data, the server (or DBMS in this example) responds by sending the requested data.
-The client-server architecture is mainly built to respond to one client at a time.
-The architecture may handle multiple connections but as these grow bigger, it may not be able to keep up with the increasing number of connections since the server might not be able to handle all the requests.
-
-\subsubsection{P2P} \label{p2p}
-A peer-to-peer (P2P) architecture has minimal to no reliance on always-on, dedicated servers.
-Peers are desktops/laptops (or users) that communicate to each other without passing through a dedicated server (or DBMS).
-These include internet, BitTorrent, and some other file sharing protocols.
-An advantage of P2P networks is their scalability, unlike the earlier type.
-Each user can benefit the whole network by sharing his/her own resources.
-
-However, since P2P doesn't have a central data management system. P2P has its own problems, which are security and incentivizing users to share their resources.
-
-\subsubsection{Hybrid}
-A hybrid architecture combines elements of P2P and client-server architectures, attempting to utilize advantages of both.
-
-For a genomic database system, it would be best to utilize Hybrid Application Architecture.
-Since there needs to be certain features that Client-Server of P2P might not be able to deliver.
-One would be scalability, another is security and centralized management of the data, which is a feature not provided by P2P.
+For an availability-priority-database, an example would be the messenger Facebook app.
+Where the focus of the app is speed.
+Here changes in a database will not lock the database, to make its app available.
+The downside is that the data is available, but it is not always up to date, and time is needed before the data is fully consistent on all the relevant databases.
 
@@
 \subsection{SeqTorr: A distributed scalable database for genomic information}
-The focus on SeqTorr was finding a way to scale databases meant specifically for genomic information.
+The focus on SeqTorr was finding a way to scale databases meant specifically for genomics information.
 There was also a focus on being able to choose information to share and use.
@@
 Having a distributed scalable local infrastructure to store genomic data instead of relying on NCBI would be beneficial since: only certain sequences in NCBI are relevant to the institutions (e.g.
-Asian sequences are needed more than Caucasian ones), researchers within the country can share data with fast speeds, and  institutions can share in the hosting of data.
+Asian sequences are needed more than Caucasian ones), researchers within the country can share data with fast speeds,and institutions can share in the hosting of data.
 
@@
 
-SeqTorr seems to go in the right direction and focus for genomic databses. But currently, SeqTorr is not available, and the code base does not have any documentation which makes future work difficult.
+SeqTorr seems to go in the right direction to focus on genomics databases.
+But currently, SeqTorr is not available, and the code base does not have any documentation which makes future work on this difficult.
 
@@
 Scientific data continues to grow, and so does the demand for easier accessibility.
-Centralized servers, as stated earlier, using HTTP or FTP cannot keep up with concurrent requests, and peer-to-peer protocols does not scale well for large files.\cite{biotorrents} One example of these centralized servers is NCBI, which will be discussed later.
+Centralized servers, as stated earlier, which are using HTTP or FTP, cannot keep up with concurrent requests, and peer-to-peer protocols does not scale well for large files.\cite{biotorrents} One example of these centralized servers is NCBI, which will be discussed later.
 BitTorrent handles both these problems.
@@
 It is much like a public tracker \cite{biotorrents}.
-The main website hosts .torrent and metadata files.
+The main website hosts torrent and metadata files.
 The main website doesn't store data, rather it lets the users keep the data and share it to other users.
@@
 that helps users in searching for relevant datasets.
-There is also a torrent file has a list of trackers, servers that "know" which peers are serving which files, so it can download from those peers simultaneously.
+There is also a torrent file that has a list of trackers, servers that "know" which peers are serving which files, so it can download from those peers simultaneously.
 These torrents act like a directory telling the user who currently has the data for a piece of the file.
@@
 The client (or user) must have their computer continuously turned on to keep their file available to others.
-The main issue with BioTorrents, which is a problem as well in P2P, is that it requires lots of users that will keep many torrents available as often as possible.\cite{biotorrents} These entail that for the system to work, many people need to use it and all these people need to keep their computers on all the time.
+The main issue with BioTorrents, which is a problem as well in P2P, is that it requires lots of users to keep torrents available as often as possible.\cite{biotorrents} These entail that for the system to work, many people need to use it and all these people need to keep their computers on all the time.
 
@@
 PeerDB is a proposal for a distributed system using P2P.
-For this to be used in the context of genomic biological work, a main website which facilitates sharing will be need to be put up.
+For this to be used in the context of genomics biological work, a main website which facilitates sharing will be need to be put up.
 PeerDB needs to be installed as well on all the users who will be participating.
@@
 \subsection{NCBI} \label{NCBI}
-The National Center for Biotechnology Information or more commonly known as NCBI, which discussed briefly in the section on \textit{BioTorrents} and in the \textit{Introduction}.
-This is the standard genomic database used by many researches and institutions.
-We will refer this later on as well as the \textit{base system}.
-This is the website used to access and store most genomic information\cite{campbell}.
+The National Center for Biotechnology Information or more commonly known as NCBI, which discussed briefly in the section on \textit{BioTorrents} and in the \textit{Introduction}, is the standard genomics database used by many researches and institutions.
+We will refer a portion of this later on as the \textit{base system}.
+This is the website used to access and store most genomics information around the world\cite{campbell}.
 It uses a client-server architecture.
@@
 
-The SeqTorr authors\cite{seqtorr} also stated that NCBI is the main and current DBMS used to get and share genomic data.
-But one difficulty using that is updating the current genomic data.
-There currently is no way to check which data (on your computer) has been updated, which sequences were corrected.
+The SeqTorr authors\cite{seqtorr} also stated that NCBI is the main and current DBMS used to get and share genomics data.
+But one difficulty using that is updating the current genomics dataset.
+There is currently no way to check which data (on your computer) have been updated, which sequences were corrected.
 And the way the authors solved this was to regularly re-download the entire NCBI database.
@@
 \section{Theoretical Framework}
+Before the system features are discussed, notes about the design, and how it connects internally is needed to gain a better appreciation of the overall system
+
+\subsection{Application Architecture}
+The system's design is guided by the application architecture.
+Here it is discussed how an app that relies on networks functions and interacts \cite{kurose}.
+There are three main types: Client-Server, Peer-to-Peer, and Hybrid.
+Client-Server or CS means there is one web application that serves the data to multiple users (or clients, hence the name).
+As the number of clients grow bigger, the server takes more stress serving all the data.
+Peer-to-Peer or P2P is different since it has no singular dedicated server.
+No singular point of stress, and data is easily distributed.
+The term client and server is merged into a term "peer" that can both receive and send data.
+All peers can share data to other peers but the catch is that there is no centralized authority moderating these data, and that an active user base or peer base is needed for this to function.
+
+Hybrid tackles some of the issues of CS and P2P, which is scalability and moderation.
+It allows easier scaling of database access (akin to adding more servers in P2P), but it also allows to have a form of moderation and security (akin to adding a moderator to CS).
+This is also more appropriate for a distributed database, and which is why a hybrid architecture is used.
 
@@
 Only users with the directory's read key can download it, and only users with the write key can make changes.
-In short, Dat has "taken the best parts of Git, BitTorrent, and Dropbox".
+In short, Dat has "taken the best parts of Git, BitTorrent, and Dropbox"\cite{dat}.
 Another important feature is that files and directories are stored as-is, meaning no modifications are done by Dat when storing files.
@@
 \subsection{Networks}
+In each architecture it was discussed that there is a form of file transfer between peers or from servers to clients.
+To achieve this transfer, years of technology created systems called networks.
+There are 5 levels in a network \cite{kurose} but only the last 2, level 4 and 5 is needed for the system.
+The other levels 1-3 create protocols that send messages to specific and unique computers without error and utilizing any erratic medium between them (whether it be undersea cables or air).
+Level 4 assumes sending between computers is possible, but erratic.
+And these protocols in this level make sure all data is complete, error free before it is acknowledged.
+An important feature here is the File Transfer Protocol which is used by the NCBI.
+Here a client needs to set up a connection to the main NCBI server, then request specific files to be transferred to the client(or the user), and of course to move these files without error.
 
-An important aspect of the proposed system is the transfer of files to other computers.
-To achieve that, there are network technologies and systems that need to be understood.
-Each layer of the network has certain features, securities, and abstraction which contributes to the overall idea of sharing files across networks.
-Networks Layers are the abstraction of how computers pass data to each other.
-Understanding this is needed to maximize the speed in which data is transferred between computers.
-As well as to realize where certain problems or bottlenecks later on.
-A network, specifically the Internet, is comprised of the following levels of abstraction.
-\cite{kurose}
-\begin{enumerate}
-    \item Level 1: Medium used to send signals (physical, electrical signals of 0
-and 1), through wire or by air
-    \item Level 2: Introduction of packets: WIFI, Ethernet, connecting together
-different interfaces. Goal of sending a group of signals.
-    \item Level 3: Introduction of networks: IP Address to uniquely identify
-entities.
-    \item Level 4: Management of error free data transmission: Sending a group
-of messages via TCP, UDP, FTP (File Transfer Protocols)
-    \item Level 5: Abstraction of Level 4: Function calls programmers will use
-to do level 4 processes.
-\end{enumerate}
+Level 5 abstracts level 4, and making the process of connecting and transferring easier to accomplish.
+Since Level 5 uses function calls to do the protocols of level 4.
+A designer of a system will want the ease-of-use of using level 5 functions, but will need knowledge of level 4 in order to better understand bugs and problems later on.
+Level 5 also allows the user to focus more on doing similar database or data manipulation functions (C.R.U.D.) as POST, GET, PUT, DELETE.
 
-Level 1 through 3 focus on building a system for sending messages from one computer entity to another.
-Every level talks to another entity on the same level.
-And a higher level talks to the lower level.
-This later on will be important to understanding the scope of what the thesis will do.
-Level 1 and 2 can be vulnerable if anyone can read the data inside the network, partly why HTTPS was developed to encrypt the data.
-
-Level 4 and 5 is what the system will mainly use and operate on.
-In these systems, there are APIs commonly used to transfer files between computers on networks.
-These levels, compared to the levels earlier discussed also are more specific to the ideas of file transfer, checking if the files are error free, and more.
-Making sure files transferred are error free will be needed to do data synchronization later on.
-Another important network concept is that computers closer on the network (when talking directly) should have a faster speed and lower latency than computers farther apart on a network.
-
-In networks, when there is data being accessed over the network, it is not called C.R.U.D.
-(the way databases access data), instead it is called POST, GET, PUT, DELETE.
-POST is analogous to CREATE meaning the user submit data to be stored (in the server).
-GET is similar to READ but is in the context of reading a file over a network (from a server).
-PUT is analogous to UPDATE as DELETE remains as DELETE.
-These are codes used to denote network commands that do the same thing as the CRUD functions discussed earlier.
-This is important since there is an added layer of difficulty when it comes to transmitting data over a network.
+All information systems must have POST, GET, PUT, DELETE or the corollary (C.R.U.D.). Since there are networks in this system, the network version of the data manipulation functions will be used.
 
 \begin{itemize}
-    \item (Move to Theoretical Framework/caDDS) The system is an information system, and all information systems must have the main API’s 
+    \item The system being proposed is an information system, and this system will have the main API’s
     \begin{itemize}
         \item POST genome data
-        \item GET genome data (this will implement a hybrid P2P transfer)
-        \begin{itemize}
-            \item This will aim to speed up number 3 in the Problems of Classical Databases
-        \end{itemize}
+        \item GET genome data, this system will implement a hybrid P2P transfer
         \item PUT genome metadata (optional: PUT data)
@@
 \subsubsection{Overview}
-The two parts of the system infrastructure are the master node and data nodes. These are also inspired from SeqTorr.
+The two parts of the system infrastructure are the master node and data nodes. This system is also inspired by SeqTorr.
 
@@
 The most frequently used data, may be put on more data nodes increasing the speed of the data transfer (see Section \ref{theoreticalspeed}).
-Another is that the features on the designs of the speed, consistency, and fault tolerance (see Section \ref{cap} ) may be calibrated to the needs of the community.
+Another is that the features on the designs of the speed, consistency, and fault tolerance  may be calibrated to the needs of the community.
 
 \section{Conclusion}
+The research designed a theoretical design for a database for genomics researchers.
+The database is distributed and uses a hybrid architecture when managing the data.
+The theoretical properties of the design was studied and compared to a base method (FTP of NCBI), notably the speed and correctness.
+The proposed system, is better suited for database systems that scale.
+And is shown to be faster if there are multiple data nodes.
 
